### 2017-12-07 article.tex
Much!

\RequirePackage[l2tabu, orthodox]{nag}
\RequirePackage{silence}\WarningFilter{typearea}{Maybe no optimal type area settings}
\documentclass[version=last, pagesize, twocolumn, twoside=off, bibliography=totoc, 12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\input{preamble/draw}
\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}

%printer margin: I take 0.5 inches = 1,27 cm to be safe. 21 cm - (2 × 1.27 cm) = 18.46 cm. 29.7 cm − 0.9 cm − (2 × 1.27 cm) = 26.26 cm
\areaset{18.46cm}{26.26cm}
\setlength{\footskip}{9mm}
\begin{document}
\title{Simple models of deliberated preference}
\author{Olivier Cailloux}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

\section{Motivation}
We define the deliberated preference of an individual. We are interested in a context of a choice of a subset of alternatives from a set of possible alternatives given a priori. We define a model of the DP and give conditions and a validation procedure that ensure that what the model claims is in the DP is indeed. We want in particular to allow for the model to be applicable to various individuals that start with different knowledge: the model may use different counter-arguments to convince different individuals.

More in detail. A full-blown proof may be so long that nobody would read it. (That’s why mathematicians do not write their proofs in set-theoretic bare notation.) Hence, shortcuts are taken, which rely on the individual knowing stuff already. More generally, assume you want to convince someone of fact $F$ with a text (we call this an argument), in our case, we want to convince a person that such alternative from the set is a good choice for her. The argument may assume the reader knows already why such or such possible counter-arguments are invalid, and the argument will thus not delve into the details of some points. When shown in interactive mode (our context here), an individual may however question some points (which we call a counter-argument). The argumenter should then be able to provide an answer to the query (a counter-counter-argument).

This article aims at defining such a procedure and conditions of its validity.

\section{Decision situation}
Here is the definition of a decision situation.
\begin{itemize}
	\item All alternatives $\allalts$.
	\item Topic $\alltopic$ = $\Set{\prop_\alt, \alt \in \allalts} ∪ \Set{\prop_{¬\alt}, \alt \in \allalts}$. Denoted simply $\alt$ and $¬\alt$. We define $¬\prop$, with $\prop = \prop_\alt$, as equal to $\prop_{¬\alt}$ and $¬\prop$, with $\prop = \prop_{¬\alt}$, as equal to $\prop_{\alt}$.
	\item All possible arguments: $\allargs$. It contains at least $\emptyset$, the empty argument. It contains all the arguments and more. As an example, $\allargs$ could be the set of all strings (assuming all arguments can be transmitted textually).
\end{itemize}

Define $f_\prop$ from any unordered pair of arguments $(\ar, \ar')$ into $\powerset{\prop\text{-sure}, ¬\prop\text{-sure}, b)}$ (where $b$ means both possible). We present both arguments to $i$ and ask which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind. The set designates all cumulated answers over time, thus allowing $i$ to be unstable (because she has learnt new things in between two questioning, or just because of time passing, or for any known or unknown reason). We assume $b$ is observable as well.
We can also use $\emptyset$ as one or both of the arguments.

We define two binary relations $\prop^+$ (always supports $\propposs$) and $\prop^−$ (always supports $¬\propsure$) on $\allargs$. We define $(\ar, \ar') \in \prop^+ ⇔ (\ar, \ar') \ileadstopst \prop ⇔ \ar' \nibeatsst[\prop] \ar ⇔ \ileadsto(\ar, \ar') \subseteq \{\propsure, b\}$, and $(\ar, \ar') \in \prop^− ⇔ (\ar, \ar') \ileadstosst ¬\prop ⇔ \ar \nibeatsst[¬\propsure] \ar' ⇔ \ileadsto(\ar, \ar') = \{¬\propsure\}$.
Hence, $(\ar, \ar') \notin \prop^+ ⇔ (\ar, \ar') \ileadstose ¬\prop ⇔ \ar' \ibeatse[\prop] \ar ⇔ ¬(\ar' \nibeatsst[\prop] \ar) ⇔ \ileadsto(\ar, \ar') \supseteq \{¬\propsure\}$, and $(\ar, \ar') \notin \prop^− ⇔ (\ar, \ar') \ileadstope \prop ⇔ \ar \ibeatse[¬\propsure] \ar' ⇔ ¬(\ar \nibeatsst[¬\propsure] \ar') ⇔ \ileadsto(\ar, \ar') ≠ \{¬\propsure\}$.

When $\prop$ is fixed, or if it can be deduced from the output set, we write $(\ar, \ar') \ileadsto O$ or $\ileadsto(\ar, \ar') = O$ instead of $f_\prop(\ar, \ar') = O$.

Actually we will use only the $\ileadstopst$ and $\ileadstosst$ ternary relations from now on, together with the property (A1) that $\ileadstopst \subseteq \ileadstope$
\footnote{Equivalently: $\ileadstosst \subseteq \ileadstose$, $(\ar, \ar') \ileadstose ¬\prop ∨ (\ar, \ar') \ileadstope \prop$, $\ileadstopst \prop ⇒ ¬ (\ileadstosst ¬\prop)$, $\ileadstosst ¬\prop ⇒ ¬ (\ileadstopst \prop)$.} 
and (A2) that $\ileadstosst \subseteq \ileadstopst$ 
\footnote{Equivalently: $\ileadstose \subseteq \ileadstope$.}
and the fact that the relations $\prop^+$ and $\prop^−$ are symmetric. 
\footnote{Unclear where we use the fact that the relations $\prop^+$ and $\prop^−$ are symmetric.}

We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$.

$\ar$ is 
\begin{itemize}
	\item $\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$
	\item $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$
\end{itemize}

We can now define $\iposs, \isure \subseteq \alltopic$:
\begin{itemize}
	\item $\prop \in \iposs ⇔ \exists \ar \ \proppossdec$. 
\footnote{\label{ft:notposs}$\prop \notin \iposs ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstose ¬\prop$.}
	\item $\prop \in \isure ⇔ \exists \ar \ \propsuredec$. 
\footnote{\label{ft:notsure}$¬\prop \notin \isure ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstope \prop$.}
\end{itemize}

We say that severe $\prop$-unstability happens iff $\forall \ar, [\exists \ar^+ \suchthat (\ar, \ar^+) \ileadstose \prop] ∧ [\exists \ar^− \suchthat (\ar, \ar^−) \ileadstose ¬\prop]$.
\footnote{Severe $\prop$-unstability happens iff severe $¬\prop$-unstability happens.}
\footnote{Here is an example of severe $\prop$-unstability. $\forall (\ar, \ar'), (\ar, \ar') \ileadstose ¬\prop, (\ar, \ar') \ileadstose \prop$.}

\begin{theorem}
	If $\prop \notin \iposs ∧ ¬\prop \notin \iposs$, severe $\prop$-unstability happens.
\end{theorem}
The proof follows from \cref{ft:notposs}.

\begin{theorem}
	$\prop \in \isure ⇒ ¬\prop \notin \iposs$.
\end{theorem}
This is because $\ar^+$ is $\propsuredec$, thus given any $\ar^−$, $(\ar^+, \ar^−) \ileadstosst \prop$ hence $(\ar^+, \ar^−) \ileadstose \prop$ (using A1).

\begin{theorem}
	$\prop \in \isure ⇒ \prop \in \iposs$.
\end{theorem}
The proof follows from (A2).

\begin{definition}[Clear-cut]
	A situation is $\prop$-clear-cut iff $\prop \notin \iposs ⇒ ¬\prop \in \isure$.
\footnote{Here is an example of non $\prop$-clear-cut situation with no severe $\prop$-unstability. $\forall (\ar, \ar'), (\ar, \ar') \ileadstose ¬\prop, (\ar, \ar') \ileadstope \prop, (\ar, \ar') \ileadstopst ¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}
\end{definition}

\subsection{Possible outcomes}
We know $\prop \in \isure ⇒ ¬\prop \notin \iposs$ and $\prop \in \isure ⇒ \prop \in \iposs$.

Here are the remaining possibilities, considering the four pairs $(\prop, \iposs)$, $(\prop, \isure)$, $(¬\prop, \iposs)$, $(¬\prop, \isure)$. (A pair $(\prop, T)$ not mentioned means $\prop \notin T$.)
\begin{labeling}{Unstability, $¬\prop$ poss}
	\item[All poss] $\prop \in \iposs$, $¬\prop \in \iposs$
	\item[$\prop$ sure] $\prop \in \iposs$, $\prop \in \isure$
	\item[$¬\prop$ sure] (symmetric)
	\item[Unstability, $\prop$ poss] $\prop \in \iposs$ (oddity: $¬\prop \notin \iposs$ should imply $\prop \in \isure$)
	\item[Unstability, $¬\prop$ poss] (symmetric)
	\item[Strong unstability] (both oddities above)
\end{labeling}

\section{Models}
A model is a triple $(\mbeats, \mleadsto, +)$ defined as follows and satisfying the constraints as indicated here.

$\mbeats$ a binary relation over $\allargs$. $\mleadsto \subseteq \allargs × \alltopic$. Define $\clargs \subseteq \allargs$ as the set of arguments used in $\mbeats ∪ \mleadsto$. Let $+$ be defined over arguments used in the model: $\ar_3 + \ar_1 = \ar'$ for some $\ar' \in \clargs$, for any $\ar_3, \ar_1 \in \clargs$.

Requirements. The maximum length of a path in $\mbeats$ is finite (which implies that $\mbeats$ is acyclic).
$\ar_3 \mbeats \ar_2 \mbeats \ar_1 ⇒ \ar_2 \mbeats \ar_3 + \ar_1$.\footnote{Necessary for definition of $\ar_3 \wibeatse \ar_2$.}

Notation. Let $\mleadstoinv(\alltopic) \subseteq \clargs$ denote the subset of arguments supporting propositions. 
$\mbeats(\ar_2)$: arguments that $\ar_2$ attacks, $\ar_1 \in \mbeats(\ar_2) ⇔ \ar_2 \mbeats \ar_1$. We write $\args \mbeats \ar$ to mean that $\forall \ar' \in \args: \ar' \mbeats \ar$, and similarly for other binary relations.

Given a decision situation, define $\wibeatse[\prop], \nwibeatse[¬\prop] \subseteq \mbeats$ as follows.

Given $\ar_3 \mbeats \ar_2, \prop \in \alltopic$: $\ar_3 \wibeatse[\prop] \ar_2$ iff $[\exists \ar_1 \in \mbeats(\ar_2) \suchthat (\ar_2 \wibeatse[\prop] \ar_1 ∧ \ar_2 \nwibeatse[\prop] \ar_3 + \ar_1)] ∨ [\ar_2 \ileadsto \prop ∧ (\ar_3, \ar_2) \ileadstose ¬\prop]$.
\footnote{Or $\ar_2 \ileadsto \prop ∧ ¬\prop \notin \mleadsto(\allargs) ∧ (\ar_3, \ar_2) \ileadstope ¬\prop$.}

Given $\ar_3 \mbeats \ar_2, \prop \in \alltopic$: $\ar_3 \nwibeatse[\prop] \ar_2$ iff $[\exists \ar_1 \in \mbeats(\ar_2) \suchthat (\ar_2 \wibeatse[\prop] \ar_1 ∧ \ar_2 \wibeatse[\prop] \ar_3 + \ar_1)] ∨ [\ar_2 \mleadsto \prop ∧ (\ar_3, \ar_2) \ileadstope \prop]$.

This is well defined thanks to the following construction.
Given a model. 
\begin{itemize}
	\item Define a root as an argument that $\mbeats$-attacks nobody. 
	\item Associate to each $\ar \in \clargs$ its depth $d(\ar)$, the distance to the farthest root: $d(\ar) = 0$ iff $\ar$ is a root and $d(\ar) = k+1$ iff $\ar$ $\mbeats$-attacks some $\ar'$ of depth $k$ and $\mbeats$-attacks no $\ar'$ of depth superior to $k$ (thus $d(\ar) = 1$ iff $\ar$ attacks some and only root nodes; $d(\ar) = 2$ iff $\ar$ attacks some node of depth 1 and only nodes of depth 1 or 0; and so on). 
	\item Observe that if $\ar_3 \mbeats \ar_2$ and $d(\ar_3) = k+1$, then $d(\ar_2) ≤ k$. 
	\item Define $\wibeatse[\prop]_{(0)} = \nwibeatse[\prop]_{(0)} = \emptyset$. %as $\ar_3 \wibeatse[\prop]_{(1)} \ar_2$ iff $d(\ar_3) = 1 ∧ \ar_2 \ileadsto \prop ∧ (\ar_3, \ar_2) \ileadstose ¬\prop$. 
%	\item Define $\nwibeatse[\prop]_{(1)}$ as $\ar_3 \nwibeatse[\prop]_{(1)} \ar_2$ iff $d(\ar_3) = 1 ∧ \ar_2 \mleadsto \prop ∧ (\ar_3, \ar_2) \ileadstope \prop$. 
	\item Define $\wibeatse[\prop]_{(k+1)} \subseteq \mbeats$, $k ≥ 0$, as $\ar_3 \wibeatse[\prop]_{(k+1)} \ar_2$ iff, first, $d(\ar_3) = k+1$, and second, $[\exists \ar_1 \in \mbeats(\ar_2) \suchthat (\ar_2 \wibeatse[\prop]_{(d(\ar_2))} \ar_1 ∧ \ar_2 \nwibeatse[\prop]_{(d(\ar_2))} \ar_3 + \ar_1)] ∨ [\ar_2 \ileadsto \prop ∧ (\ar_3, \ar_2) \ileadstose ¬\prop]$. Proceed similarly for $\nwibeatse[\prop]_{(k+1)}$.
	\item Define $\wibeatse[\prop] = \bigcup_{k ≥ 0} \wibeatse[\prop]_{(k)}$, and similarly for $\nwibeatse[\prop]$.
\end{itemize}

{Check: Given $\ar_4 \mbeats \ar_3, \ar_3 \notin \mleadstoinv(\alltopic)$, with $\ar_2 \in \mbeats(\ar_3)$ ⇒ ($\ar_2$ and $\ar_4 + \ar_2$ $\mbeats$-attack only root nodes and do not support any proposition): $\ar_4 \wibeatse[\prop] \ar_3$ iff
\begin{itemize}
	\item $\exists \ar_2 \in \mbeats(\ar_3) \suchthat [(\ar_3 \wibeatse[\prop] \ar_2) ∧ (\ar_3 \nwibeatse[\prop] \ar_4 + \ar_2)]$ iff 
	\item $\exists \ar_2 \in \mbeats(\ar_3) \suchthat [(\exists \ar_1 \in \mbeats(\ar_2) \suchthat \ar_2 \wibeatse[\prop] \ar_1 ∧ \ar_2 \nwibeatse[\prop] \ar_3 + \ar_1) ∧ (\exists \ar_1 \in \mbeats(\ar_4 + \ar_2) \suchthat \ar_4 + \ar_2 \wibeatse[\prop] \ar_1 ∧ \ar_4 + \ar_2 \wibeatse[\prop] \ar_3 + \ar_1)]$ iff 
	\item $\exists \ar_2 \in \mbeats(\ar_3) \suchthat [(\exists \ar_1 \in \mbeats(\ar_2) \suchthat \ar_1 \ileadsto t ∧ (\ar_2, \ar_1) \ileadstose ¬\prop ∧ \ar_3 + \ar_1 \ileadsto \prop ∧ (\ar_3 + \ar_1, \ar_2) \ileadstope \prop ∧ (\exists \ar_1 \in \mbeats(\ar_4 + \ar_2) \suchthat \ar_1 \mleadsto \prop ∧ (\ar_4 + \ar_2, \ar_1) \ileadstose ¬\prop ∧ \ar_3 + \ar_1 \ileadsto \prop ∧ (\ar_4 + \ar_2, \ar_3 + \ar_1) \ileadstose ¬\prop)]$.
\end{itemize}
}

Given $\ar_3 \in \clargs, \ar_2 \in \clargs$, with $\exists \ar_1 \in \mbeats(\ar_2) \suchthat \ar_2 \wibeatse[\prop] \ar_1$, we have: $\ar_3 \wibeatse[\prop] \ar_2 ∨ \ar_3 \nwibeatse[\prop] \ar_2$.

Define $\ar_2 \wibeatse \ar_1 ⇔ \exists \prop \in \alltopic \suchthat \ar_2 \wibeatse[\prop] \ar_1$.

Hence, given $\ar_3 \in \clargs, \ar_2 \in \clargs, \ar_2 \notin \mleadstoinv(\alltopic)$: $¬ (\ar_3 \nwibeatse \ar_2)$ iff $\forall \ar_1 \in \mbeats(\ar_2) ∩ \wibeatse(\ar_2): ¬ (\ar_2 \wibeatse \ar_3 + \ar_1)$.

\section{Conditions}
All these conditions assume that a decision situation $(\allalts, \allargs, \{\ibeatse[\prop]\}, \{\ibeatse[\prop, \text{sure}]\})$ and a model $\eta = (\mbeats, \mleadsto, +)$ are given.

Define $\argsdec = \clargs \setminus \text{im}(\wibeatse)$ the decisive arguments according to $\wibeatse$, or $\wibeatse$-decisive arguments for short: $\ar \in \argsdec ⇔ \wibeatseinv(\ar) = \emptyset$.

\begin{definition}[Reinstatement]
	Given $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1, \ar_3 \in \argsdec$: $\mbeats(\ar_1) \subseteq \mbeats(\ar_3 + \ar_1)  ∧ \mbeatsinv(\ar_3 + \ar_1) \subseteq \mbeatsinv(\ar_1) \setminus \mbeats(\ar_3)$.
	\footnote{TODO the condition must be $\wibeatseinv(\ar_3 + \ar_1) \subseteq \mbeatsinv(\ar_1) \setminus \mbeats(\ar_3)$ to allow $\ar_5 \mbeats \ar_4 \mbeats \ar_3 \mbeats \ar_2 \mbeats \ar_1$ and $\ar_5 \mbeats \ar_4 \mbeats \ar_3 + \ar_1$, considering that possibly $\ar_3$ is $\wibeatse$-decisive. This should not invalidate the conditions, but it does currently. But it’s not a problem: the model would actually not be built this way. In this scenario the argument $\ar_3 + \ar_1$ is useful only in case $\ar_3$ is decisive, thus $\ar_4 \mbeats \ar_3 + \ar_1$ must not be planned. Rather $\ar_5 + \ar_3$ decisive, then $(\ar_5 + \ar_3) + \ar_1$. Alternatively, also $\ar_4 \mbeats \ar_1$ and then no problem as well.}
	\footnote{The stronger condition mandating $\mbeatsinv(\ar_3 + \ar_1) \subseteq \wibeatseinv(\ar_1) \setminus \mbeats(\ar_3)$ would be more difficult to check: when some $\ar_2 \wibeatse \ar_3 + \ar_1$, we’d need to check not only that $\ar_2 \mbeats \ar_1$ but also $\ar_2 \wibeatse \ar_1$.}
	\footnote{We do not mandate that $\ar_3 + \ar_1 \ileadsto \prop$, so that the model can afford not resisting to the counter-attacks to $\ar_3 + \ar_1$ (resistance to c-a to $\ar_1$ suffice). We need: Obs applies to restricted supports (one per prop decided by model); Covering applies to extended supports (restricted supports plus those obtained by reinstatement). Replacement-1 applies to all and requires attack at least as large; Replacement-2 applies to restricted supports and requires no new $\ibeatse$-attacks.}
\end{definition}

\begin{definition}[Justifiable unstability]
	$\forall \ar_2 \mbeats \ar_1 \suchthat \ar_2 \wibeatse \ar_1, \ar_2 \nwibeatse \ar_1: \exists \ar_3 \mbeats \ar_2 \suchthat \ar_3 \wibeatse \ar_2$.
\end{definition}

\begin{definition}[Finite defense]
	If $\wibeatseinv(\ar) \subseteq \wibeatse(\argsdec)$, then $\exists \args \subseteq \argsdec, \card{\args} ≤ j \suchthat \mbeatsinv(\ar) \subseteq \wibeatse(\args)$.
\footnote{To satisfy Finite defense, in presence of the other conditions, suffice to limit the width of the model (TODO check). But it may be interesting to not limit it and declare that the model has specific replies to any counter-argument, but promises to use only a few rebuttals and that afterwards, the dm will stop using those kind of arguments (but we don’t know in advance which ones will be chosen).}
\end{definition}
Thus, if the attackers of $\ar$ are attacked by decisive arguments, then $j$ defenders are enough to defend $\ar$.

Define $R$, the reinstates relation, as follows: $\ar_3 R \ar_1$ iff $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1$ (for some $\ar_2$), $\ar_3 \in \argsdec$. Define $\gdargs$ as the transitive closure of $\mleadstoinv(\alltopic)$ under $R$.
\begin{definition}[Covering]
	$\forall \ar \in \gdargs, \ar' \in \allargs: \ar' \ibeatse \ar ⇒ \ar' \mbeats \ar$. \footnote{Specify $\ibeatse$.}
\end{definition}

\begin{definition}[Observable validity]
	$\forall \ar_2 \mbeats \ar_1 \mleadsto \prop: ¬(\ar_2 \wibeatse[\prop] \ar_1) ∨ \exists \ar_3 \mbeats \ar_2 \suchthat \ar_3 \wibeatse[\prop] \ar_2$. Furthermore, if $¬(\clargs \mleadsto \prop), \forall \ar_1 \mleadsto ¬\prop, \ar \in \allargs: \ar_1 \ibeatse[\prop] \ar$.
	\footnote{If the model claims $¬\prop \in \isure$, this requires clear-cut (for that prop), so we must mandate it (hopefully A3 or an equivalent such as Justifiable unstability fits). Thus we only need to prove $\prop \notin \iposs$, for which $\ar_1 \ibeatse[\prop] \ar$ suffices.}
\end{definition}

\section{Theorem}
\begin{theorem}[Validity]
	Given a decision situation and a model $\eta$, if all our conditions are satisfied, $\mleadsto(\clargs) \subseteq \iposs$. Furthermore, if $¬\prop \in \mleadsto(\clargs) ∧ \prop \notin \mleadstoinv(\clargs)$, $¬\prop \in \isure$.
\end{theorem}
\begin{proof}
$\ar$ is defended iff its $\wibeatse$-attackers are $\wibeatse$-attacked by $\wibeatse$-decisive arguments.

First, we want to prove that $\ar_1$ defended implies $\ar_1$ replaceable by some $\wibeatse$-decisive $\ar$, and if $\ar_1 \in \gdargs$, then its replacer $\ar$ is in $\gdargs$ as well.

By hypothesis, $\wibeatseinv(\ar_1) \subseteq \wibeatse(\argsdec)$. Thus, $\exists \args \subseteq \argsdec \suchthat \mbeatsinv(\ar_1) \subseteq \wibeatse(\args)$, $\args$ finite [Finite defense].

Pick any $\ar_{3, 1} \in \args$ such that $\ar_{3, 1} \wibeatse \ar_2 \wibeatse \ar_1$ (if there’s none, $\ar_1 \in \argsdec$ and we’re done). $\ar_{3, 1} + \ar_1$ replaces $\ar_1$, and $\mbeatsinv(\ar_{3, 1} + \ar_1) \subseteq \mbeatsinv(\ar_1) \setminus \mbeats(\ar_{3, 1})$ [Reinstatement]. Hence, $\mbeatsinv(\ar_{3, 1} + \ar_1) \subseteq \wibeatse(\args) \setminus \mbeats(\ar_{3, 1})$. Iterate by picking any $\ar_{3, 2} \in \args$ such that $\ar_{3, 2} \wibeatse \ar_2 \wibeatse \ar_{3, 1} + \ar_1$ (if there’s none, $\ar_{3, 1} + \ar_1 \in \argsdec$ and we’re done) and obtaining $\ar_{3, 2} + (\ar_{3, 1} + \ar_1)$ replacing $\ar_{3, 1} + \ar_1$ (hence, replacing $\ar_1$) with $\mbeatsinv(\ar_{3, 2} + (\ar_{3, 1} + \ar_1)) \subseteq \mbeatsinv(\ar_{3, 1} + \ar_1) \setminus \mbeats(\ar_{3, 2})$. Hence, $\mbeatsinv(\ar_{3, 2} + (\ar_{3, 1} + \ar_1)) \subseteq \wibeatse(\args) \setminus \mbeats(\ar_{3, 1}) \setminus \mbeats(\ar_{3, 2})$. Iterating in such a way over the finite set $\args$ will finally yield an element that is $\wibeatse$-decisive. The last point, $\ar_1 \in \gdargs ⇒ \ar \in \gdargs$, follows from the definition of $\gdargs$.

Second, we want to prove that if $\ar_1$ not defended and has no decisive $\wibeatse$-attackers (meaning that $\wibeatseinv(\ar_1) \subseteq \overline{\argsdec}$), then $\ar_1$ is $\wibeatse$-attacked by some $\ar_2$ that is not defended and has no decisive $\wibeatse$-attacker.

Consider $\ar_1$ not defended and having no decisive $\wibeatse$-attackers. Because $\ar_1$ is not defended, by definition, it is $\wibeatse$-attacked by some $\ar_2$ that has no decisive $\wibeatse$-attacker. Because $\ar_1$ has no decisive attacker, $\ar_2$ is not decisive. If $\ar_2$ was defended, by the first part of this proof, it would be replaceable by a decisive argument, and $\ar_1$ would have a decisive attacker. Thus, $\ar_2$ is not defended.

Third, consider an argument $\ar_1 \in \mleadstoinv(\alltopic)$. It has no decisive $\ibeatse$-attacker: as $\ar_1 \in \gdargs$, any $\ibeatse$-attack is a $\wibeatse$-attack [Covering], and $\ar_1$ has no decisive $\wibeatse$-attacker [Obs val]. Also, $\ar_1$ is defended: assume it is not, then by our second point in this proof some $\ar_2 \wibeatse \ar_1$, with $\ar_2$ not defended and with no decisive $\wibeatse$-attacker, and iterating and using finiteness of $\wibeatse$ leads to a contradiction. Hence, by our first point in this proof, $\ar_1$ is replaceable by some $\wibeatse$-decisive $\ar \in \gdargs$. As $\ar \in \gdargs$, any $\ibeatse$-attack is a $\wibeatse$-attack [Covering], thus $\ar$ is $\ibeatse$-decisive.
\end{proof}

\appendix
\section{Todo}
\begin{itemize}
	\item If $i$ does not consider $\ar$ as supporting $\prop$, it also works: if $\prop$ is not weakly acceptable by default, then any $\ar'$ is considered by $i$ as a better argument than $\ar$ in favor of certain $¬\prop$, and so on. In fact, whether $\emptyset \ibeatse[\prop] \emptyset$ determines whether $\prop$ is weakly supported by default.
	\item I should define $\ar' (□\ibeatse[\prop]) \ar$ as an observable: “Assuming $\ar'$ would survive, do you consider $\ar'$ as leading to certainty of $¬\prop$, even when considering $\ar$?”. It distinguishes our knowledge and the truth: $\ar' (□\ibeatse[\prop]) \ar ⇒ \ar' \ibeatse[\prop] \ar$, thus, implies $¬ (\ar' \nibeatsst[\prop] \ar)$. But out of $¬ (\ar' (□\ibeatse[\prop]) \ar)$, nothing.
	\item Partition (objectively) $\allargs$ (or $\allargs × \alltopic$) into arguments in favor of $\prop, \text{sure}$, $¬\prop, \text{sure}$, and similarly for possible. Use only one rel $\ibeatse$, defined on contradictory arguments only, instead of $\ibeatse[\prop, \text{sure}]$ and others. Define $\ar' \ibeatse[\prop] \ar$ equals no when $¬(\ar' \ileadsto ¬\prop, \text{sure})$, equals $\ibeatse$ for adequate arguments, and equals yes when $¬(\ar \ileadsto \prop, \text{possible})$ and $\ar' \ileadsto ¬\prop, \text{sure}$, with probably some complications needed for the argument $\emptyset$ (and related default attitude towards $\prop$).
\end{itemize}

Questions:
Q1. Relationship with $\ar \ibeatse[\prop] \ar$?

We want to exclude: $s$ supports $p$ perhaps, attacked by $s2$ (supporting $¬p$ sure), but then $s2$ is attacked by $s$. Exclude $\ar' \ibeatse[\prop] \ar$ and $\ar \ibeatse[¬\prop, \text{sure}] \ar'$. Require to assume that this situation implies another argument $\ar_3$ “attacking” $\ar'$, thus, such that $\ar_3 + \ar$ is no more attacked by $\ar_2$.

\subsection{Attack rel}
Given $\prop$, define a (symmetric) strong attack rel, $Q_\prop$, as $\ar Q_\prop \ar'$ iff $(\ar, \ar') \ileadstosst ¬\prop$.

Define a (symmetric) weak attack rel, $R_\prop$, as $\ar R_\prop \ar'$ iff $(\ar, \ar') \ileadstopst ¬\prop$.

Given $\prop$, we want to check the arguments that defend it. We want to build an asymmetric attack relation.
$\ar_1 M \ar_0$, and not the converse, iff $\ar_1$ attacks $\ar_0$ and, would $\ar_1$ be attacked, $\ar_0$ reinstated would defend $\prop$.
$\ar_1$ should be able to support a sub-claim $\prop'$, and then $\ar_2$ attack $\ar_1$ on that sub-claim precisely. $\ar_2$ decisive on this sub-claim and $\ar_0 + \ar_2$ decisive implies $\ar_1$ attacks $\ar_0$ and not conversely.

Or simply. $\ar_1$ decisive for $¬\prop$ means $\ar_1$ attacks $\ar_0$ and not conversely.

Then define $i$’s DJ in terms of argumentation semantics.

\section{To think}
Propositions weakly self-supported $T \subseteq \alltopic$: weakly accepted if no arg is given. Examples: $m$ = “eat miam”; $¬b$ = “beurk is to exclude”; or, in a problem where there’s no particularly good aliments, both $a$ = “eat this” and $¬a$.

When given $(\ar, \prop)$, $i$ may say: $\ar$ does not survive; or: assuming $\ar$ survives, then $\ar$ supports $\prop$, or, assuming $\ar$ survives, then $\ar$ does not support $\prop$ anyway.

When given $\ar'$ against $\ar$, $i$ may say: $\ar'$ does not survive, or: assuming $\ar'$ survives, then $\ar'$ supports $¬\prop$, …

Given $(\ar_2, \prop), (\ar_1, ¬\prop) \in D$, define $¬(\ar_2 \wibeatse^\text{neg}_{¬\prop} \ar_1)$ iff for some $(\ar, \prop) \in D$, where $\ar_1 \ibeatse[\prop] \ar$: $\ar_1 \ibeatse[\prop] \ar + \ar_2$. Equivalently: $\ar_2 \wibeatse^\text{neg}_{¬\prop} \ar_1$ iff for all $\ar$, where $\ar_1 \ibeatse[\prop] \ar$: $¬(\ar_1 \ibeatse[\prop] \ar + \ar_2)$. (This does not seem right: if given $\ar_3$ attacking $\ar_2$, and not given $\ar_4$ which would convincingly rebut $\ar_3$, then temporarily it may hold again that $\ar_1 \ibeatse[\prop] \ar + \ar2$ (in the sense that $\ar_1 + \ar_3 \ibeatse[\prop] \ar + \ar_2$).)

$\ar_2 \wibeatse_{¬\prop} \ar_1$ can perhaps be queried directly by asking (in the context of some $\ar_1 \ibeatse[\prop] \ar$): “assume $\ar_2$ survives, then does $\ar_2$ counter $\ar_1$?” (In the sense that $\ar_2$ is sufficiently convincing that $\prop$ holds perhaps, to cancel the argument $\ar_1$ according to which $¬\prop$ surely holds.)

\section{Certainties}
Looking for certainties. Those propositions that are in the reflexive preferences in a demanding sense: there is a strong enough reason to prefer it than its contrary.
\begin{itemize}
	\item $\ar' \wibeatse \ar$: weak attack; $\ar'$ renders $\ar$ invalid (can’t be used to say that $t$ holds for sure) (assuming $\ar'$ survives)
	\item Propositions strongly self-supported: strongly accepted if no arg is given. Examples: $m$ = “eat miam”; $¬b$ = “beurk is to exclude”. We might have neither $c$ nor $¬c$ in that set.
\end{itemize}

\begin{definition}[Sure acceptance]
	Define a situation $(\allalts, \allargs, \set{\ibeatse[\prop]})$. A proposition $\prop \in \alltopic$ is accepted as sure iff $\exists \ar' \in \allargs \suchthat \forall \ar \in \allargs: \ar \nibeatsst[\prop, \text{sure}] \ar'$.
\end{definition}

Assume we use rather: if $p$ is not sure, then $¬p$ is weakly accepted (by def). Then we have never problems of inconsistency! But we could be in a situation where $p$ is not accepted as sure but nobody can tell why because it is fundamentally unstable (sometimes $p$ being accepted, sometimes not).

\section{Example about model instanciation}
The general conditions are Reinstatement, Justifiable unstability, Finite defense and Covering.
A general model is a model that claims it satisfies the general conditions.

TODO give up general models. In this example, $\ar_1$ would need to be planned as attacking sometimes $\ar_2$. Better consider an instanciation mechanism. An instantiated model is particular, and can be tested (especially against another one).
\begin{example}
	$\ar_3 \mbeats \ar_2 \mbeats \ar_1 \mleadsto \prop$, $\ar_2 \mleadsto ¬\prop$; $\ar_3 + \ar_1 \mleadsto \prop$.
\end{example}
This model is compatible (meaning that it satisfies the general conditions) with the following decision situations. We describe $\ibeatse$ fully (no attack iff not mentioned).
\begin{itemize}
	\item Sure of $\prop$: $\ar_1 \mleadsto \prop$; $\ar_3 \ibeatse \ar_2$ (the rest is implied, for example $\forall \ar_4 \in \allargs: \ar_1 \ibeatse[¬\prop] \ar_4$ because of covering).
	\item Sure of $\prop$ with reinstatement: $\ar_3 \ibeatse[\prop] \ar_2 \ibeatse[\prop] \ar_1 \mleadsto \prop$; $\ar_1 + \ar_3 \mleadsto \prop$; $\ar_3 \ibeatse[¬\prop] \ar_2$
	\item Sure of $¬\prop$: $\ar_2 \mleadsto ¬\prop$; $\ar_2 \ibeatse \ar_1$
	\item Both: $¬(\ar_2 \ibeatse \ar_1), \ar_1 \mleadsto \prop$, $¬(\ar_3 \ibeatse \ar_2), \ar_2 \mleadsto ¬\prop$
\end{itemize}
This situation falsifies the model. $\ar_4 \ibeatse \ar_1$, $\ar_4$ not attacked.

\section{Model certainties}
Assume we define $\ar_1 \wibeatse[¬\prop, \text{sure}] \ar_2$ ⇔ $\ar_2 \nwibeatse[\prop] \ar_1$. Then, indeed, given $\ar_1 \mleadsto \prop$, $\ar_2 \wibeatse[¬\prop, \text{sure}] \ar_1$ ⇔ $\ar_2 \ibeatse[¬\prop, \text{sure}] \ar_1$. But it gives the wrong conclusion. For $\ar_3 \mbeats \ar_2 \mbeats \ar_1 \mleadsto t: \ar_3 \wibeatse[\prop, \text{sure}] \ar_2$ iff $\ar_2 \nwibeatse[¬\prop] \ar_3$ iff $\exists \ar_1 \in \mbeats(\ar_3) \suchthat \ar_3 \wibeatse[¬\prop] \ar_1 ∧ \ar_3 \wibeatse[¬\prop] \ar_2 + \ar_1$.

Given $\ar_1 \mleadsto \prop$, define $\ar_2 \wibeatse[\prop, \text{sure}] \ar_1$ iff $\ar_2 \ibeatse[\prop, \text{sure}] \ar_1$.

Given $\ar_1 \mleadsto \prop$, define $\ar_2 \nwibeatse[\prop, \text{sure}] \ar_1$ iff $\ar_1 \ibeatse[¬\prop] \ar_2$.

Given $\ar_3 \in \clargs, \ar_2 \in \clargs, \ar_2 \notin \mleadstoinv(\alltopic), \prop \in T$: $\ar_3 \wibeatse[\prop, \text{sure}] \ar_2$ iff $\exists \ar_1 \in \mbeats(\ar_2) \suchthat \ar_2 \wibeatse[\prop, \text{sure}] \ar_1 ∧ \ar_2 \nwibeatse[\prop, \text{sure}] \ar_3 + \ar_1$.

Given $\ar_3 \in \clargs, \ar_2 \in \clargs, \ar_2 \notin \mleadstoinv(\alltopic), \prop \in T$: $\ar_3 \nwibeatse[\prop, \text{sure}] \ar_2$ iff $\exists \ar_1 \in \mbeats(\ar_2) \suchthat \ar_2 \wibeatse[\prop, \text{sure}] \ar_1 ∧ \ar_2 \wibeatse[\prop, \text{sure}] \ar_3 + \ar_1$.

\section{Example about default arguments}
s2 argues in favor of p against s1: s "le monde n’est pas fiable". s1 "le monde est fiable, bhl l’a dit". s2 "bhl est un clown, il s’est planté sur l’Irak". s3 "il avait raison sur l’Irak : l’Irak a des ADM". s4 "l’Irak n’a pas d’ADM, Bush l’a reconnu".
Does s4 attack s3?
"bhl est un clown, il s’est planté sur l’irak" + "l’irak n’a pas d’ADM, Bush l’a reconnu" VS "il avait raison sur l’Irak : l’Irak a des ADM" !

Measure problem?

\section{Alternative definitions of finite defense}
Define Finite defense-$\wibeatse$-$\wibeatse$-$\mbeats$-dec as: $\wibeatseinv(\ar) \subseteq \wibeatse(\clargs \setminus \text{im}(\mbeats)) ⇒ \wibeatseinv(\ar) \subseteq \wibeatse(\args)$. Finite defense-$\wibeatse$-$\wibeatse$-$\mbeats$-dec is insufficient to provide $T_\eta = \iposs$. Define $\ar' \mbeats^\text{fail} \ar$ iff $\ar' \mbeats \ar ∧ ¬(\ar' \wibeatse \ar)$. Consider $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1$, $\ar_3' \wibeatse \ar_2' \wibeatse \ar_1$, and so on, and $\ar_4 \mbeats^\text{fail} \{\ar_3, \ar_3', …\}$. Then I really need infinitely many arguments to defend $\ar_1$ but Finite defense-$\wibeatse$-$\wibeatse$-$\mbeats$-dec is artificially satisfied because the antecedent fails to trigger.

Define Finite defense-$\wibeatse$-$\wibeatse$-subsets as: $\wibeatseinv(\ar) \subseteq \wibeatse(\args) ⇒ \wibeatseinv(\ar) \subseteq \wibeatse(\args')$. Finite defense-$\wibeatse$-$\wibeatse$-subsets is insufficient to provide $T_\eta = \iposs$. This is because Reinstatement allows for new attacks in $\wibeatse$ (it only forbids new attacks in $\mbeats$), thus we can forever transform previously failing attacks to new attacks, hence always satisfying Finite defense (always finite cover of $\wibeatse$, but infinite cover of $\mbeats$) but still not converging. Consider $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1, \ar_3' \wibeatse \ar_2' \mbeats^\text{fail} \ar_1$, and so on; and $\ar_3' \wibeatse \ar_2' \wibeatse \ar_3 + \ar_1, \ar_3'' \wibeatse \ar_2'' \mbeats^\text{fail} \ar_3 + \ar_1$, and so on.

Define Finite defense-$\mbeats$-$\wibeatse$-startdec as: $\mbeatsinv(\ar) \subseteq \wibeatse(\argsdec) ⇒ \mbeatsinv(\ar) \subseteq \wibeatse(\args)$. Finite defense-$\mbeats$-$\wibeatse$-startdec is insufficient to provide $T_\eta = \iposs$. Consider $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1, \ar_3' \wibeatse \ar_2' \wibeatse \ar_1$, and so on, and $\ar_5 \mbeats^\text{fail} \ar_4 \mbeats^\text{fail} \ar_1$. Then I really need infinitely many arguments to defend $\ar_1$ but Finite defense-$\mbeats$-$\wibeatse$-startdec is satisfied as there is no cover of the $\mbeats$ attacks to $\ar_1$.

Define Finite defense-$\mbeats$-$\mbeats$-startdec as: $\mbeatsinv(\ar) \subseteq \mbeats(\argsdec) ⇒ \mbeatsinv(\ar) \subseteq \mbeats(\args)$. Finite defense-$\mbeats$-$\mbeats$-startdec is insufficient to provide $T_\eta = \iposs$. Consider $\ar_3 \wibeatse \ar_2 \wibeatse \ar_1$, $\ar_3' \wibeatse \ar_2' \wibeatse \ar_1$, and so on, and $\ar \mbeats^\text{fail} \{\ar_2, \ar_2', …\}$. Then I really need infinitely many arguments to defend $\ar_1$ but Finite defense-$\mbeats$-$\mbeats$-startdec is artificially satisfied because of $\ar$. Define Finite defense-$\mbeats$-$\mbeats$-subsets as: $\mbeatsinv(\ar) \subseteq \mbeats(\args) ⇒ \mbeatsinv(\ar) \subseteq \mbeats(\args')$ (for any $\args \subseteq \argsdec$). Finite defense-$\mbeats$-$\mbeats$-subsets is (rightly) non satisfied in this example.

Define Finite defense-$\wibeatse$-$\wibeatse$-subsets as: $\wibeatseinv(\ar) \subseteq \wibeatse(\args) ⇒ \wibeatseinv(\ar) \subseteq \wibeatse(\args')$. Finite defense-$\mbeats$-$\wibeatse$-subsets $⇏$ Finite defense-$\wibeatse$-$\wibeatse$-subsets. Consider $\ar_2 \mbeats^\text{fail} \ar_1$ (to be continued…)

%\bibliography{mybib}

\end{document}