### 2023-11-30 simple.tex
Much new stuff

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=last, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\usepackage[normalem]{ulem}
%\input{preamble/draw}
%\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}
\DeclareMathAlphabet{\mathup}{OT1}{\familydefault}{m}{n}

%I find these settings useful in draft mode. Should be removed for final versions.
	%Which line breaks are chosen: accept worse lines, therefore reducing risk of overfull lines. Default = 200.
		\tolerance=2000
	%Accept overfull hbox up to...
		\hfuzz=2cm
	%Reduces verbosity about the bad line breaks.
		\hbadness 5000
	%Reduces verbosity about the underful vboxes.
		\vbadness=1300

\begin{document}
\title{Theories of deliberated choice}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
  \href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
\hypersetup{
  pdfsubject={Epistemology},
  pdfkeywords={Decision aiding, Decision making, Argumentation}
}
\makeatother
\maketitle

\begin{abstract}
  To be written.
\end{abstract}

\section{Introduction}
Observing subjects perform acts of choice is the basis of much work in econometrics. Such acts have been used, among others, to give an observable basis to the concept of preference \citep{samuelson_foundations_1983}. %Acts of choice are considered here in the broad sense of choice between courses of actions, encompassing the simple setting of choosing between various elementary consumption bundles as well as settings involving more abstract acts such as choosing a strategy in a game-theoretical setting.

Most of these approaches consider only acts of choice that are performed spontaneously in the normal course of action of an individual, or that are given a specific frame depending on the psychological aspect that is to be studied. As an example of the former kind, typical willingness to pay studies asks the subject whether she prefers an object at some price or another object at another price, describing the objects in a way that is considered immediately understandable by the subject (or simply showing the objects). Examples of the latter kind include the celebrated work of Kahneman and Tversky \citep{bell_descriptive_1988, kahneman_thinking_2012}.

On the other hand, since at least the seminal work of \citet{fishkin_when_2011}, another sort of judgment involving subjective desirability is emerging as also worth studying. These are judgments given by subjects after arguments in favor and against different possible stance have been considered. Such judgments may differ from immediate judgments of desirability given in natural conditions. Indeed, individuals may change opinion on the desirability of some course of action, and therefore revise their choice, while learning about the properties of objects, the consequences of some acts (or empirical knowledge useful to estimate the likelihood of some consequences), or the logical or empirical impossibilities between consequences. This is particularly likely to happen when choosing among non everyday objects or acts whose consequences are not well known by the subject.

While numerous articles have discussed the concept of deliberation in the last decades, it remains unsettled, to the best of my knowledge, how to observe a deliberated judgment in a systematic and reproducible fashion. There is a need to define this concept sufficiently precisely that, first, it is clear what counts as a deliberated judgment (i.e., when “sufficient” \citep{meinard_justification_2020} or “correct” deliberation has happened); and second, it is clear which part of the claim is empirical and which part stems from an unfalsifiable conception of what “deliberated” means. This permits to study deliberated judgments empirically, meaning here that disagreements about someone’s deliberated judgment in a given context can be solved, at least in principle, by an empirical experiment, provided the very conception of what constitutes a deliberated judgment is shared. This article is interested in defining the concept of deliberated judgement of a subject and show how one can study it in an empirical manner. The “classical” counterpart of deliberated judgments, that does not involve confrontation with arguments, will henceforth be called the natural judgment. As we deal with choice acts, we will also name the concept “deliberated choice”, as opposed to “natural choice”.

While the concept of deliberation in the deliberative democracy literature usually refers to multiple individuals debating, this article considers deliberation as related to a single individual: an individual deliberates (in the sense of: carefully weights) the stance she considers most appropriate while being confronted to arguments. The deliberated judgment of an individual is thus defined with no requirement that the arguments be uttered by peers in a debate; they can be given by an automaton, for example. In this sense, deliberated judgments come close to what \citet{rawls_theory_1999} calls reflective equilibrium (referring to an idea of \citet{goodman_fact_1983}), a concept that has much inspired the present proposal.

To define deliberated judgments, one needs to start with the set of arguments under consideration, and the way the subject gets to know them, called hereafter the “exposure protocol”. As a result, we need always talk about the deliberated judgment \emph{given some exposure protocol}. The notion of argument invoked here is an extremely large one, including texts, images, sounds, experiences (a hiking trip), and so on: anything that can be transmitted to a subject qualifies as a possible argument. In particular, such an “argument” need not have the logical structure of what is considered a proper argument in argumentation theory. This important feature of this proposal permits to avoid taking a position on what is a correct argument, and permits to study wide enough conceptions of deliberated judgment, including those involving no paternalism \citep{cailloux_formal_2020}: anything that influences the judgment of a subject, as judged by the subject herself and not by any external standard, may a priori be considered worth including in at least some conceptions of deliberated judgment, thus including whatever some could consider as “incorrect reasoning” or “bad reasons”.

The dependence of the deliberated judgment concept on the exposure protocol yields a first qualification about the concept of deliberated judgments when one desires to use it normatively: it is only as reasonable to call the resulting judgments “deliberate” in a normative sense as the exposure protocol is normatively relevant to deliberation. A sufficiently restricted, or even purposefully chosen, set of arguments may yield judgments that are only deliberate in the technical sense used here, but that pertain more to brain-washing than to reflective judgments. While it is indeed a long-term goal of this research program to yield a normatively relevant conception of preference (appropriate, for example, for recommendation), it is not a claim of this article that anything that corresponds to deliberated judgments as defined here qualify as being appropriate for any normative use: much depends on the chosen exposure protocol. The goal of this article is one of a conceptual and practical clarification. It proposes a sharp distinction between the empirical content and the normative content of the concept of deliberated judgments: the definition proposed here transfers the full normative weight on the choice of the exposure protocol; and leaves the rest of the disagreements to empirical settling. This is methodologically akin, mutatis mutandis, to the axiomatic study of justice such as done in social choice, only complementing mathematical deduction with empirical claims. In social choice, the choice of axioms bears the full normative weight; once axioms are chosen, it is a matter of logic which social choice rules (if any) correspond to that choice. In our case, once an exposure protocol is chosen, it is an empirical matter which deliberated judgments result.
In supplement to a possibly helpful conceptual clarification, such a sharp separation may reveal practically useful for separation of concern as it permits to study deliberated judgments empirically under various exposure protocols independently of possible disagreements about which norms are more appropriate for deliberation in which circumstances.

The deliberated judgment of a subject is said to depend on the exposure protocol, that is, on the set of arguments and the way they are submitted to the subject, and not only on the set of arguments, because the very way that they are transmitted to the individual may affect her judgment differently. For example, reading an image description and showing an image to a subject may impress them differently \citep{railton_facts_2003}.

The exposure protocol defines a set of argument and, importantly, how arguments are sent to the subject and her subsequent choice observed. There is no explicit dependency on the internal process that occurs in the subject’s mind while she (possibly) processes the arguments, as it is not desirable to assume that the internal process could be observed. We consider subjects as black boxes and only require a capacity of observation of their choices after exposure to arguments. Depending on the exposure protocol and the subject, it is of course possible for the arguments to never affect any choice of the subject, a conclusion that may be interesting in its own right.

The deliberated judgment of a subject is defined as a function of the exposure protocol, as accepting every claims that resists every counter-arguments. The notion of resistance is itself defined according to the exposure protocol.

Studying judgments resulting from exposure to arguments includes a difficulty that is not present with natural choice experiments: exposing an individual to arguments may change the individual’s stance, an effect that cannot be erased in order to submit the individual to an unrelated sequence of arguments. The observer thus cannot in general assume that multiple unrelated sequences of arguments can be tested on a given individual. This difficulty will be worked around by considering general theories of deliberated judgments, that apply to sets of individuals. This is methodologically similar to defining theories of how objects break. One can break a given (unrepairable) object only once, so such theories must apply to sets of objects if one wants to be able to test multiple ways of falsifying the theories.

For conceptual clarity, I will focus on the case of a single binary decision to take, akin to a choice situation facing a bundle of two courses of action, exactly one of which must be adopted.

\section{Exposure protocol and deliberated choice}
I assume given two antagonistic acts, constituting the decision at stake, such as “buy this object for the price of \$15” and “do not buy this object for the price of \$15”, or “take this job offer” and “decline this job offer”.
%The letter $\phi$ will be used to denote a generic act (one of these two), and given an act $\phi$, $¬\phi$ denotes the other act.
Let the set $P = \set{\phi, ¬\phi, 0}$ denote the set of possible preferences, where $\phi$ denote a strict preference for the first act against the other one, $¬\phi$ conversely, and $0$ denotes indifference (both acts are considered equally worth).

\subsection{Exposure protocol}
An exposure protocol is a tuple $(I, \allargs, {\allleadsto})$, where $I$ is a set of individuals, $\allargs$ is a set of arguments, and ${\allleadsto}$, “leads to”, is the behavior of the individuals as a function of the arguments they have been exposed to.

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero and $\N^* = \N \setminus \set{0}$.
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N^*: \intvl{j, 0} = \emptyset$).
Define $\allhists = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence.
A generic element of $\allhists$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path. Its length is denoted by $\card{\hist} \in \N$ and its elements, $\hist(k)$ or $\hist_k$.
Thus, given an argumentative path $\hist \in \allhists$, $\forall k \in \intvl{1, \card{\hist}}: \hist(k) = \hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$. A path $\hist$ is a sub-path of another path $\histp$ whenever $\hist \subseteq \histp$, thus, $\forall k \in \intvl{1, \card{\hist}}: \hist_k = \histp_k$.

The function ${\allleadsto}: I × \allhists → P$ associates to each individual and path an element of $P$, ${\allleadsto}(i, \alpha)$ representing the choice of $i$ following exposure to the arguments of $\alpha$, in order. That is, ${\allleadsto}(i, \alpha) = \phi$ denotes that $i$ chooses the first act after exposure to $\alpha$, ${\allleadsto}(i, \alpha) = ¬\phi$ denotes $i$ choosing the second act and ${\allleadsto}(i, \alpha) = 0$ is used to represent indifference of $i$.
The notation $\alpha \ileadsto p$, pronounced “$\alpha$ leads $i$ to $p$”, will be used to mean ${\allleadsto}(i, \alpha) = p$, considering ${\ileadsto} \subseteq \allhists × P$ as a binary relation.
As usual with binary relations, $\hist \nileadsto p$ means ${\allleadsto}(i, \alpha) ≠ p$ or equivalently ${\allleadsto}(i, \alpha) \in P \setminus \set{p}$.

\subsection{Deliberated choice}
The deliberated choice of $i$ given the exposure protocol $(I, \allargs, {\allleadsto})$ is denoted $P_i \subset P$.
Its formal definition requires the notion of decisive argument.

Given $\hist \in \allhists$, $\range(\hist) = \hist(\intvl{1, \card{\hist}})$ denotes the range of $\hist$, that is, all arguments contained in the sequence $\hist$.
Let $\histend = \hist(\intvl{\max(1, \card{\hist} - 1), \card{\hist}})$ denote the set containing the last two arguments of the sequence $\hist$ if the sequence has at least two elements; the unique argument of the sequence if it has exactly one element; and the empty set if it is empty.

Given $i \in I, p \in P, \ar \in \allargs$, define $\ar \ileadstost p$, pronounced “$a$ leads definitely $i$ to $p$”, iff $\forall \hist \in \allhists \suchthat \ar \in \histend: \hist \ileadsto p$.
When $\ar \ileadstost p$, $\ar$ is said to be a decisive argument for $\prop$ from $i$’s point of view given the exposure protocol.

The deliberated choice of $i$ is then considered as the set of propositions supported by decisive arguments.
\begin{definition}[Deliberated choice]
  \label{def:decisive}
  $\forall i \in I, p \in P:
    p \in P_i ⇔ [\exists \ar \in \allargs \suchthat \ar \ileadstost p].$
\end{definition}

It follows from the consideration of “end” as the last \emph{two} arguments that decisive arguments can argue for at most one proposition, thus, that the deliberated choice cannot hold more than one proposition.
\begin{theorem}[Deliberated choice singularity]
  $\forall i \in I: \card{P_i} ≤ 1$.
\end{theorem}
\begin{proof}
  Assume that $\exists \ar, \ar' \in \allargs, p, p' \in P \suchthat \ar \ileadstost p \land \ar' \ileadstost p'$.
  Since $\ar \ileadstost p$, $(\ar, \ar') \ileadsto p$, and since $\ar' \ileadstost p'$, $(\ar, \ar') \ileadsto p'$, hence $p' = p$.
\end{proof}

The intended semantics is that $P_i$ represents the preference of $i$ over the proposed set of two acts, after deliberation, given the exposure protocol. In other words, if $P_i = \set{\phi}$, the individual considers the first act strictly better than the second one, after deliberation (and conversely for $P_i = \set{¬\phi}$); if $P_i = \set{0}$, the deliberated judgment of $i$ is that both acts are equally worth; and if $P_i = \emptyset$, no act is deliberately strictly better than the other one and they are also not deliberately exactly equivalent to each other.

This latter possibility may look unpleasant, but it is a fact that facing some difficult problems, individuals sometimes are undecided, and it is an a priori empirical possibility that the individual remains undecided (or even, \emph{becomes} undecided) when considering multiple arguments.
Being undecided is not the same state as being decidedly indifferent about two issues, and those states can be distinguished empirically, as we will see.

\subsection{An example and a discussion about observability}
The following toy example is inspired by an argument from \citet[p.~59]{railton_facts_2003}. Here is his argument.
\begin{quote}
  An experiment in cognitive psychology reveals that when subjects observe two individuals taking a test, one of whom answers a high percentage of the early questions correctly but then falters, while the other does
  poorly initially but then answers a high percentage of the later questions
  correctly, the first is usually viewed as more able, even though in the end
  each answers the same number of questions correctly in all. If the subjects had been asked which test-taker they would prefer as a math tutor,
  they likely would have answered “The first,” and perhaps they would also
  have been willing to pay a premium to secure his services. But if they
  were to become convinced of the influence of sheer order on their evaluations of relative ability, they presumably would no longer want their
  initial preference for the first test-taker to influence their choices – or
  willingness to pay – in this way.
\end{quote}
\begin{example}[Math tutor]
  \label{ex:tutor}
  Let $I$ be a set of individuals desiring to evaluate two candidates $\set{m_1, m_2}$ for math tutoring.
  Consider one individual $i \in I$.
  Let ${\allleadsto}(i, a_1)$ represent $i$ being shown a test of candidate $m_1$ wherein $m_1$ answers 70\% of the first half of questions correctly but then only 40\% of the other half correctly.
  Let ${\allleadsto}(i, a_2)$ represent $i$ being shown a test of candidate $m_2$ wherein $m_2$ answers 45\% of the first half of questions correctly, then 75\% of the other half correctly.
  %	Let ${\allleadsto}(i, a_e)$ represent $i$ being given an explanation about the psychological phenomenon described in the quote above.
  Let ${\allleadsto}(i, a_e)$ represent $i$ being given the following argument. “When tested, $m_1$ answers 70\% of the first half of questions correctly but then only 40\% of the other half correctly; whereas $m_2$ answers 45\% of the first half of questions correctly, then 75\% of the other half correctly. It follows than $m_1$ answers on average less well than $m_2$. Observing the candidates take the tests may yield a contrary impression; this is a well-known psychological phenomenon depending on the order of the questions that the candidates answer correctly.”
  Let $\prop_1$ represent the choice of $m_1$ and $\prop_2$ that of $m_2$.

  If the prediction described in the quote applies to $i$, the relation $\ileadsto$ is as follows, $\forall k \in \N^*$ (recall that for $\args \subseteq \allargs$, $\args^k$ represents the set of sequences of length $k$ using arguments from $\args$): $\forall \hist \in \set{a_e}^k, \hist \ileadsto \prop_2$; $\forall \hist \in \set{a_1, a_2}^k, \hist \ileadsto \prop_1$; $\forall \hist \in \set{a_1, a_e}^k \suchthat a_e \in \histend, \hist \ileadsto \prop_2$; $\forall \hist \in \set{a_2, a_e}^k \suchthat a_e \in \histend, \hist \ileadsto \prop_2$; $\forall \hist \in \allargs^k \suchthat a_e \in \histend, \hist \ileadsto \prop_2$. The values of ${\allleadsto}(i, \hist)$ for $\hist = \emptyset$ and for $\hist \in \allargs^k \suchthat a_e \notin \histend$ are undetermined by the prediction, but whatever they are, $a_e$ is a decisive argument for $\prop_2$, hence, $i$’s deliberated choice is $\prop_2$.
\end{example}

Now that deliberated choices are defined, we need to explore what one can learn about deliberated choices using our empirically available data.
This requires to define claims about deliberated choices and to explore how to validate claims.

\section{Claims about deliberated choice}
Given an exposure protocol $(I, \allargs, {\allleadsto})$, consider a claim about $\allleadsto$, such as “every individuals in $I$ choose deliberately $\phi$”, or, in \cref{ex:tutor}, “$a_e$ is a decisive argument for $i$ in favor of $\phi_2$”.
Validating claims about deliberated choice is no trivial matter, for two reasons.

First,
such claims must entail an infinity of constraints about $\allleadsto$:
in order to deduce that $p \in P_i$ for some $p \in P, i \in I$, the claim must constrain $\ileadsto$ on every possible paths $\alpha \in \allhists$.
It follows that a claim cannot be validated empirically exhaustively by observing the value of $\allleadsto$ on every points that it constraints.

Second,
our observations are limited in important ways: once we have shown some arguments to a given individual $i$, we cannot undo the effects of this action on $i$ and show to her a different argument as if starting from scratch (unless assuming $i$ forgets arguments after a while and waiting for sufficiently long, which would create practical and theoretical difficulties).
In other words, once we have shown a path $\hist$ to $i$, we can only continue the protocol with $i$ by showing paths $\histp$ that are suffixes of $\hist$.

To illustrate this point with \cref{ex:tutor}, let $\hist = (a_1, a_2, a_e)$ designate the argumentative path where the arguments are shown in the order they are described in \cref{ex:tutor}, and let $\hist' = (a_e, a_2, a_1)$ designate the reverse order.
The deliberated choice concept is defined so as to not depend on the order of the arguments shown to $i$: for having $\prop \in P_i$, it must be that directing to $i$ the path $\hist$ or the path $\hist'$ both lead $i$ to choose $\prop$.
But it is impossible to observe both the value of ${\allleadsto}(i, \hist)$ and the value of ${\allleadsto}(i, \hist')$.

The first difficulty is familiar as it occurs in any endeavor to validate general claims in science: claims that cover a large (or infinite) spectrum cannot be validated by observing the system exhaustively. One must rather develop a falsification protocol, that is, a way to invalidate the claim, and then observe that the claim is not invalidated. This is the approach taken in the next section.

This approach also permits to tackle the second difficulty, by letting us study theoretically the needs and limits of falsification given our observable basis.
% Such theories contain an observable claim and a conclusion claim, so as to satisfy two properties. First, such a theory must be falsifiable in principle, meaning that it must be possible, using our observational basis (propositions of the sort $\alpha \ileadsto \phi$), to invalidate the observable claim of the theory. As a counter-example, using \cref{ex:tutor}, a theory that claims that $\forall i \in I: \alpha_{123} \ileadsto \prop_1 \lor \alpha_{321} \ileadsto \prop_1$ is not falsifiable in principle, as, whatever the real $\ileadsto$ relation, with a given $i$, one can never show that $\alpha_{123} \nileadsto \prop_1 \land \alpha_{123} \nileadsto \prop_1$. Second, such a theory must be so that if its observable claim holds, then some given alternative is in the deliberated judgment of the individuals in $I$.

\subsection{Claims and observable basis}
% Such a deduction should use, as much as possible, the observational basis, that is, the propositions of the form $\alpha \ileadsto p$ that can be obtained experimentally.
Note first that given a relation ${\allleadstog} \in \allleadstos$, $\powerset{\allleadstog}$ denotes the set of relations over $(I × \allhists) × P$ that refine $\allleadstog$ (thus, each element $f \in \powerset{\allleadstog}$ is right-unique but non necessarily total, meaning that it does not necessarily cover $(I × \allhists)$).

Given a pair $(I, \allargs)$, a claim $C$ is a set of possibilities for $\allleadsto$, thus, $C \subseteq \allleadstos$.
Given a formula $\psi$, let the corresponding claim be $C_\psi = \set{{\allleadstog} \in \allleadstos \suchthat {\allleadstog} \text{ satisfies } \psi}$.
A claim is trivial iff it is the trivial claim, defined as $C = \allleadstos$.
A claim is finite iff its cardinal is finite (meaning that it is compatible with a finite number of possibilities).
Let $C^\wsub = \cup_{{\allleadstog} \in C} \powerset{\allleadstog} = \set{c \subseteq (I × \allhists) × P \suchthat \exists {\allleadstog} \in C \suchthat c \in \powerset{\allleadstog}}$ denote the set of non necessarily complete possibilities compatible with $C$ (thus, each element $c \in C^\subseteq$ is right-unique but non necessarily total, meaning that it does not necessarily cover $(I × \allhists)$).
Note that $C^\subseteq$ is downwards closed, meaning that $\forall c \in C^\subseteq: c' \subseteq c \implies c' \in C^\subseteq$.

Given a pair $(I, \allargs)$, an observation $θ \in \Theta$ is a right unique finite subset of $(I × \allhists) × P$ such that $\forall i \in I, (i, \hist, p), (i, \histp, q) \in θ: \card{\hist} ≤ \card{\hist'} \implies \hist \subseteq \hist'$.
The set $\Theta$ contains the a priori possible observations, that is, those that can be considered possible when nothing is known about $\allleadsto$.
We say that an observation $\theta$ is compatible with a claim $C$ iff
$\theta \in C^\subseteq$, equivalently, iff
$\theta \in \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
Thus, given ${\allleadstog} \in \allleadstos$, the set $\Theta \cap \powerset{\allleadstog}$ is the set of observations compatible with ${\allleadstog}$.

Given an observation $\theta$, let $\theta^{-1}(P) \subseteq I × \allhists$ denote the pairs $(i, \hist)$ in $\theta$.
Let $\incr(\allhists) = \set{\emptyset} \cup \cup_{\emptyset \neq N \subseteq \N} \set{A(N) \suchthat A \in (\allhists)^N \land [\card{A_i} = i] \land [i < j \implies A_i \subsetneq A_j]} \subseteq \powerset{\allhists}$ denote the possibly empty embedded sequences of paths.
\commentOC{TODO define $\incr(\allhists) = \set{H \wsub \allhists \suchthat \card(H) ≤ \card(H') \implies H \wsub H'}$.}
An element $\theta^{-1}(P)$ can also be seen as a (finite) function associating to each $i \in I$ an element of $\incr(\allhists)$.
Therefore, we let $\finite(\incr(\allhists)^I) = \set{\theta^{-1}(P) \suchthat \theta \in \Theta}$ denote the set of possible observations restricted to $I × \allhists$, the elements of which being called \emph{a priori observation}, where $\finite(\incr(\allhists)^I)$ denote the functions $f: I \mapsto \incr(\allhists)$ with finite support, thus $\card{f^{-1}(\incr(\allhists) \setminus \emptyset)} < \N$.
Note that $\allpriobss$ and $\incr(\allhists)^I$ are downwards closed, meaning that $\forall O \in \allpriobss: O' \subseteq O \implies O' \in \allpriobss$ and $\forall O \in \incr(\allhists)^I: O' \subseteq O \implies O' \in \incr(\allhists)^I$.
Let also $\max\incr(\allhists) = \set{A \in \incr(\allhists) \suchthat \nexists A' \in \incr(\allhists) \suchthat A \subset A'} = \set{A \in \incr(\allhists) \cap (\allhists)^\N}$ denote the maximal elements of $\incr(\allhists)$, in the sense of containing sequences that are not a subsequence of another sequence; equivalently, containing those sequences whose sequence of cardinals of their elements equals $\N$.

Let the \emph{simple a priori observations} be defined as $\cup_{J \subseteq_\finite I} (\allhists)^J$, equivalently, as the right-unique finite relations over $I × \allhists$.
Given a simple a priori observation $O: J \mapsto \allhists$, we can define the a priori observation $O_O = \set{(i, \hist) \in J × \allhists \suchthat \forall i \in J: \hist \subseteq O(i)}$ as the completion of $O$ to all sub-paths in $O$.

Similarly, we can view an observation $\theta$ as a function $\theta: I \mapsto \cup_{A \in \incr(\allhists)} P^A$ and thus $\Theta = \finite((\cup_{A \in \incr(\allhists)} P^A)^I)$, where $\finite$ designates the finite relations (considered as a set of triples).
Note that $(\cup_{A \in \max \incr(\allhists)}P^A)^I$ is the set of functions that associate to each $I$ a maximal increasing sequence of paths, together with values from $P$.

Given a claim $C$, we can also define the set $\Theta \cap C^\subseteq$ of possibilities for increasing sequences corresponding to the claim.
This set is downwards-closed, as both $\Theta$ and $C^\subseteq$ are.

\subsection{Rejection and falsification}
Given an exposure protocol $(I, \allargs, {\allleadsto})$, and a claim about ${\allleadsto}$, we develop here a procedure to deduce that the claim is correct.

Given an exposure protocol $\allleadsto$, we say that a claim $C$ \emph{holds} iff $C \ni {\allleadsto}$.

At our disposal is the procedure of empirical rejection, that consists in directly invalidating a claim by observing something that contradicts it.
\begin{definition}
  Given an exposure protocol $\allleadsto$, we say that $C$ is \emph{rejectable} iff
  $\Theta \cap \powerset{\allleadsto} \nsubseteq C^\wsub$.
\end{definition}

The following theorem verifies that when $C$ holds, it is not rejectable.
\begin{theorem}
  \label{th:ex-post}
  $\forall C \ni {\allleadsto}:
    \Theta \cap \powerset{\allleadsto} \subseteq C^\wsub$.
\end{theorem}
\begin{proof}
  This holds by definition of $C^\wsub = \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
\end{proof}

The converse, however, does not hold.
\begin{example}[Disjunct with incompatible paths, not rejectable]
  \label{ex:disjR}
  Given $\ar \neq \ar' \in \allargs$,
  the claim $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not rejectable.
  In fact,
  $\Theta \wsub C^\wsub$.
  Consider any $θ \in \Theta$ and let us prove that $\theta \in \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
  By definition of $\Theta$, $\theta^{-1}(P)$ may not contain both $((\ar), i)$ and $((\ar'), i)$, thus letting $t$ denote one of these pairs not in $\theta^{-1}(P)$ and defining $r = (t, \phi)$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have $\theta \in \powerset{\allleadstog'}$ and ${\allleadstog'} \in C$.
\end{example}
\Cref{ex:disjR} shows a claim that is not rejectable, irrespective of the exposure protocol.
In particular, the example applies when $C$ does not hold (that is, when ${\allleadsto} \notin C$). We conclude that a claim may fail to hold but still be not rejectable.
It follows that non-rejection cannot be a sufficient condition for validating a claim.

\begin{remark}
  \Cref{ex:bigor,ex:threeFs} present situations where, similarly, the claim may fail to hold but still be not rejectable.
  In \cref{ex:bigor}, non rejection is due to finiteness of observations.
  In \cref{ex:threeFs}, rejection depends on ${\allleadsto}$.
\end{remark}

\Cref{def:falsw} fills the gap to obtain that $C$ holds iff it is not rejectable.
The property demands that whatever the attitude of individuals, if the attitude is not compatible with the claim, then it is possible to observe that it is not.
\begin{definition}
  \label{def:falsw}
  A claim $C$ is \emph{falsifiable or trivial} iff
  $\forall {\allleadstog} \notin C: \Theta \cap \powerset{\allleadstog} \nsubseteq C^\wsub$.
\end{definition}

It is important to note that falsifiability does not depend on $\allleadsto$: whether falsifiability holds can be determined analytically, meaning, by considering only the content of the claim, without considering the behavior of individuals.

\Cref{th:holdsiff} shows that falsifiability indeed fills that gap:
falsifiability guarantees that a claim holds if, and only if, it is not rejectable.
\begin{theorem}
  \label{th:holdsiff}
  For any falsifiable or trivial claim $C$: $C \ni {\allleadsto}$ iff $\Theta \cap \powerset{\allleadsto} \wsub C^\wsub$.
\end{theorem}
\begin{proof}
  Consider any falsifiable $C$.
  Using the contrapositive of \cref{def:fals}, $\forall {\allleadstog} \in \allleadstos: \nexists θ \in \Theta \cap \powerset{\allleadstog} \suchthat \theta \notin C^\wsub \implies {\allleadstog} \in C$.
  In particular,
  $\nexists θ \in \Theta \cap \powerset{\allleadsto} \suchthat \theta \notin C^\wsub \implies {\allleadsto} \in C$.
  The reverse implication is proven by \cref{th:ex-post}.
\end{proof}

\begin{definition}
  A claim is \emph{strongly unfalsifiable} iff
  $\Theta \wsub C^\wsub$.
\end{definition}
\Cref{ex:disjR,ex:bigor} illustrate, as well as \cref{ex:existsParticular} when $I$ is infinite.
\begin{definition}
  A claim is \emph{strictly unfalsifiable} iff
  it is unfalsifiable and not strongly unfalsifiable.
\end{definition}

Making a strongly unfalsifiable claim more precise ($C = (C_1 \cap C_2) \sub C_1$ in \cref{th:strongUToStrictU}) can make it strictly unfalsifiable.
Also, a strictly unfalsifiable claim can be made falsifiable by picking a single relation from the claim.

But, it does not hold that more precise claims are always more falsifiable:
making a falsifiable claim more precise ($C = (C_1 \cap C_2) \sub C_2$ in \cref{ex:partlyU}) can make it strictly unfalsifiable.
(In fact, \cref{ex:partlyU} shows that the relation is not one-way: going from a strongly unfalsifiable ($C_1$), or from a falsifiable ($C_2$), claim, to a more precise one ($C = C_1 \cap C_2$) yields a strictly unfalsifiable claim.)

\begin{example}
  \label{ex:allParticular}
  Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
  Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $θ = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}

The following procedure permits to determine whether a claim is falsifiable.
Given a set of observations $S \wsub \Theta$, define the relations that are observation-compatible with $S$ as $E(S) = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: {\allleadstog} \cap \theta \in S}$.
Given a claim $C$, define the set of its observation-compatible relations as $E(C) = E(C^\wsub \cap \Theta)$ (reusing the letter $E$ by abuse of notation).
Note that this set extends $C$, in the sense that $C \wsub E(C)$.
The claim $C$ is falsifiable iff it is closed under extension to observation-compatible relations, that is, iff $E(C) \subseteq C$.
\begin{theorem}
  \label{thm:caractFals}
  A claim $C$ is falsifiable or trivial iff $E(C) \wsub C$, or equivalently, iff 
  $\forall {\allleadstog} \in \allleadstos$:
  \[\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq \implies {\allleadstog} \in C.\]
\end{theorem}
\begin{proof}
  Note that
  $E(C) = E(C^\wsub \cap \Theta) = \set{{\allleadstog} \in \allleadstos \suchthat \forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\wsub}$.

  The condition $E(C) \wsub C$ is equivalent to
  $\forall {\allleadstog} \in \allleadstos: ({\allleadstog} \in E(C) \implies {\allleadstog} \in C)$.
  Note that ${\allleadstog} \in E(C) \Leftrightarrow \set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq$.

  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \Theta \cap \powerset{\allleadstog} \wsub C^\wsub$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.
  And, $\forall {\allleadstog} \in \allleadstos$: $\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} = \set{θ \subseteq {\allleadstog}}$.
\end{proof}

Let $S_θ = \set{{\allleadstog} \in \allleadstos \suchthat θ \subseteq {\allleadstog}}$ denote the possibilities for $\allleadsto$ compatible with the observations.
Given $O$ and ${\allleadstog}$, it follows that $S_{\restr{\allleadstog}{O_O}}$ denotes the “leads to” relations that complete $O_O$ using $\allleadstog$.
Define that a relation ${\allleadstog} \in {\allleadstos}$ is \emph{obs-compatible} with a set $S$ iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S$.

\begin{remark}
  \label{rq:closedcomb}
  Define the closure under combinations of $C$ as the relations that can be formed by picking parts of relations in $C$, formally, $(\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos$.
  We have $E(C) \wsub (\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos$.

  Being closed under combinations is sufficient to be falsifiable, but not necessary, as seen with \cref{ex:consti2} (in other words, for some claims, $(\cup_{S \in C^\wsub \cap \Theta} S) \cap \allleadstos \wsub E(C)$ does not hold).
\end{remark}

Define the \emph{suffix} relation $R = \set{(\hist, \histp) \in \allhists × \allhists \suchthat \histp = (\hist, .)}$ as the (infinite) tree whose set of nodes is $\allhists$ and which links  argumentative paths to all of their direct suffixes.
Note that a path in such a tree (an increasing sequence of elements in $\allhists$) corresponds to an argumentative path, that is, to an element of $\allhists$.

Any function ${\ileadstog}: \allhists \mapsto P$ can be considered as having this tree as a domain.
% Define a forest as an association of each $i \in I$ to such a tree.
Any relation ${\allleadstog} \in \allleadstos$ can then be viewed as a forest associating each $i \in I$ to a tree labeled by the function ${\ileadstog}$ (see \cref{fig:forest}).

\begin{figure}
  \begin{tikzpicture}[level distance=0.7cm]

    \tikzstyle{treenode}=[execute at begin node=$, execute at end node=$]
    \tikzstyle{treelabel}=[treenode, anchor=west, inner sep=0]
    \tikzstyle{bottomnode}=[execute at begin node=$\scriptstyle, execute at end node=$]
    \tikzstyle{bottomlabel}=[bottomnode, anchor=west, inner sep=0]
    \tikzstyle{tree}=[level 1/.style={sibling distance=3.4cm, every node/.style=treenode},
    level 2/.style={sibling distance=1.6cm, every node/.style=bottomnode},
    level 3/.style={sibling distance=0.7cm, every node/.style=treenode}
    ]
    \newlength{\lhi}
    \setlength{\lhi}{3.5cm}
    \newlength{\lvi}
    \setlength{\lvi}{1.2cm}

    \path node (all) {$\allleadstog$};

    \path (-\lhi, -\lvi) node[anchor=north] (tree1) {\begin{tikzpicture}
        \path[tree] node (root) {$\emptyset$}
        child {
            node (a) {(a)}
            child {
                node (aa) {(aa)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
            child {
                node (ab) {(ab)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
          }
        child {
            node (b) {(b)}
            child {
                node (ba) {(ba)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
            child {
                node (bb) {(bb)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
          }
        ;
        \begin{scope}[style=treelabel]
          \path (root.east) node {\mapsto p};
          \path (a.east) node {\mapsto p'};
          \path (b.east) node {\mapsto p};
        \end{scope}
        \begin{scope}[style=bottomlabel]
          \path (aa.east) node {\mapsto p'};
          \path (ab.east) node {\mapsto p'};
          \path (ba.east) node {\mapsto p'};
          \path (bb.east) node {\mapsto p};
        \end{scope}
      \end{tikzpicture}};
    \path[->] (all) edge node[midway, above] {$i_1$} (tree1.north);

    \path (\lhi, -\lvi) node[anchor=north] (tree2) {\begin{tikzpicture}
        \path[tree] node (root) {$\emptyset$}
        child {
            node (a) {(a)}
            child {
                node (aa) {(aa)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
            child {
                node (ab) {(ab)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
          }
        child {
            node (b) {(b)}
            child {
                node (ba) {(ba)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
            child {
                node (bb) {(bb)}
                child {
                    node {…}
                  }
                child {
                    node {…}
                  }
              }
          }
        ;
        \begin{scope}[style=treelabel]
          \path (root.east) node {\mapsto p};
          \path (a.east) node {\mapsto p};
          \path (b.east) node {\mapsto p'};
        \end{scope}
        \begin{scope}[style=bottomlabel]
          \path (aa.east) node {\mapsto p};
          \path (ab.east) node {\mapsto p'};
          \path (ba.east) node {\mapsto p'};
          \path (bb.east) node {\mapsto p};
        \end{scope}
      \end{tikzpicture}};
    \path[->] (all) edge node[midway, above] {$i_2$} (tree2.north);
  \end{tikzpicture}
  \caption{A forest}
  \label{fig:forest}
\end{figure}

Given an attitude $\allleadsto$, deliberated choices are represented by an element $\pi({\allleadsto}) \subseteq I × P$ (thus with $\pi({\allleadsto})(i) = P_i$) which is right unique (meaning that $\forall i \in I: \card{\pi({\allleadsto})(i)} ≤ 1$).
Equivalently, let $(P \cup \set{\emptyset})^I$ denote the right unique relations over $I × P$; and then the deliberated choices are an element $\pi({\allleadsto}) \in (P \cup \set{\emptyset})^I$.

A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
% A claim $C$ is \emph{indicative} iff $\exists H \in \finite(\incr(\allhists)), i \in I \suchthat \pi(\set{{\allleadstog} \suchthat H \wsub \ileadstoinvg(P)} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub \suchthat \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
\begin{remark}
  Informativeness permits mere counting: the claim that exactly two persons deliberately believe $\phi$ is informative.
  Informativeness may be a desirable property (rather than informativeness beyond counting) as one may want to know whether a given proposition is believed by a majority of individuals, or whether a given proposition is believed by nobody.
  Note also that strong informativeness beyond counting implies informativeness beyond counting and indicativeness.
  Note also that informativeness beyond counting implies informativeness.
\end{remark}
\begin{remark}
  We could weaken indicativeness by allowing for disjuncts: observing either that $\hist \ileadsto \phi$ or that $\hist \ileadsto \psi$ permits to deduce something about $P_i$.
  (Or more complex formulas under the form of a set of possible functions in $P^H$?)
  (Actually I am not sure that this really weakens the condition.)
\end{remark}

\begin{definition}[Anonymity]
  Given a permutation $\sigma$ over $I$ and any ${\allleadstog} \in \allleadstos$, let ${\allleadstog} \circ \sigma$ denote $\set{(i, \hist, \phi) \suchthat (\sigma(i), \hist, \phi) \in {\allleadstog}}$ and let $C \circ \sigma$ denote $\set{{\allleadstog} \circ \sigma \suchthat {\allleadstog} \in C}$.
  A claim $C$ is \emph{anonymous} iff $\forall \sigma: C = C \circ \sigma$.
\end{definition}
\Cref{ex:anon} illustrates.

We are now interested in the interplay between properties of holding, anonymity, (strong) informativeness beyond counting, and falsifiability.
\Cref{th:hsif,th:hif,th:asif} show their compatibility when one of holding or anonymity is not required (except that informativeness beyond counting cannot be made strong in the edge case where $\forall i \in I: P_i = \emptyset$).

We thus turn to requiring both anonymous and holding.
They are always satisfiable when only these two are required (\cref{th:ha}).
\Cref{th:infoBeyonCKO} (and extensions, \cref{th:infoBeyonCKOext,th:strinfoBeyonCKO}) shows that informativeness beyond counting (and variants) is too demanding if one wants to investigate situations where individuals have differing deliberate preferences.

It is possible under reasonable conditions of regularity to satisfy anonymous and holding together with indicativeness. This happens for example when everybody has a deliberate judgment and a finite set of decisive arguments suffices to lead everybody to their deliberate judgment (\cref{th:suffIndicDec}).

\begin{theorem}
  \label{th:suffFals}
  Assume that $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Then, there exists a claim that holds and is anonymous and falsifiable.
\end{theorem}
\begin{proof}
  Define $C = \set{{\allleadstog} \suchthat \forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}}$. It holds and is anonymous.
  Pick any ${\allleadstog} \notin C$.
  Then, $\exists i \in I \suchthat {{\ileadstog}} \notin \set{{\ileadsto[k]} \suchthat k \in I}$.
  Pick any such $i$ and any $H$ such that $\forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Define $\theta = \set{(\hist, i, {\ileadstog}(\hist)) \suchthat \hist \in H} \in \Theta \cap \powerset{\allleadstog}$.
  Then, $\theta \notin C^\wsub$, because $\forall j \in I: \theta \nsubseteq {\ileadsto[j]}$.

  For another proof, note that $C^\wsub \cap \Theta = \set{\theta \in \Theta \suchthat \forall i \in I: \exists j \in I \suchthat \theta \cap (\allhists × \set{i} × P) \wsub {\ileadsto[j]}}$.
  Pick any ${\allleadstog} \in E(C^\wsub \cap \Theta)$ and let us show that ${\allleadstog} \in C$.
  Note that ${\allleadstog} \notin C \iff \exists i \in I \suchthat {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, in which case the consequent in the theorem holds.
  We need to show that $\nexists i \in I, H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  In other words, given any $i \in I$, we need to show that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.
\end{proof}

\begin{remark}
  The opposite of \cref{th:suffFals} is NOT $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Thus, for some ${\allleadstog} \in \allleadstos$ and some $i \in I$ with ${\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, we have that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.

  For example, $\allargs = \set{\ar_1, \ar_2}$ and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and the attitude contains only people constant to $\phi$ (except that they map $\emptyset$ to anything) and people constant to $¬\phi$ (except that they map $\emptyset$ to anything).
\end{remark}
\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}

\subsection{Ideas for no suitable}
$C$ is \emph{suitable} iff it holds and is anonymous, informative and falsifiable.
\begin{example}
  Assume that $\exists \ar, \phi \suchthat \forall i: \ar \ileadstost \phi$.
  Then $\forall i: \ar \ileadstostg \phi$ is suitable.
  Note that it is falsifiable but not verifiable (that is, no $O$ implies $C$).
\end{example}

Our goal now is to find an exposure protocol $\allleadsto$ with no suitable claim.

Could we have that if it holds then it is not informative?
$i_1$ thinks $p$ and $i_2$ thinks $¬p$.

Given $\histk \in \allhists$ and $\ar_l \in \allargs$, let $(\histk, …, \ar_l, ?)$ denote the set
$\set{\histl \suchthat \histk \subset \histl \land \ar_l \in \histlend}$.
Given $\hist \in \allhists$, ${\ileadstog}$ and $\ar_k, \ar_l \in \allargs$, let $(\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set
$\set{\histl \suchthat \exists \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given $\ar_k, \ar_l \in \allargs$, let
$(…, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set
$\cup_{\hist \in \allhists} (\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?) =
  \set{\histl \suchthat \exists \hist, \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend} =
  \set{\histl \suchthat \exists \histk \suchthat \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given ${\ileadstog}$, $p \in P$, $s \in \allargs$ and $S \subseteq \allargs$, note that $s \in \ileadstoinvg(p)$ stand for $s \ileadstog p$ and $S \subseteq \ileadstoinvg(p)$ stand for $\forall s \in S: s \ileadstog p$.

\begin{theorem}
  Given any $p \in P, \ar_k, \ar_l \in \allargs$, consider the following claim $C$:
  \begin{multline}
    \label{eq:cl}
    \forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
    \forall \hist \in (…, \ar_1, ?) \cap \nileadstoinvg(p): \\
    (\hist, …, \ar_2, ?) \subseteq \ileadstoinvg(p).
  \end{multline}
  Assume that ${\allleadsto} \in C$. Then
  $\exists i \in I \suchthat [\ar_k \nileadstost p \land \ar_l \nileadstost p]$
  is equivalent to:
  \begin{multline}
    \label{eq:cln}
    \exists i \in I \suchthat \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l}:\\
    \exists \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p) \suchthat
    (\hist, …, \ar_2, ?) \subseteq \ileadstoinv(p).
  \end{multline}
\end{theorem}
\begin{proof}
  $\ar_k \nileadstost p$ iff $\exists \hist \suchthat \ar_k \in \histend \land \hist \nileadsto p$.
\end{proof}
\begin{remark}
  Idea.
  That consequence is unfalsifiable so we cannot include anything like this in $C$ (only something stronger, but of course a stronger statement could be false). Therefore, with no addition, we are doomed that exists some situations where no claim is suitable.
\end{remark}
\begin{theorem}
  Pick any $p \in P$ and $\ar_k ≠ \ar_l \in \allargs$ and consider the claim \eqref{eq:cl} and not \eqref{eq:cln}, thus, \eqref{eq:cl} and $\forall i, \ar_k \ileadstostg p \lor \ar_l \ileadstostg p$. It is not falsifiable.
\end{theorem}
\begin{proof}
  %	Intuitively, because it might require (with $\hist = (\ar_1)$) to observe that for a given $i$, $(a_1) \nileadsto p \land (\ar_1, \ar_2) \ileadsto p$ and also $(\ar_2) \nileadsto p \land (\ar_2, \ar_1) \ileadsto p$.
  We need to exhibit some ${\allleadstog} \notin C$ such that $\forall θ \subseteq {\allleadstog}: \theta \in C^\wsub$.
  Define $\allleadstog$ as, $\forall i \in I$, $(\ar_k) \nileadstog p, (\ar_l) \nileadstog p, \forall \hist \in \allhists \setminus \set{(\ar_k), (\ar_l)}: \hist \ileadstog p$.
  It follows that $\ar_k \nileadstostg p \land \ar_l \nileadstostg p$. Thus, ${\allleadstog} \notin C$.
  Now pick any $θ \subseteq {\allleadstog}$.
  % The set $O$ can’t have both $(\ar_k) \nileadstog p$ and $(\ar_l) \nileadstog p$.
  % Assume, wlog, that $(\ar_k, i, {\allleadstog}(\ar_k, i)) \notin O$.
  %	Then, intuitively, $\ar_k$ decisive is in $S_θ \cap C$.
  Define $\allleadstog'$ as, $\forall i \in I: \forall \hist \in \allhists \setminus \set{(\ar_k), (\ar_l)}, \hist \ileadstog' p$
  and $\forall \ar \in \set{\ar_k, \ar_l}$, $(\ar) \ileadstog' ({\allleadstog}(\ar, i))$ if $(\ar, i, {\allleadstog}(\ar, i)) \in O$, $(\ar) \ileadstog' p$ otherwise.
  %Pick any $p' ≠ p$ and define $\allleadstog'$ as, $\forall i \in I: \forall \hist \in \allhists \setminus \set{(\ar_l)}, \hist \ileadstog' p \land (\ar_l) \ileadstog' p' ≠ p$.
  The set $O$ can’t have both $(\ar_k) \nileadstog p$ and $(\ar_l) \nileadstog p$, thus, ${\allleadstog'} \in C$.
  We conclude that ${\allleadstog'} \in S_θ \cap C$.
\end{proof}
\begin{theorem}[Informal]
  Assume that the NOT consequence holds and the claim holds.
  Prove that there is no suitable claim.
  If the claim is just the NOT cons and claim, then it is not fals.
  If the claim is weaker, it does not permit to prove that $\ar_k$ or $\ar_l$ is decisive. (BUT does it possibly prove that some arg is decisive? But it may be false.)
  If the claim is stronger, it may be false.
\end{theorem}
\begin{theorem}[Formal]
  Pick any $p \in P$ and $\ar_k ≠ \ar_l \in \allargs$ and assume that $\allleadsto$ fails \cref{eq:cln}, that is, satisfies:
  \begin{multline}
    \label{eq:answerable}
    \forall i \in I, \exists \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \suchthat
    \forall \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p): \\
    (\hist, …, \ar_2, ?) \cap \ileadstoinv(p) ≠ \emptyset.
  \end{multline}
  Assume further that $\allleadsto$ satisfies \eqref{eq:cl}, thus:
  $\forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
    \forall \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p): (\hist, …, \ar_2, ?) \subseteq \ileadstoinv(p)$.
  Then, there is no suitable claim.
\end{theorem}
\begin{proof}
  Note that $\allleadsto$ satisfies, $\forall i \in I, \exists \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \land \forall \hist \in (…, \ar_1, ?): \hist \ileadsto p$; equivalently, $\ar_k \ileadstost p \lor \ar_l \ileadstost p$.

  Assume the claim is just \eqref{eq:cl} $\land$ \eqref{eq:answerable}. Then if is not fals.

  If the claim is sufficiently stronger (to be falsifiable) than this (thus with $C' \subset C$, thus with ${\allleadstog} \notin C'$ as defined above), then it may be false.
  Indeed, to be falsifiable, we need ${\allleadstog'} \notin C'$.
  Now we need $C'$ to hold so it must contain ${\allleadsto}$ but not contain ${\allleadstog'}$. It does not contain ${\allleadstog'}$ iff it implies $¬(\land_{\hist \in \allhists \setminus \set{(\ar_l)}} \hist \ileadstog' p \land (\ar_l) \ileadstog' p')$.
  %So it must claim that $\ar_l$ is not decisive. Or something like, for some $\hist_1$: if $\ar_l \ileadstog' p'$, then $\hist_1 \ileadstog' ¬p$.

  Intuitively, need that when we do not observe $(\ar_k)$ for a given $i$, but only $(\ar_l)$ for that $i$, the claim says something contrary about that $(\ar_l, i)$.
\end{proof}
\begin{remark}
  Idea: then use contrapositive. Not cons implies some dec arg.
  Prove equivalence to establish that we need to assume not cons to make such claims informative (in such a way).
  We could also attempt to prove that $\exists \ar \suchthat \ar \ileadstost \phi$ but assuming the claim $C$ is everything we know, there is no hope in proving something about some argument that we ignore.

  As an example, with the shortest possible paths, we get $\histk = (\ar_k)$ and $\histl = (\ar_l)$, thus:
  \begin{equation}
    \left\{\begin{aligned}
       & (\ar_k ) \nileadsto \phi \land \set{\histp \suchthat (\ar_k) \subseteq \histp \land \ar_l \in \histpend} \ileadsto \phi; \\
       & (\ar_l) \nileadsto \phi \land \set{\histp \suchthat (\ar_l) \subseteq \histp \land \ar_k \in \histpend} \ileadsto \phi.
    \end{aligned}\right.
  \end{equation}

  Idea: first consider that $a_l$ equals $a_j \cup a_k$ so that we can reasonably use an axiom.
  Then in the more general case, assume $\phi$ is not in DJ at all, which also implies that $a = a_k \cup a_l$ fails; and we fall back to the same.
\end{remark}
\begin{remark}
  The theorem is realistic in the following sense.
  Some need $a_k$ only; others need $a_j \cup a_k$ (ignorant of $a_j$ at start).
  Consider $C$ saying that if $a_k$ not convincing, then $a_j \cup a_k$ is convincing afterwards and if $a_j \cup a_k$ not convincing, then $a_k$ convincing afterwards.
  Define $a_l = a_j \cup a_k$.

  In other words.
  Assume that knowing both $a_j$ and $a_k$ is required to get to $\phi$.
  Assume that $J$ knows $a_j$ and when told it, gets bored, thus not convincing. For $j$, only $a_k$ is decisive.
  But $K$ ignores $a_j$. For $k$, only $a_j \cup a_k$ is decisive.
  %	
  %	Could have:
  %	a_k then a_l OK
  %	ak KO
  %	al KO
  %	al then ak OK
\end{remark}

\begin{example}[To be completed]
  Assume that $\exists \ar_j ≠ \ar_k, J, K \subset I \suchthat J \cup K = I \land \card{J} \notin \N \land \card{K} \notin \N \land \forall j \in J: \ar_j \ileadstost[j] \phi \land \forall k \in K: \ar_k \ileadstost[k] \phi \land \forall \ar \in \allargs \setminus \set{\ar_j, \ar_k}, i \in I: \ar \nileadstost \phi$.
  Then no suitable $C$ exists?

  Let’s find some suitable $C$ that shows that $\phi \in P_i$.
\end{example}

Intuition: if $a_j$ and afterwards $a_k$ is convincing, then either $a_j$ is unuseful in which case $a_k$ is decisive; or $a_j$ is useful in which case $a_j \cup a_k$ is decisive. However, could also be that $a_j$ is essential as first argument; thus $a_j$ is useful but $a_j \cup a_k$ is not decisive as $a'$ then $a_j \cup a_k$ is not convincing. Note that the latter part can be excluded by a falsifiable claim: $\nexists a' \suchthat a', a_j \cup a_k \nileadsto \phi$.

The following antecedent excludes two of these three hypothesis, leaving only the one that $a_j$ is useful.
$((a_j), …, a_k)$ convincing ; $a_k$ not decisive (or $(a_k)$ not convincing?); $\nexists \ar \suchthat [(\ar, …, a_k)$ not convincing $\land(\ar, …,  a_j \cup a_k)$ not convincing].
  The conclusion is: $a_j \cup a_k$ decisive.

  Is this an axiom or is this provable?
  And: is the part with $\nexists \ar$ a falsifiable claim?

  \subsection{Alternative definition of syntactic falsifiability}
  A claim $C$ is attempt-falsifiable iff $\exists θ \suchthat θ \implies ¬C$, by which I mean $\theta \notin C^\wsub$.
  \begin{example}
    $(\ar) \ileadstog \phi \lor (\ar, \ar) \ileadstog \phi \lor …$ is not falsifiable.
    %	Its negation is $\forall k: (\ar^k) \nileadsto \phi$.
    %	The claim represents the set $\set{{\allleadsto} \in \allleadstos \suchthat (\ar) \ileadsto \phi \lor  (\ar, \ar) \ileadsto \phi \lor …}$.
    Given any $O$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in O} \card{\hist} + 1$. Pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_θ \cap C ≠ \emptyset$.
  \end{example}
  \begin{example}
    $(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
    Given any $O$, it may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land t \in {\allleadstog}$. We have ${\allleadstog} \in S_θ \cap C ≠ \emptyset$.
  \end{example}
  \begin{example}[Fals. but shouldn’t]
    \label{ex:fbs}
    $\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is falsifiable.
    Consider $θ = \set{(j, \hist, ¬\phi)}$.
  \end{example}
  \begin{example}
    $\forall i: \ar \ileadstostg \phi$ is falsifiable.
    Consider any $i \in I$ and $θ = \set{(i, (\ar), ¬\phi)}$.
  \end{example}
  \Cref{ex:fbs} shows that this definition of falsifiability is wrong.

  \subsection{Argumentative theories of deliberated choice}
  \label{sec:dynamic}
  An argumentative theory of deliberated choice $\gamma$ is defined as a tuple $(\pga, \dg)$, where $\pga \in P$ is the proposition supported by the theory and $\dg$ is a defense relation.

  The defense relation $\dg \subseteq \allhists × \allargs$ defines the argumentation strategy of the theory. Informally speaking, it defines the arguments that the theory claims will convince the individuals in $I$ to choose $\pga$. Assuming that some path $\hist$ challenges the theory’s intended conclusion in favor of $\pga$ (as judged by the individual), $\dg(\hist)$ is the argument that the theory plays as a counter-argument to the path $\hist$.

  Formally speaking, a defense relation $\dg \subseteq \allhists × \allargs$ is a right-unique relation satisfying bounded depth (to be defined shortly). Right-unique means that given any $\hist \in \allhists$, $\dg(\hist) = \set{\ar \in \allargs \suchthat (\hist, \ar) \in \dg}$ is empty or is a singleton.
  As is usual for right-unique relations, we write $\dg(\hist) = \ar$ instead of $\dg(\hist) = \set{\ar}$ when $\dg(\hist)$ is a singleton (and $\dg(\hist) = \emptyset$ otherwise).
  Given $H \subseteq \allhists$, define $\dg(H) = \set{\dg(\hist) \suchthat \alpha \in H} \subseteq \allargs$.
  Given $\ar_0, \ar_1 \in \allargs$, say that $\ar_1$ defends $\ar_0$
  iff $\exists \hist \in \allhists \suchthat \ar_0 \in R(\hist) \land \dg(\hist) = \ar_1$, equivalently,
  iff $\ar_0 \in R(\dg^{-1}(\ar_1))$, equivalently,
  iff $\ar_1 \in \dg(\set{\hist \suchthat \ar_0 \in R(\hist)})$.

  Say that $\gamma$ has depth at least $k \in \N^*$ iff $\exists (\alpha_i, 1 ≤ i ≤ k) \suchthat \forall 1 ≤ i ≤ k, \dg(\set{\alpha_j \suchthat 1 ≤ j ≤ i - 1}) \subsetneq \dg(\set{\alpha_j \suchthat 1 ≤ j ≤ i}) \land \forall 1 ≤ i ≤ k - 1, \alpha_i \cup \dg(\alpha_i) \subseteq \alpha_{i + 1}$.
  The bounded depth condition on the defense relation imposes that for some $k \in \N^*$, $\gamma$ does not have depth at least $k$.
  The depth of a theory is defined as the smallest value $k \in \N$ such that it does not have depth at least $k + 1$.
  For example, $\gamma$ has depth $0$ iff $\dg = \emptyset$; and $\gamma$ has depth $1$ iff $[\exists \alpha_1 \suchthat \dg(\alpha_1) ≠ \emptyset] \land [\forall \alpha_1, \alpha_2 \suchthat \dg(\alpha_1) ≠ \emptyset \land \alpha_1 \cup \dg(\alpha_1) \subseteq \alpha_2: \dg(\alpha_2) \subseteq \dg(\alpha_1)]$.

  \commentOC{Intuition. Invoke $\dg$ at $\alpha$ only when $\alpha$ fails. Permit to invoke a previous defense again as maybe the counter was too long to perform when summarized. If $a_2$ fails, we must invoke $\dg(…, a_2, x)$; failures of $\dg(…, a_2, …, x)$ should not count. We should invoke $\dg$ only when $\alpha$ fails with some defense produced by $\dg()$ belongs to $\histend$.}

  %Falsification seems done when $\exists \alpha, \beta \suchthat d(\beta) \in \histend \land \alpha \ileadsto p$. But then seems to be required to tell $d$ what has been tried already!}

  An argumentative theory claims that its argumentation strategy, embodied by $\arga$ and $\dg$, will convince the individuals in $I$.
  More precisely, given $i \in I$, it claims that $\arga$ either is a decisive argument for $\pga$, or, if $\arga$ is not, that the theory will defend the argument by replying to every counter-argument to $\arga$ with an argument that either is a decisive argument for $\pga$ or, in turn, will be defended by the theory, and so on as long as the theory has new counter-arguments to produce; the theory claiming that some argument along that chain is decisive.

  \commentOC{Condition: $\hist \nileadsto \pga \land \exists \hist' \subseteq \hist \suchthat \dg(\histp) \in \histend \land \dg(\hist)$ is not decisive ($\exists \beta \subseteq \hist \suchthat \dg(\hist) \in \beta_\text{end} \land \beta \ileadsto \pga$).}

  To define this formally, note that either $\arga$ is a decisive argument for $\pga$, or some paths challenge this claim, namely those $\hist \in \allhists$ such that $[\arga \in \histend \land \hist \nileadsto \pga]$.
  Generalizing this observation, given $i \in I$, define $\dgi(\ar) = \set{\dg(\ar, \hist) \suchthat \ar \in \histend \land \hist \nileadsto \pga} \subseteq \dg(\ar)$ as the subset of defenders of $\ar$ required for convincing $i$ of choosing $\pga$.
  Define $\gargsi \subseteq \allargs$ as the transitive closure of $\arga$ under $\dgi$: $\gargsi = \cup_{k \in \N} (\dgi)^k(\arga)$. The set $\gargsi$ is the set of arguments required to convince $i$ of choosing $\pga$.

  An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
  It is said to be truthful iff this claim is correct.
  More precisely, its claims are about paths of the sort $\hist = (\arga, …, \ar_2, …, \ar_3, …)$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …
  The following definition makes this precise by mandating that whenever increasing subsequences of $\hist$, involving arguments $\ar_j$ proposed by the theory, are not convincing, then the theory is able to produce yet another argument in defense of its claim.%
  \footnote{Intuition.
    If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
    If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
    $(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
    $\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
  \begin{definition}[Truthfulness]
    \label{def:truth}
    An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\allleadsto})$ is \emph{truthful} iff
    $\forall i \in I, \hist \in \allhists, n \in \N, k \in \intvl{1, \card \hist}^n, \ar \in \allargs^n$:
    \begin{equation}
      \left.\begin{aligned}
         & \ar_1 = \arga                                                                            \\
         & \forall 1 ≤ j < n: \ar_{j + 1} = \dg(\ar_j, α_{\intvl{1, k_j}}) \in \histbend{k_{j + 1}} \\
         & \forall 1 ≤ j ≤ n: \restr{\hist}{\intvl{1, k_j}} \nileadsto \pga
      \end{aligned}\right\}
      ⇒ \dg(\ar_n, \hist) ≠ \emptyset.
    \end{equation}
  \end{definition}
  Given the paths $\alpha, \alpha'$, the sequence $(\alpha, \alpha')$ denotes a path constituted of the arguments in $\alpha$ followed by the arguments of $\alpha'$.
  \begin{axiom}[Answerability-1]
    $\forall …$:
    \begin{equation}
      \left.\begin{aligned}
         & (\ar_4, \ar_2, \ar) \nileadsto \pga         \\
         & (\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
         & (\ar_5, \ar_2, \ar) \nileadsto \pga
      \end{aligned}\right\}
      ⇒ (\ar_5, \ar_2, \ar, \ar_3) \ileadstost \pga.
    \end{equation}
  \end{axiom}
  Requires excluding:
  \begin{equation}
    \left.\begin{aligned}
       & (\ar_4, \ar_2, \ar) \nileadsto \pga               \\
       & (\ar_4, \ar_2, \ar, \ar_3, \ar') \ileadsto \pga   \\
       & (\ar_5, \ar_2, \ar) \nileadsto \pga               \\
       & (\ar_5, \ar_2, \ar, \ar_3, \ar') \nileadsto \pga.
    \end{aligned}\right\}
  \end{equation}

  To exclude:
  \begin{equation}
    \left.\begin{aligned}
       & (\ar_4, \ar_2, \ar) \nileadsto \pga         \\
       & (\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
       & (\ar_5, \ar_2, \ar) \ileadsto \pga          \\
       & (\ar_5, \ar_2, \ar_3, \ar') \nileadsto \pga
    \end{aligned}\right\}
  \end{equation}
  Using:
  \begin{equation}
    \left.\begin{aligned}
       & (\ar_4, \ar_2, \ar) \nileadsto \pga         \\
       & (\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
       & (\ar_5, \ar_2, \ar) \ileadsto \pga
    \end{aligned}\right\}
    ⇒ (\ar_5, \ar_2, \ar_3) \ileadstost \pga.
  \end{equation}

  \begin{axiom}[Answerability]
    \label{ax:ans}
    $\forall i \in I, \hist \in \allhists, \ar \in \gargsi$:
    \begin{equation}
      \left.\begin{aligned}
         & \hist \nileadsto \pga                                        \\
         & (\hist, \ar) \ileadsto \pga                                  \\
         & \forall \ar' \in \allargs: (\hist, \ar, \ar') \ileadsto \pga
      \end{aligned}\right\}
      ⇒ \ar \ileadstost \pga.
    \end{equation}
  \end{axiom}
  \footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
    Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
    Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?
    In summary, we want to avoid that $a$ is not decisive but we can’t observe it when starting with some $α$ even though $α$ has also no decisive arguments: if $α$ has no decisive argument, then it must be possible to counter it.

    In fact, we might need (a weakening of) the following. If $a_1$ is “decisive” when starting with some $α$ (meaning $(α, a_1, a')$ YES for all arguments $a'$), then $a_1$ is decisive. Hopefully: if $(α, a')$ YES for all arguments $a'$, then some $a \in α$ is decisive.}
  \footnote{
    There is a problem. Consider the following may hold for some $i$. $a_1, a_5, a_2$ not convincing, $a_1, a_4$ not convincing, whil $a_1, a_5$ and $a_1, a_4, a_2$ and other variants convincing. Then for all $i$ one of $a_1$ or $a_2$ will appear to be decisive although for those people neither is.

    Problem might be solved by observing that $(a_1, a_5, a_2, a_1)$ convincing means that $a_1$ is decisive after all; and not convincing means that we have observed not truthfulness.

    We want that if the theory is invalid (thus some $i$ is not convinced), then either axiom fails or the theory is experimentally falsifiable.
    In the following example, the theory seems invalid but the axiom holds and the theory can’t seem to be defeated experimentally.

    Imagine $i$ thinks as follows. $a_1$ counter $a_5$; $a_2$ counters $a_4$; but $a_4$ and $a_5$ together make $a_1$ and $a_2$ invalid, so that $(a_4, a_5)$ counter both $a_1$ and $a_2$ (and also $a_1$ and $a_2$ together). Also assume that $i$ has short memory so that $a_1, a_4, a_5$ does not convince ($a_4$ and $a_5$ win) while $a_4, a_5, a_1$ do ($a_1$ wins as $a_4$ is too far away). BUT $a_5, a_4, a_1$ should then not convince.

    The solution is: after $a1, a5, a2$, $a1$ appears decisive, thus also $a1,a5, a2, a1, a4$ ok. Then some subset $a$ of $a1 \cup a_5 \cup a2$ should satisfy $a, a4$ OK and replace $a1$. And $a$ is decisive.
  }

  \begin{axiom}[Answerability - using newly built arguments (newest so far)]
    \label{ax:ansn}
    $\forall i \in I, \hist^j \in (\allhists)^n$:
    \begin{equation}
      [\forall j: \hist^j \ileadstost \pga] ⇒ [\exists \ar \in \allargs \suchthat \forall j: (\hist^j, \ar) \ileadstost \pga].
    \end{equation}
  \end{axiom}

  \begin{axiom}[Reinstatement]
    \label{ax:r}
    $\forall i \in I, \hist \in \allhists, \ar_1, \ar_2 \in \gargsi, \ar \in \allargs$:
    \begin{equation}
      \left.\begin{aligned}
         & (\hist, \ar_1, \ar) \nileadsto \pga            \\
         & (\hist, \ar_1, \ar, \ar_2) \ileadsto \pga      \\
         & (\hist, \ar_1, \ar, \ar_2, \ar) \ileadsto \pga
      \end{aligned}\right\}
      ⇒
      \left\{\begin{aligned}
         & (\hist, \ar_2) \ileadsto \pga                                                      \\
         & (\hist, \ar_2, \ar) \ileadsto \pga                                                 \\
         & \forall \ar': (\hist, a_2, a') \nileadsto \pga ⇒ (\hist, a_1, a') \nileadsto \pga.
      \end{aligned}\right.
    \end{equation}
  \end{axiom}
  \begin{example}[Reinstatement is required]
    Consider $a_1$ not decisive because of $a$ and then $(a_1, a, a_2) \ileadsto p$. But still $a_2$ not decisive: $a2 \nileadsto p$.
  \end{example}

  A path can be shortened to its first $k \in \intvl{0, \card{\hist}}$ elements, which is denoted by $\restr{\hist}{\intvl{1, k}}$, and “shortening” a path beyond its length yields the same path again, thus given $k \in \N$, $\restr{\hist}{\intvl{1, k}} = \restr{\hist}{\intvl{1, \min\set{\card{\hist}, k}}}$.
  Define a path as being decisive for $p$, $\hist \ileadstost p$, iff $\hist \ileadsto p \land \forall \ar \in \allargs: (\hist, \ar) \ileadsto p$, or equivalently, iff $\forall \hist' \suchthat \card{\hist} ≤ \card{\hist'} ≤ \card{\hist} + 1 \land \restr{\hist'}{\intvl{1, \card{\hist}}} = \hist: \hist' \ileadsto p$. Thus, $\hist \nileadstost p ⇔ \exists \ar \in \allargs \suchthat (\hist, \ar) \nileadsto p$ (or…?).

  \begin{axiom}[Self-sufficiency of positive arguments]
    $\forall \ar_1, \ar_2 \in \gargsi, \ar \in \allargs, \hist \in \allhists$:
    \begin{equation}
      \left.\begin{aligned}
         & (\hist, \ar_1) \nileadstost \pga \\
         & (\hist, \ar_2) \nileadstost \pga
      \end{aligned}\right\}
      ⇒ (\hist, \ar_1, \ar_2) \nileadstost \pga.
    \end{equation}
  \end{axiom}

  \begin{theorem}[Validity]
    If an argumentative theory $(\pga, \arga, \dg)$ is truthful and satisfy \cref{ax:ans,ax:r} and $\allargs$ is finite, then the theory is correct.
  \end{theorem}
  \begin{proof}
    Consider any $i \in I$.
    We have to show that $\pga \in \Phi_i$.
    Assume depth $3$.
    Assume that $\arga$ and $\ar_2$ are not decisive. Let’s show that $\ar_3$ is decisive.
    Assume that $(a_4, \arga, \ar_2, a) \nileadsto \pga$.
    By TF, $(a_4, \arga, \ar_2, a, \ar_3) \ileadstost \pga$.
    By Reinst, $(a_4, \arga, \ar_2, \ar_3) \ileadstost \pga$.
    By Self-s, $(a_4, \arga, \ar_3) \ileadstost \pga$.
    By Self-s, $(a_4, \ar_3) \ileadstost \pga$.

    Now consider $(\ar_5, \ar_2) \ileadsto \pga$. (Or any such path not ending with $\ar_3$.)
    We want $(\ar_5, \ar_2, \ar_3) \ileadstost \pga$.
    By Answ, $\exists a \suchthat (\ar_5, \ar_2, a) \nileadsto \pga$.
  \end{proof}
  \begin{proof}[Sketch of proof]
    Consider any $i$. We have to show that $\pga \in \Phi_i$.%
    \footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$).
      By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
      We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
      By reinst: $(\ar_2)$ YES.}

    Assume that $\arga$ is not decisive (otherwise, done). Thus, for some $\hist: (\hist, \arga, \ar) \nileadsto \pga$ or $(\hist, \ar, \arga)$.
    Consider that $\hist: (\hist, \arga, \ar) \nileadsto \pga$.
    Consider $\hist'$ and wonder whether $(\hist', \arga, \ar) \ileadsto \pga$. If so,
    By Answ, this means that $\forall \hist, (\hist, \arga, \ar) \nileadsto \pga$.
    By truthfulness, $\ar_2 = \dg(\ar_1, (\hist, \arga, \ar)) ≠ \emptyset \land (\hist, \arga, \ar, \ar_2) \ileadsto \pga$.
    Assume now that $(\hist, \arga, \ar, \ar_2, …) \ileadsto \pga$.
    Then by Reinstatement, $(\hist, \ar_2) \ileadsto \pga$.

    %	Assume, to start with, that the theory is of depth 2, more precisely, $\arga$ is not decisive because $\hist \nileadsto \pga, \arga \in \histend$ and $\dgi(\arga) = \ar_2$ with no further hist against it.
  \end{proof}

  This definition of truthfulness is adopted because of our goal of distinguishing empirical claims. Truthfulness, so defined, can be falsified empirically: if it does not hold, it is possible to prove this empirically by exhibiting a counter-example in the form of a tuple $i, k, (a_j)_{j \in \intvl{1, k}}, \alpha$ such that … (to be completed).

  Note that a static theory $(p, \ar)$ can be considered as an argumentative theory $(p, \ar, \emptyset)$, and $(p, \ar)$ is truthful in the sense of \cref{sec:static} iff $(p, \ar, \emptyset)$ is truthful in the sense of \cref{def:truth}.

  Associated with an argumentative theory also comes the following axiom, that mandates a more general version of the following requirement: if some argument $\ar_2$ defends some argument $\ar_1$, and if $\ar_2$ leads $i$ to opt for $\pga$ whenever it immediately follows $\ar_1$, then omitting $\ar_1$ would still have lead $i$ to opt for $\pga$.

  \begin{axiom}[Omitting defended arguments]
    \label{ax:omitting}
    If the theory is truthful, then $\forall i \in I, k \in \N^*, l \in \intvl{2, k}, (\ar_j)_{j \in \intvl{1, k}} \in \allargs^k, \hist, \beta \in \allhists \suchthat [\ar_k \in \histend \land \forall j \in \intvl{2, k}: \ar_j \in \dgi(\ar_{j - 1})]:
      (\beta, \ar_{l - 1}, \ar_l, \hist) \ileadsto \pga ⇒ (\beta, \ar_l, \hist) \ileadsto \pga$.
  \end{axiom}
  This is called an axiom because it cannot be falsified using observables. It has to be accepted on the basis of the rest of the observations and on the contents of the defending arguments. As \cref{ex:lichtenstein} suggests, theories can be built so as to make this axiom plausible by embedding within defending arguments the content of the defended argument.

  An argumentative theory is said to be correct iff $\forall i \in I: \pga \in \iprops$.
  As previously, it is said to be valid iff its truthfulness leads to its correctness, except that the above axiom needs now be included as a price for the added flexibility of the theory.
  \begin{theorem}[Validity]
    If an argumentative theory is truthful, then \cref{ax:omitting} implies that the theory is correct.
  \end{theorem}
  \begin{proof}[Sketch of proof]
    Assume, to start with, that the theory is of depth 2, more precisely, $\arga$ is not decisive because $\hist \nileadsto \pga, \arga \in \histend$ and $\dgi(\arga) = \ar_2$ with no further hist against it.

    Consider any $i \in I$.
    Assume $\arga$ is not decisive.
    Consider any $\hist \suchthat \ar_2 \in \histend$. We want to show that $\hist \ileadsto \pga$.

    IF $\hist$ contains $\arga$ before $\ar_2$, done by TF.
    Otherwise, define $hist'$ which has a1 just before a2.
    Apply TF on hist' then apply axiom.

    Consider now a theory of depth 3.
    Consider any $i \in I$.
    Assume $\arga$ not decisive, $\ar_2$ not decisive.
    Consider any $\hist \suchthat \ar_3 \in \histend$. We want to show that $\hist \ileadsto \pga$.

    If it includes a1 then a2, ok thanks to truthfulness.
    Otherwise, build hist3 that has a1 then a2 just before a3.
    That’s ok thanks to TF.
    By axiom, hist2 with only a2 just before a3 is thus ok.
    By axiom, hist is thus ok.
  \end{proof}

  \appendix
  \section{Supplementary material}
  \label{sec:supplementary}
  The following example presents a situation where non rejection is due to finiteness of observations.
  \begin{example}[Big disjunct, non rejectable]
    \label{ex:bigor}
    Consider $\ar \in \allargs$.
    The claim $[\forall i: (\ar) \ileadstog \phi] \lor [\forall i: (\ar, \ar) \ileadstog \phi] \lor …$,
    thus, $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$,
    is non rejectable: $\Theta \wsub C^\wsub$.

    Given any $θ \in \Theta$, define $l = \max_{\hist \suchthat \exists i \in I, p \in P \suchthat (i, \hist, p) \in θ} \card{\hist} + 1$ if $\exists p \in P \suchthat (i, \hist, p) \in θ$, and $l = 1$ otherwise. Pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have
    $\theta \in \powerset{\allleadstog}$ and ${\allleadstog} \in C$, thus, $\theta \in C^\wsub$.
  \end{example}
  \begin{example}[Big disjunct unfals iff infinite]
    \label{ex:existsParticular}
    Given any $\ar \in \allargs$ and an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is strongly unfalsifiable: any $θ \in \Theta$ only has finitely many $i$ in its triples, thus, some $i$ is not reached by $\theta$, thus $\theta \in C^\wsub$.
    If $I$ is finite, then the claim is falsifiable: if it is wrong, observe for each $i$ some adequate triple $(i, \hist_i, p ≠ \phi)$.
  \end{example}

  \begin{example}[Three functions, may be rejectable]
    \label{ex:threeFs}
    Consider $i \in I$, $ar \neq ar' \in \allargs$ and the three paths $\hists = \set{\emptyset, (\ar), (\ar')}$.
    The claim “Exactly two of the three paths in $\hists$ lead $i$ to $\phi$ and the remaining one leads $i$ to $¬\phi$” is not rejectable iff ${\ileadsto}$ associates, in order, $(\emptyset, (\ar), (\ar'))$ to either
    $(¬\phi, \phi, \phi)$,
    $(\phi, ¬\phi, \phi)$,
    $(\phi, \phi, ¬\phi)$,
    $(\phi, \phi, \phi)$ or
    $(\phi, ¬\phi, ¬\phi)$.

    Indeed, if ${\ileadsto}$ satisfies this, then either it holds and by \cref{th:ex-post} it is not rejectable, or it associates this triplet to $(\phi, \phi, \phi)$ or
    $(\phi, ¬\phi, ¬\phi)$ in which case, as $(\ar)$ and $(\ar')$ cannot be together in $\theta$, any $\theta \in \Theta$ with $\theta \sub {\allleadsto}$ satisfy $\theta \in C^\wsub$.

    Conversely, if $C$ is not rejectable, then for any $\theta_1$ containing $\emptyset$ and $(\ar)$ and $\theta_2$ containing $\emptyset$ and $(\ar')$, with $\theta_1, \theta_2 \sub {\allleadsto}$, we have $\theta_1, \theta_2 \in C^\wsub$, which requires that ${\ileadsto}$ associates $(\emptyset, (\ar))$ to $(¬\phi, \phi)$, $(\phi, ¬\phi)$ or $(\phi, \phi)$ and $(\emptyset, (\ar'))$ to $(¬\phi, \phi)$, $(\phi, \phi)$ or $(\phi, ¬\phi)$.
  \end{example}

  \begin{example}[Disjunct with incompatible paths, not falsifiable, to be deleted]
    \label{ex:disj}
    Given $\ar \neq \ar' \in \allargs$, the claim appearing in \cref{ex:disjR}, namely,
    $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$, is not falsifiable.
    Consider any ${\allleadstog} \notin C$, thus satisfying, for some $i$, $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
    Consider any $θ \subseteq {\allleadstog}$ and let us prove that $\theta \in C^\wsub$.
    (This is a consequence of \cref{ex:disjR} so this proof should not be useful.)
    By definition of a general observation, $θ$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $θ$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have
    $\theta \in \powerset{\allleadstog'}$ and ${\allleadstog'} \in C$, thus, $\theta \in C^\wsub$.
  \end{example}

  \begin{example}
    \label{ex:anon}
    Given $\ar$ and $\phi$, the claim $\ar \ileadstog \phi$ is not anonymous iff $\card{I} ≥ 2$. Indeed, pick any $j ≠ i$ and define $\sigma(i) = j$, $\sigma(j) = i$ and $\forall k \in I \setminus \set{i, j}: \sigma(k) = k$, then $C = \set{{\allleadstog} \suchthat {\allleadstog}(i, \ar) = \phi}$, which contains in particular ${\allleadsto}$ defined as $\forall \hist: [\forall k \in I \setminus \set{j}, {\allleadsto}(k, \hist) = \phi \land {\allleadsto}(j, \hist) = ¬\phi$; while $C \circ \sigma = \set{{\allleadstog} \suchthat {\allleadstog}(j, \ar) = \phi}$ which does not contain ${\allleadsto}$.
  \end{example}

  \begin{example}
    Consider $I = \set{i}$ and $\ar_1 \neq \ar_2 \in \allargs$.
    The claim $C = \set{{\allleadstog} \in \allleadstos \suchthat (\forall \hist \neq \emptyset: \hist \ileadstog \phi) \lor (\forall \hist \neq \emptyset: \hist \ileadstog ¬\phi)}$ is not falsifiable.

    Consider ${\allleadstog}$ defined as $\hist \ileadstog \phi$ iff $\hist = \emptyset \lor \hist_1 = \ar_1$ and $\hist \ileadstog ¬\phi$ otherwise.
    We have ${\allleadstog} \notin C$.
    But any $θ \subseteq {\allleadstog}$ satisfies $\theta \in C^\wsub$.
  \end{example}

  \begin{example}
    \label{ex:falsAnonIfBig}
    Consider $I = \set{i_1, i_2}$ and $\ar_1 \neq \ar_2 \in \allargs$.
    Given $p_0, p, p'_0, p' \in P$,
    define ${\allleadstog^{i_1 \mapsto (p_0, p); i_2 \mapsto (p'_0, p')}}$ as $\emptyset \ileadstog[i_1] p_0 \land \emptyset \ileadstog[i_2] p'_0 \land \forall \hist \in \allhists \setminus \set{\emptyset}: \hist \ileadstog[i_1] p \land \hist \ileadstog[i_2] p'$.

    The claim $C = \set{{\allleadstog^{i_1 \mapsto (p_0, p); i_2 \mapsto (p'_0, p')}} \suchthat p_0, p, p'_0, p' \in \set{\phi, ¬\phi}}$ is not falsifiable.
    Consider ${\allleadstog}$ defined as, $\forall i \in I$, $\hist \ileadstog \phi$ iff $\hist = \emptyset \lor \hist_1 = \ar_1$ and $\hist \ileadstog ¬\phi$ otherwise.
    We have ${\allleadstog} \notin C$.

    Consider now any bigger claim $C'$, thus with $C \wsub C'$.
    I believe that $C'$ is falsifiable iff it includes all relations that are constant on all paths starting with a given argument (thus all $\allleadstog$ such that $\forall \hist, \histp \in \allhists: \hist_1 = \histp_1 \implies {\ileadstog}(\hist) = {\ileadstog}(\histp)$).
  \end{example}

  \begin{example}
    Consider $I = \set{i_{p_0, p} \suchthat p_0, p \in \set{\phi, ¬\phi}}$ (thus $\card{I} = 4$).
    Given $p_0, p \in P$,
    define $R^{(p_0, p)} = R^{\emptyset \mapsto p_0; p}: \allhists \rightarrow P$ as $R(\emptyset) = p_0 \land \forall \hist \in \allhists \setminus \set{\emptyset}: R(\hist) = p$.

    Define ${\allleadsto}$ as $\forall p_0, p \in \set{\phi, ¬\phi}: {\ileadsto[i_{p_0, p}]} = R^{\emptyset \mapsto p_0; p}$.

    Given $f: I \rightarrow \set{\phi, ¬\phi}^2$,
    define ${\allleadstog^f}$ as $\forall i \in I: {\allleadstog^f_i} = R^{f(i)}$.
    Note that any $f \in (\set{\phi, ¬\phi}^2)^I$ can be seen as a permutation $\sigma$ over $I$ and that ${\allleadstog^f} = {\allleadsto} \circ \sigma$.

    The claim $C = \set{{\allleadstog^f} \suchthat f \in (\set{\phi, ¬\phi}^2)^I}$ is falsifiable: consider any ${\allleadstog}$ observably compatible with $C$; if there is $\hist^1, \hist^2, \hist^3 \in \allhists \setminus \set{\emptyset}$ and $i_1 \neq i_2 \neq i_3 \neq i_1 \in I$ with $\forall k \in \intvl{1, 3}: \hist^k \ileadstog[i_k] \phi$, then ${\allleadstog}$ is not observably compatible with $C$, so there must be at most 2 individuals mapping some non empty path to $\phi$ (thus at least 2 mapping all non empty paths to $¬\phi$) and for the same reason, at least 2 individuals mapping all non empty paths to $\phi$; thus (together with noting that two individuals mapping every paths to $\phi$ is not observably compatible with $C$ and adding the missing bits), ${\allleadstog} \in C$.
  \end{example}

  \begin{example}
    Consider $I = \cup_{p_0, p \in \set{\phi, ¬\phi}} I_{p_0, p}$, with each $I_{p_0, p}$ infinite, and $\ar_1 \neq \ar_2 \in \allargs$.
    Given $p_0, p \in P$,
    define $R^{(p_0, p)} = R^{\emptyset \mapsto p_0; p}: \allhists \rightarrow P$ as $R(\emptyset) = p_0 \land \forall \hist \in \allhists \setminus \set{\emptyset}: R(\hist) = p$.
    Define ${\allleadsto}$ as $\forall p_0, p \in \set{\phi, ¬\phi}, i \in I_{p_0, p}: {\ileadsto} = R^{\emptyset \mapsto p_0; p}$.

    The claim $C = \set{{\allleadsto} \circ \sigma \suchthat \sigma: I \leftrightarrow I}$ is not falsifiable:
    consider ${\allleadstog}$ defined as, $\forall i \in I$, $\hist \ileadstog \phi$ iff $\hist = \emptyset \lor \hist_1 = \ar_1$ and $\hist \ileadstog ¬\phi$ otherwise; we have ${\allleadstog} \notin C$
    but any $θ \in \Theta, \theta \subseteq {\allleadstog}$ satisfies $\theta \in C^\wsub$.

    Now consider $C' = \set{{\allleadstog} \suchthat \forall i \in I, \hist, \histp \in \allhists: \hist_1 = \histp_1 \implies {\ileadstog}(\hist) = {\ileadstog}(\histp)}$. (Thus $C \wsub C'$.)
    This one is falsifiable: if ${\allleadstog} \notin C$, then $\exists i \in I, \hist, \histp \suchthat \hist_1 = \histp_1 \land {\ileadstog}(\hist) ≠ {\ileadstog}(\histp)$, now define $\histpp$ as the common prefix to $\hist$ and $\histp$; as ${\ileadstog}(\hist) ≠ {\ileadstog}(\histp)$, we have at least one of ${\ileadstog}(\hist) ≠ {\ileadstog}(\histpp)$ or ${\ileadstog}(\histp) ≠ {\ileadstog}(\histpp)$, and as $\histpp$ satisfies $\histpp_1 = \hist_1$ and $\histpp_1 = \histp_1$, observing the corresponding pair of paths for $i$ falsifies $C$.
  \end{example}

  \begin{example}
    Imagine that the real relation is composed of individuals who map $(a_1, a_2)$ to $\phi$ and $(a_1, a_3)$ to $¬\phi$ and individuals converse (with both, say, mapping everything else to $0$).
    Then the claim $C = \set{{\allleadstog} \suchthat \forall i \in I: {\ileadstog}(\ar_1, \ar_2) \in \set{\phi, ¬\phi} \land {\ileadstog}(\ar_1, \ar_3) \in \set{\phi, ¬\phi}}$ is falsifiable.

    Imagine that the real relation is composed of individuals who map $(a_1, a_2)$ to $\phi$ and $(a_1, a_3)$ to $¬\phi$ and individuals converse (with both, say, mapping everything else to $0$) and individuals who map $(a_4, a_5)$ to $\phi$ and $(a_6)$ to $¬\phi$ and individuals converse, with both mapping everything else to $0$.
    Then the claim $C = \set{{\allleadstog} \suchthat \forall i \in I: [{\ileadstog}(\ar_1, \ar_2) \in \set{\phi, ¬\phi} \land {\ileadstog}(\ar_1, \ar_3) \in \set{\phi, ¬\phi} \land {\ileadstog}(H) = 0] \lor [{\ileadstog}(\ar_4, \ar_5) \in \set{\phi, ¬\phi} \land {\ileadstog}(\ar_6) \in \set{\phi, ¬\phi} \land {\ileadstog}(H) = 0]}$, where $H$ designates all “other” paths (abusing notation), is falsifiable.

    Imagine that the real relation is composed of individuals who map $(a_1, a_2)$ to $\phi$ and $(a_1, a_3)$ to $¬\phi$ and individuals converse (with both, say, mapping everything else to $0$) and individuals who map $(a_1, a_4)$ to $\phi$ and $(a_6)$ to $¬\phi$ and individuals converse, with both mapping everything else to $0$.
    Then the claim $C = \set{{\allleadstog} \suchthat \forall i \in I: [{\ileadstog}(\ar_1, \ar_2) \in \set{\phi, ¬\phi} \land {\ileadstog}(\ar_1, \ar_3) \in \set{\phi, ¬\phi} \land {\ileadstog}(H) = 0] \lor [{\ileadstog}(\ar_1, \ar_4) \in \set{\phi, ¬\phi} \land {\ileadstog}(\ar_6) \in \set{\phi, ¬\phi} \land {\ileadstog}(H) = 0]}$, where $H$ designates all “other” paths (abusing notation), is falsifiable: could have an individual mapping both $(\ar_1, \ar_2)$ and $(\ar_1, \ar_4)$ to $0$ but then if she maps $\ar_6$ to $0$, observe both $(\ar_1, \ar_2)$ and $\ar_6$ and falsify $C$; if she maps $\ar_6$ in $\set{\phi, ¬\phi}$, both $(\ar_1, \ar_4)$ and $\ar_6$ and falsify $C$ (here assuming falsely in the whole example that $\ar_6$ can be observed together with $(\ar_1, .)$, so we probably should replace $\ar_6$ by $\emptyset$).
  \end{example}

  \subsection{More examples}
  \begin{example}[Partly falsifiable is still unfalsifiable]
    \label{ex:partlyU}
    Given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
    the claim
    $C = \set{{\allleadstog} \in \allleadstos \suchthat \hist \ileadstog[j] \phi \land \forall i \in I: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]}$ is unfalsifiable but not strongly unfalsifiable.

    Define
    $C_1 = \set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]}$ and
    $C_2 = \set{{\allleadstog} \in \allleadstos \suchthat \hist \ileadstog[j] \phi}$,
    so that $C = C_1 \cap C_2$.
    Observe that $C_1$ is the claim of \cref{ex:disj} and $C_2$ is falsifiable and coverable. 
    \Cref{th:strongUToStrictU} concludes.
  \end{example}

  \begin{example}[Partly again]
    Considering again \cref{ex:partlyU},
    given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
    $C = \set{{\allleadstog} \suchthat \hist \ileadstog[j] \phi \land \forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]}$ is not falsifiable but \cref{th:veryunfals} is not applicable.
    Observe that
    $C^\subseteq = \set{R \in I × \allhists × P \suchthat [(j, \hist) \in R^{-1}(P) \implies (j, \hist, \phi) \in R] \land \forall i: [(i, (\ar)) \in R^{-1}(P) \land (i, (\ar')) \in R^{-1}(P)] \implies [(i, (\ar), \phi) \in R \land (i, (\ar'), \phi) \in R]}$ and 
    $C^\subseteq \cap \Theta = \set{\theta \in \Theta \suchthat (j, \hist) \in \theta^{-1}(P) \implies (j, \hist, \phi) \in \theta}$.
    % Pick any $\allleadstog$ satisfying $\hist \ileadstog[j] \phi$ and $\exists i \suchthat (\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
    % The definition of non-falsifiability applies: pick any $\theta \subseteq {\allleadstog}$, we have $\theta \in C^\subseteq$ as $(j, \hist, \phi)$ is compatible with $\theta$ and the other condition excludes nothing in $\Theta$.
    % We can apply Thm 4 similarly: given any $O$, we have $\restr{\allleadstog}{O} \in C^\subseteq$.
    % We should be able to apply a corollary similarly.
    The relations obs-compatible with $C^\subseteq \cap \Theta$ are those satisfying $\hist \ileadstog[j] \phi$.
    We conclude that $C$ is not falsifiable as $C$ is smaller than that set.
  \end{example}

  \begin{example}[Come back]
    In \cref{ex:allParticular} we considered a given $\ar$ and $\phi$ and defined the falsifiable claim $C = \set{{\allleadstog} \suchthat \forall i: \ar \ileadstostg \phi}$.
    Observe that $C^\subseteq \cap \Theta = \set{\theta \in \Theta \suchthat \forall \hist \suchthat \ar \in \histend: \theta(i, \hist) = \emptyset \lor \theta(i, \hist) = \phi}$.
    The relations obs-compatible with $C^\subseteq \cap \Theta$ are exactly those in $C$.

    \Cref{ex:existsParticular} assumed an infinite $I$, in which case the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
    Observe that $C^\subseteq \cap \Theta = \set{\theta \in \Theta \suchthat \exists i \suchthat \forall \hist \suchthat \ar \in \histend: \theta(i, \hist) = \emptyset \lor \theta(i, \hist) = \phi}$.
    Any relation ${\allleadstog}$ is obs-compatible with that set as $C^\subseteq \cap \Theta = \Theta$: given any $\theta \in \Theta$, there is an $i$ such that $\forall \hist: \theta(i, \hist) = \emptyset$ as $\theta$ must have a finite support.
  \end{example}

  \begin{example}
    \label{ex:cardPart}
    Pick $i \in I$, $p ≠ p' \in P$ and $\hist, \histp \in \allhists$ with $\hist$ and $\histp$ non prefix related.
    The claim $\psi = (\hist \ileadstog p \lor \hist \ileadstog p') \land (\histp \ileadstog p \lor \histp \ileadstog p') \land (\forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \histpp \ileadstog[j] p)$ is falsifiable.

    Given $q, q' \in P$, let ${\allleadstog^{q, q'}}$ be defined as $\hist \ileadstog^{q, q'} q$, $\histp \ileadstog^{q, q'} q'$ and $\forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \histpp \ileadstog[j]^{q, q'} p$.
    Note that $C_\psi = \set{{\allleadstog^{p, p}}, {\allleadstog^{p, p'}}, {\allleadstog^{p', p}}, {\allleadstog^{p', p'}}}$.

    Then, $C_\psi^\wsub \cap \Theta =
      \set{\theta \in \Theta \suchthat \theta(\hist, i) \in \set{\emptyset, p, p'} \land \theta(\histp, i) \in \set{\emptyset, p, p'} \land \forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \theta(\histpp, j) \in \set{\emptyset, p}}$.
    % and we see that $C_\psi^\wsub \cap \Theta =
    % \set{\theta \in \Theta \suchthat [\theta(\hist, i) = \theta(\histp, i) = \emptyset \lor \theta(\hist, i) \in \set{p, p'} \lor \theta(\histp, i) \in \set{p, p'}] \land \forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \theta(\histpp, j) \in \set{\emptyset, p}}$.

    Any obs-compatible relation has to satisfy $\forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \histpp \ileadstog[j] p$.
    It also must satisfy ${\allleadstog}(\hist, i) \in \set{p, p'}$, for otherwise, with $\theta = \set{(\hist, i, {\allleadstog}(\hist, i))}$, we get $({\allleadstog} \cap \theta) \notin C_\psi^\wsub \cap \Theta$.
    Similarly, it must satisfy ${\allleadstog}(\histp, i) \in \set{p, p'}$.
    % are $\set{{\allleadstog} \in \allleadstos \suchthat [{\allleadstog}(\hist, i) \in \set{p, p'} \lor {\allleadstog}(\histp, i) \in \set{p, p'}] \land \forall (j, \histpp) \in (I × h) \setminus \set{(i, \hist), (i, \histp)}: \histpp \ileadstog[j] p}$, which, given $p \neq p^{\prime\prime} \neq p'$, contains ${\allleadstog^{p, p^{\prime\prime}}}$, contrary to $C_\psi$.
  \end{example}

  \begin{example}
    Consider any $i$, $\ar_1 ≠ \ar_2 \in \allargs$, $p, p' \in P$.
    The claim $C = \set{{\allleadstog} \suchthat (\ar_1) \ileadstog p \land (\ar_2) \ileadstog p'}$ is falsifiable.

    We have $C^\subseteq \cap \Theta = \set{\theta \suchthat (\theta((\ar_1), i) = \emptyset \lor \theta((\ar_1), i) = p) \land (\theta((\ar_2), i) = \emptyset \lor \theta((\ar_2), i) = p')}$.
    The obs-compatible relations are $\set{{\allleadstog} \suchthat (\ar_1) \ileadstog p \land (\ar_2) \ileadstog p'}$.
  \end{example}
  \begin{example}
    \label{ex:consti2}
    Assume that $\card{I} ≥ 2$.
    Consider any $p, p' \in P$ and the constant functions ${\allleadstog^p}$ to $p$ and ${\allleadstog^{p'}}$ to $p'$ and the claim $C = \set{{\allleadstog^p}, {\allleadstog^{p'}}}$.
    We have $C^\wsub \cap \Theta = \set{\theta \suchthat \theta(\theta^{-1}(P)) = \set{p} \lor \theta(\theta^{-1}(P)) = \set{p'}}$.
    Consider any relation $\allleadstog$ that is obs-compatible with $C^\wsub \cap \Theta$. Assume that for some $\hist, i$, $\hist \ileadstog p$. Then, pick any $j \neq i$, and observe that by obs-compatibility, for any $\hist'$, $\hist' \ileadstog[j] p$.
    It follows, still by obs-compatibility, that for any $\hist'$, $\hist' \ileadstog p$ (otherwise pick $\theta$ including $(\hist', i)$ and any $(\hist^{\prime\prime}, j)$). Conclude that $\allleadstog$ is constant to $p$.
    It follows that the obs-compatible relations are $\set{{\allleadstog^p}, {\allleadstog^{p'}}}$, thus, $C$ is falsifiable.
  \end{example}
  \begin{example}
    \label{ex:consti1}
    Assume that $\card{I} = 1$.
    Consider any $p, p' \in P$ and the constant functions ${\allleadstog^p}$ to $p$ and ${\allleadstog^{p'}}$ to $p'$ and the claim $C = \set{{\allleadstog^p}, {\allleadstog^{p'}}}$.
    We have $C^\wsub \cap \Theta = \set{\theta \suchthat \theta(\theta^{-1}(P)) = \set{p} \lor \theta(\theta^{-1}(P)) = \set{p'}}$.
    Consider any relation $\allleadstog$ that is obs-compatible with $C^\wsub \cap \Theta$. Assume that for some $\hist, i$, $\hist \ileadstog p$. Observe that by obs-compatibility, $\emptyset \ileadstog p$.
    It follows, still by obs-compatibility, that for any $\hist'$, $\hist' \ileadstog p$ (otherwise pick $\theta$ including $(\hist', i)$ and $(\emptyset, i)$). Conclude that $\allleadstog$ is constant to $p$.
    It follows that the obs-compatible relations are $\set{{\allleadstog^p}, {\allleadstog^{p'}}}$, thus, $C$ is falsifiable.
  \end{example}
  \begin{example}
    \label{ex:constwout0}
    Consider any $p, p' \in P$ and the claim $C = \set{{\allleadstog} \suchthat (\forall \hist \neq \emptyset, i \in I: \hist \ileadstog p) \lor (\forall \hist \neq \emptyset, i \in I: \hist \ileadstog p')}$. That claim is falsifiable iff $\card I ≥ 2$ or $\card \allargs = 1$. The proof if so is similar to \cref{ex:consti2}; otherwise, pick any $\ar \in \allargs$ and note that any relation satisfying $(a.*) \ileadstog p$ and $\forall b \neq a: (b.*) \ileadstog p'$ is obs-compatible with $C^\wsub \cap \Theta$ but is not in $C$.
  \end{example}
  \begin{example}
    \label{ex:unfor}
    Consider any $i$, $\hist, \histp \in \allhists$, $p, p' \in P$ where $\hist$ and $\histp$ are not prefix related.
    The claim $C = \set{{\allleadstog} \suchthat \hist \ileadstog p \lor \histp \ileadstog p'}$ is unfalsifiable.

    We have $C^\wsub \cap \Theta = \set{\theta \suchthat (\theta(\hist, i) = \emptyset \lor \theta(\hist, i) = p) \lor (\theta(\histp, i) = \emptyset \lor \theta(\histp, i) = p')}$.
    Observe that any $\theta \in \Theta$ satisfy $\theta(\hist, i) = \emptyset \lor \theta(\histp, i) = \emptyset$, thus, $C^\wsub \cap \Theta = \Theta$.
    It follows that the obs-compatible relations are $\allleadstos$, which is bigger than $C$.
  \end{example}
  \begin{example}
    Consider any $i$, $\ar \neq \ar' \in \allargs$, $p, p' \in P$.
    The claim $C = \set{{\allleadstog} \suchthat (\ar) \ileadstog p \lor (\ar') \ileadstog p'}$ is unfalsifiable (per \cref{ex:unfor}).

    Now consider the smaller claim $C' = \set{{\allleadstog} \suchthat (\ar_1) \ileadstog p}$ (thus $C' \ssub C$).
    It yields a smaller set ${C'}^\wsub \cap \Theta = \set{\theta \suchthat (\theta((\ar_1), i) = \emptyset \lor \theta((\ar_1), i) = p)}$.
    The obs-compatible relations are the smaller set $\set{{\allleadstog} \suchthat (\ar_1) \ileadstog p}$.
    Thus, $C'$ is falsifiable.

    Note that any claim $C^{\prime\prime}$ bigger than $C$ is unfalsifiable as well, unless trivial:
    from $C \wsub C^{\prime\prime}$, we deduce $\Theta = C^\wsub \cap \Theta \wsub {C^{\prime\prime}}^\wsub \cap \Theta$, thus ${C^{\prime\prime}}^\wsub \cap \Theta = \Theta$ and the obs-compatible relations are $\allleadstos$, which equals $C^{\prime\prime}$ only if $C^{\prime\prime}$ is trivial.

    Now consider the alternative smaller claim of \cref{ex:consti2}, $C^{(2)} = \set{{\allleadstog^p}, {\allleadstog^{p'}}}$ (thus $C^{(2)} \ssub C$).
    It yields a smaller set ${C^{(2)}}^\wsub \cap \Theta$.
  \end{example}
  \begin{example}
    \label{ex:for}
    Consider any $i$ and $p, p' \in P$.
    The claim $C = \set{{\allleadstog} \suchthat (\forall \hist: \card{\hist} \in 2\N \implies \hist \ileadstog p) \lor (\forall \hist: (\card{\hist} \in 2\N + 1 \lor \hist = \emptyset) \implies \hist \ileadstog p')}$, namely, all paths of even lengths lead $i$ to $p$ OR all paths of odd lengths or empty lead $i$ to $p'$, is falsifiable.

    Note that $C^\wsub \cap \Theta = \set{\theta \in \Theta \suchthat (\forall \hist: \card{\hist} \in 2\N \implies \theta(\hist) \in \set{\emptyset, p}) \lor (\forall \hist: (\card{\hist} \in 2\N + 1 \lor \hist = \emptyset) \implies \theta(\hist) \in \set{\emptyset, p'})}$.
    Consider any obs-compatible ${\allleadstog}$.
    Assume that $\exists \hist \suchthat \card{\hist} \in 2\N \land \hist \nileadstog p$.
    Pick $\theta = \set{(\emptyset, i, {\ileadstog}(\emptyset)), (\hist, i, {\ileadstog}(\hist))}$. We have $({\allleadstog} \cap \theta) \in C^\wsub \cap \Theta$, and ${\allleadstog} \cap \theta = \theta$, and from $\theta \in C^\wsub \cap \Theta$ we obtain ${\ileadstog}(\emptyset) = p'$.
    Similarly, using $\theta = \set{(\emptyset, i, {\ileadstog}(\emptyset)), (\histp, j, {\ileadstog[j]}(\histp))}$, we obtain that $\allleadsto$ conforms to the relation in $C$.
  \end{example}

  \begin{remark}
    Intuition (to be explored) is that when $C^\wsub \cap \Theta$ has the covering property, then the obs-compatible relations are easily deduced from the set (something like, just consider compatibility point per point, one singleton observation at a time). Covering is what is used above, something like $\exists i, \hist, \histp$ such that $\forall \hist^{\prime\prime}: [\forall j \neq i: \set{(\hist, i, .), (\histp, j, .)} \in C^\wsub \cap \Theta] \land [\exists j \neq i \suchthat \set{(\hist', i, .), (\hist^{\prime\prime}, j, .)} \in C^\wsub \cap \Theta]$.
  \end{remark}

  \subsection{Compatibility of claim properties}
  \begin{theorem}
    \label{th:hsif}
    For all exposure protocols, there exists a claim that holds and is strongly informative beyond counting and falsifiable iff $\exists i \suchthat P_i \neq \emptyset$.
  \end{theorem}
  \begin{proof}
    Define $C = \set{\allleadsto}$.
  \end{proof}
  \begin{theorem}
    \label{th:hif}
    For all exposure protocols, there exists a claim that holds and is informative beyond counting and falsifiable.
  \end{theorem}
  \begin{proof}
    Define $C = \set{\allleadsto}$.
  \end{proof}
  \begin{theorem}
    \label{th:asif}
    For all exposure protocols, there exists a claim that is anonymous, strongly informative beyond counting and falsifiable.
  \end{theorem}
  \begin{proof}
    Define the claim as $\forall i \in I, \hist \in \allhists: \hist \ileadstog \phi$.
  \end{proof}

  \begin{theorem}
    \label{th:ha}
    For all exposure protocols, there exists a claim that holds and is anonymous.
  \end{theorem}
  \begin{proof}
    Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}$.
  \end{proof}

  \begin{theorem}
    \label{th:HAinformativeness}
    Given any $Q \wsub P \cup \set{\emptyset}$,
    the following two conditions on $\allleadsto$ are equivalent:
    there exists a claim that holds and is anonymous and satisfies 
    $\exists j \in I \suchthat \pi(C)(j) \ssub Q$;
    and 
    $\set{P_i, i \in I} \ssub Q$.
  \end{theorem}
  \begin{proof}
    Consider any claim $C$ that holds and is anonymous, pick any $j \in I$, and let us show that $\set{P_i, i \in I} \wsub \pi(C)(j)$.
    Pick any $p \in \set{P_i, i \in I}$. 
    Pick any $i \in I$ such that $\pi({\allleadsto})(i) = p$.
    Define $\sigma$ as the permutation exchanging $i$ and $j$.
    As $C$ holds, it contains $\allleadsto$, and by anonymity, it contains ${\allleadsto} \circ \sigma$. Observing that $\pi({\allleadsto} \circ \sigma)(j) = \pi({\allleadsto})(i) = p$, we obtain $p \in \pi(C)(j)$.

    It follows that for any claim $C$ that holds and is anonymous, if $\exists j \in I \suchthat \pi(C)(j) \ssub Q$ then $\set{P_i, i \in I} \wsub \pi(C)(j) \ssub Q$ thus $\set{P_i, i \in I} \ssub Q$.

    Turning to the other direction, if $\set{P_i, i \in I} \ssub Q$, then pick any $p \in Q \setminus \set{P_i, i \in I}$; the claim $\forall i \in I: P_i \neq p$ satisfies the desired properties.
  \end{proof}

  \begin{theorem}
    \label{th:infoBeyonCKO}
    There exists a claim that holds and is anonymous and informative beyond counting iff $\set{P_i, i \in I} \neq P \cup \set{\emptyset}$.
  \end{theorem}
  \begin{proof}
    From \cref{th:HAinformativeness} with $Q = P \cup \set{\emptyset}$, we have that 
    there exists a claim that holds and is anonymous and satisfies 
    $\exists j \in I \suchthat \pi(C)(j) \ssub P \cup \set{\emptyset}$ iff
    $\set{P_i, i \in I} \ssub P \cup \set{\emptyset}$.
  \end{proof}

  \begin{theorem}
    \label{th:strinfoBeyonCKO}
    There exists a claim that holds and is anonymous and strongly informative beyond counting iff $\set{P_i, i \in I} \ssub P$.
  \end{theorem}
  \begin{proof}
    From \cref{th:HAinformativeness} with $Q = P$, we have that 
    there exists a claim that holds and is anonymous and satisfies 
    $\exists j \in I \suchthat \pi(C)(j) \ssub P$ iff
    $\set{P_i, i \in I} \ssub P$.
  \end{proof}

  \begin{theorem}
    \label{th:infoBeyonCKOext}
    Given $Q = \set{\phi, ¬\phi}$,
    the following two conditions on $\allleadsto$ are equivalent:
    there exists a claim that holds and is anonymous and satisfies 
    $\exists j \in I \suchthat \pi(C)(j) \ssub Q$;
    and 
    $\set{P_i, i \in I} \ssub Q$.
  \end{theorem}
  \begin{proof}
    From \cref{th:HAinformativeness} with $Q = \set{\phi, ¬\phi}$, we have that 
    there exists a claim that holds and is anonymous and satisfies 
    $\exists j \in I \suchthat \pi(C)(j) \ssub \set{\phi, ¬\phi}$ iff
    $\set{P_i, i \in I} \ssub \set{\phi, ¬\phi}$.
  \end{proof}

  \begin{theorem}
    \label{th:suffIndic}
    Assume that for some $H \in \finite(\incr(\allhists))$ and $i_1 \in I \suchthat P_{i_1} \in P$, we have $\forall i_2 \suchthat P_{i_2} \neq P_{i_1}$ that $\exists \hist \in H \suchthat {\allleadsto}(\hist, i_1) \neq {\allleadsto}(\hist, i_2)$.
    Then, there exists a claim that holds and is anonymous and indicative.
  \end{theorem}
  \begin{proof}
    Define $p = P_{i_1}$.
    Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}$.
    The claim is indicative for $i = i_1$ and $\theta = \set{(\hist, i_1, {\allleadsto}(\hist, i_1)) \suchthat \hist \in H}$ for the following reason.
    The attitudes in $C$ compatible with $\theta$ are $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}})(i_1) = p$.
  \end{proof}
  \begin{corollary}
    \label{th:suffIndicDec}
    Assume that for some $p \in P$, we have $\exists i \in I \suchthat P_i = p$ and $\forall i_2 \suchthat P_{i_2} \neq p$ we have that $\exists A \in \finite(\allargs) \suchthat \forall i \in I \suchthat P_i \neq p, \exists a \in A \suchthat {\ileadstost}(a) \in P \setminus \set{p}$, then there exists a claim that holds and is anonymous and indicative.
  \end{corollary}
  \begin{proof}
    We know that $i_1$ has some decisive argument $a_1$ in favor of $p$.
    Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
    Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist_i = (\hist_{i - 1}, a_1, i)$.
    Define $H = \set{\hist_i \suchthat i \in \intvl{2, \card{A} + 1}}$.
    Apply \cref{th:suffIndic}.
  \end{proof}

  \section{Theories of deliberated choice}
  This section assumes that an exposure protocol $(I, \allargs, {\allleadsto})$ is defined, and defines theories of deliberated choice associated with that exposure protocol. We will start with very simple theories (\cref{sec:static}) in order to build an intuition about the conditions required on theories of deliberated choice, then justify the need for more complex theories (\cref{sec:lichtenstein}) and finally define more general theories of deliberated choice (\cref{sec:dynamic}).

  \subsection{Theories using a static decisive argument}
  \label{sec:static}
  \Cref{ex:tutor} suggests a simple form for theories of deliberated choice: suffice to determine an argument $\ar \in \allargs$ and a preference $p \in P$ such that the theory claims that $\forall i \in I: \ar \ileadstost p$ and that $\forall i \in I: p \in P_i$.

  \begin{definition}[Static theory]
    \label{def:static}
    A static theory is a pair $(p, \ar)$ for some $p \in P$ and some $\ar \in \allargs$.
  \end{definition}
  A static theory $(p, \ar)$ is said to be truthful iff $\forall i \in I: \ar \ileadstost p$, and to be correct iff $\forall i \in I: p \in P_i$.
  We can then deduce that such theories are valid, meaning that their truthfulness implies their correctness. Let us make this (very simple) result stand out as a theorem, because a parallel will be usefully made with more complicated theories.
  %Note that this result immediately follows from \cref{ax:decisive}.
  \begin{theorem}[Static theories are valid]
    Any truthful static theory is correct.
  \end{theorem}
  \begin{proof}
    This follows immediately from \cref{def:decisive}.
  \end{proof}
  Importantly, truthfulness can indeed be checked using our observational basis, that is, if $\exists i \in I \suchthat \ar \nileadstost p$, then it is possible to exhibit this fact experimentally.

  Such theories have the merit of simplicity, but there is a need to design more adaptive theories: in many interesting cases, no single argument will be considered decisive by all individuals in $I$. We then need theories that argue with individuals by reacting to what they consider good arguments. The following section illustrates this need.

  \subsection{Some decision problems with no single decisive argument}
  \label{sec:lichtenstein}
  One ambition of the theories to be proposed here is to permit exhibiting the existence of a deliberated consensus among a large set of persons about issues that involve numerous arguments and counter-arguments.

  In many decision problems, individuals will not be attracted by the same set of arguments and counter-arguments, even though they may finally settle to the same preference. Imagine that some individual $i_1$ considers an argument $a_1$ for $\nprop$ relevant and some individual $i_2$ considers an argument $a_2$ for $\nprop$ relevant, but $i_1$ does not consider $a_2$ relevant and $i_2$ does not consider $a_1$ relevant (for example, because they already have some counter-arguments in mind to those arguments). Imagine further that there is some counter-argument $a'_1$ that can be given to $i_1$ to rebut $a_1$, and a different counter-argument $a'_2$ that can be given to $i_2$ to rebut $a_2$, so that they both have $\prop$ in their deliberated judgment when considering all these arguments.
  In such situation, static theories as defined in \cref{sec:static} are ill suited: these theories need to present the same argument to all individuals and would thus need to define a bigger argument $a$ containing both $a'_1$ and $a'_2$. Such an argument could be too long to be convincing to $i_1$ or $i_2$, as it needlessly spends efforts countering an argument that the individual considers irrelevant to start with. This  will become an important problem if there are not only two but many a priori possible arguments $a_k$ for $\nprop$, a small subset of which is considered relevant by any given individual, each with their own counter-argument $a'_k$ for $\prop$.
  This section is devoted to describing two examples to illustrate that point. Bear in mind that these examples may look somehow naïve on the aspect of the hypothesis they make about what convinces individuals. Real-world complex decision situations will require complex theories that do not fit the purpose of simple illustrative examples. Their goal is to argue that non-static theories are required, but not to fully describe such theories.

  %For a partial illustration about an ambitious setting, consider a decision to be taken about the relevance of some concrete action to help cope with climate change, and let $I$ be a set of citizens interested in the climate issue. The complete set of all arguments that will be considered relevant by at least one individual may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ against the proposed concrete action, it is possible to give counter-arguments $A'$ that will convince the individual to opt for the proposal.
  One of the settings discussed by \citet{thaler_nudge_2009} permits such an illustration. %to illustrate this point more concretely.
  Let $I$ be a set of workers interested in establishing a savings plan.
  The authors claim that workers tend to not save enough because of psychological biases.
  It is therefore tempting to conjecture that the deliberated choice of those workers is to save more, in the sense defined here, than what their natural choice (the one revealed in absence of arguments) leads to.
  A difficulty related to this decision problem is that the complete set of all arguments that will be considered relevant by at least one individual for such a complex issue may well be very large. It may also be that, for each individual who considers relevant some arguments $A$ for saving not much (i.e., spending more in the short term), it is possible to give counter-arguments $A'$ that will convince the individual to opt for saving more.
  But no single argument could be decisive for each $i \in I$: it could be that only counter-arguments built specifically according to what a given individual considers relevant could convince her, given the complexity of the issue.
  This illustrates that the deliberated choice of each individuals in $I$ may be to save more even though no static theory (\cref{def:static}) may achieve showing it, as it could be that no such theory is truthful.

  %The above setting illustrates the ambition of this research program, but at the price of rendering impossible in the confines of this article to define fully the set of arguments and giving a complete account of a concrete theory of deliberated choice in such a complex setting, an endeavor of its own.

  The following example illustrates this same point in a constrained setting which permits a complete definition of $\allargs$ and $\allleadsto$.%, which permits defining the setting completely at the price of a reduced realism.
  \begin{example}[A decision problem about two lotteries]
    \label{ex:lichtenstein}
    In a famous article about preference reversals, \citet{lichtenstein_reversals_2006} discuss an example involving two lotteries. The lottery $L_P$ lets the subject win four dollars with 99\% chance and lose ten cents with 1\% chance; in summary, $L_P = (.99, +\$4; .01, −\$0.1)$. The lottery $L_D$ is, using the same notations: $(.33, +\$16; .67, −\$2)$. The letter $P$ stands for probability: one salient aspect of $L_P$ is the very high chance of winning. The letter $D$ stands for dollar: one salient aspect of $L_D$ is the high gain it gives (compared to $P$) in dollar amount if the subject is lucky.

    This example is interesting because when asked to bid for lotteries, the theory discussed by the authors predicts, and their experiments confirm, that subjects tend to bid more for $L_D$ than for $L_P$, because the bidding question tend to make individuals focus on dollar amounts; whereas when asked to choose one lottery, the lotteries being presented in pairs, subjects tend to choose $L_P$ among the pair $\set{L_P, L_D}$, giving greater attention to probabilities.

    Define $P = \set{L_P, L_D, 0}$.

    Define $\ar_e$ as a text arguing that the choice between $L_P$ and $L_D$ should be determined by their expected revenue, namely, $\$3.959$ for $L_P$ and $\$4$ for $L_D$.%
    \footnote{Here is a possible text for $\ar_e$. “Imagine you play $L_P$ 100 times. On average, you would win 99 times and lose once, which means gaining 99 × \$4 = \$396 and losing 1 × \$0.10 = \$0.10, hence a total net gain of \$395.90. Thus, your average net gain for the $L_P$ bet would be \$3.9590. [Similar text for $L_D$.] Therefore, we suggest you choose $L_D$ for a higher expected revenue.”}
    Define $\ar_u$ as a text arguing that the utility of money should matter, not the revenue, and arguing that a loss hurts more than a gain.% 
    \footnote{Here is a possible text for $\ar_u$. “$L_D$ is better than $L_P$ in terms of expected gain, but it is in general absurd to reason in terms of expected gain. For example, consider a choice between winning $10^6$ dollars with probability $0.8$ and winning $2 × 10^6$ dollars with probability $0.4$. These two lotteries have the same expected gain, but the first one is clearly preferrable. Therefore, we recommend to rather pick $L_P$ as it presents a much better chance of earning money”.}
    Define a counter-argument $\ar_l$ to $\ar_u$, indicating that for small monetary amounts, it is perfectly reasonable to consider utility as linear, especially given that in this experiment, multiple lotteries will be played.%
    \footnote{Here is a possible text for $\ar_l$. “The lottery $L_D$ has a higher expected revenue, and thus, a higher utility than $L_P$, because for small monetary amounts, it is perfectly reasonable to consider utility as linear, especially given that in this experiment, multiple lotteries will be played.”}
    Define $\ar$ so as to include the contents of $\ar_e$, $\ar_u$ and $\ar_l$.%
    \footnote{Here is a possible text for $\ar$. [Text of $\ar_e$] “Note that one could say that the utility of money should matter, but for small monetary amounts, it is perfectly reasonable to consider utility as linear, especially given that in this experiment, multiple lotteries will be played.”}
    Define $\allargs = \set{\ar_e, \ar_u, \ar_l, \ar}$.

    Define $I$ as a set of individuals facing a choice between $L_P$ and $L_D$ in a laboratory experiment.
    Define $\ileadsto$ as exposing $i$ to $\alpha$ and observing the resulting choice by asking $i$ to indicate which lottery she prefers playing, or declare being indifferent.
    %Given an argumentative path $\hist$, $\hist \ileadsto (P ≥ D)$ iff, in a situation of $i$ having been exposed, in order, to the arguments in $\hist$, letting $i$ choose between $P$ and $D$ or declare indifference leads $i$ to choose $P$, or declare indifference. Here is a possible text presented to $i$ to ask for a choice: “Considering these arguments, would you please choose either to play $P$, or play $D$, or declare to be indifferent (in which case one will be picked randomly for you)”. 
    %Similarly, define $\hist \ileadsto (P > D)$ iff showing the arguments in $\hist$ to $i$ and letting $i$ choose between $P$ and $D$ or declare indifference leads $i$ to choose $P$ and not $D$. 
    %Define similarly $\hist \ileadsto \prop$ for $\prop \in \set{D ≥ P, D > P}$.

    Assume that some individuals $i \in I$ do not see the point of argument $\ar_u$ and some do but are convinced by the counter-argument $\ar_l$. Accordingly, define $\ileadsto[e]$, representing the first kind of individuals, as follows. $\forall \hist \in \allhists: \hist \ileadsto[e] L_D ⇔ [\ar_e \in R(\hist)]$, and $\hist \ileadsto[e] L_P$ otherwise. Note that in particular, if $\hist$ contains $\ar$ but not $\ar_e$, $\ar \ileadsto[e] L_P$, as we assume such individuals are confused by the longer argument $\ar$ referring to abstract notions and therefore opt for $L_P$, falling back to the common psychological bias.
    The second kind of individuals is represented by $\ileadsto[u]$ defined as $\forall \hist \in \allhists: \hist \ileadsto[u] L_D ⇔ [\ar_u \notin R(\hist) \lor \ar \in R(\hist) \lor \set{\ar_u, \ar_l} \subseteq R(\hist)]$, and $\hist \ileadsto[u] L_P$ otherwise.
    We assume here that such individuals think about $\ar_u$ as a correct argument in favor of $L_P$ until shown the rebuttal $\ar_l$ (or $\ar$, which contains a similar rebuttal).
    Assuming finally $\allleadsto$ as equal to $\ileadsto[e]$ for individuals of the first kind and $\ileadsto[u]$ otherwise concludes the definition of the exposure protocol $(I, \allargs, {\allleadsto})$.

    It follows from this definition that $\forall i \in I: L_D \in P_i$, and that no static theory of that exposure protocol is truthful: only $\ar_e$ is a decisive argument for $L_D$ for the first kind of individuals whereas for the second kind, only $\ar$ is.
  \end{example}
  This example illustrates the need for theories that are able to react to arguments depending on what a given individual considers relevant. This is the object of the next section.

  \subsection{Form of theories (draft)}
  Given an exposure protocol $(I, \allargs, {\allleadsto})$, we want to phrase some claims under the form of a logical sentence $C$ such that:
  if $C$ holds, then it implies something about the dj of some $i$;
  if $C$ fails, it is possible to falsify it.

  We can observe set $O$ that forms a function $f(i) = \set{\alpha \ileadsto \phi, \alpha' \ileadsto ¬\phi, …}$, with $\alpha \subset \alpha \subset …$ and with finite $O$.
  Our second condition is that $¬C \implies \exists θ \suchthat θ \implies ¬C$.
  For example, if $C$ is $\alpha \ileadsto \phi \lor \beta \allleadsto[j] \phi$, $O$ is $\alpha \ileadsto ¬\phi \land \beta \allleadsto ¬\phi$, then $O$ implies $¬C$.
  As another example, $C$ being $\exists \alpha \suchthat \alpha \ileadsto \phi$ is unfalsifiable, as its negation is $\forall \alpha: \alpha \nileadsto \phi$, which requires infinite $O$.

  Consider $\allleadsto$ such that $\exists \ar, \phi \suchthat \forall i: \ar \ileadsto \phi$.
  Then some suitable $C$ exists.
  It is $\forall i: \ar \ileadstost \phi$.
  If false, it is falsifiable.
  Even though it is not provable by any $O$.

  Now consider $\allleadsto$ such that $\exists J, K \subset I \suchthat \ar \ileadstost[j] \phi, \ar' \ileadstost[k] \phi$.
  Then no suitable $C$ exists?

  If $C$ has the form, whenever $a$ fails, $a'$ is then decisive, it is falsifiable, but does not lead to a decisive argument.
  Want to show: $j \ileadstost \phi$. Need to show: $ar$ or $ar'$ is decisive.
  Assume we know who is a $j$. Then we can!
  Claim: $\ar \ileadsto \phi$.
  Now assume that our claim has to be anonymous.
  Also, assume the formulas involve one $i$ at a time (not needed, probably, but not much restrictive so okay for now).
  So it is of the form $(i, \alpha, \phi) \lor (i, \alpha, \phi') \land (i, \alpha', \phi)…$ implies something decisive.
  So it must be of the form: all $\alpha$ ending with $\ar$ do $\ileadsto \phi$. Or must imply this.
  Let’s make it very strong, as much as possible, thus equal to $\ileadsto \cap \ileadsto[j]$.
  That’s not strong enough!

  Escaping?
  Assume that we know that 50\% are J and 50\% are K.

  \section{Technical notes about falsification}
  \begin{definition}
    \label{def:fals}
    A claim $C$ is \emph{falsifiable} iff
    $C \neq {\allleadstos} \land
    \forall {\allleadstog} \notin C: \exists θ \in \Theta \cap \powerset{\allleadstog} \suchthat \theta \notin C^\wsub$.
  \end{definition}
  \begin{definition}
    \label{def:unfals}
    A claim is \emph{unfalsifiable} iff
    $C = {\allleadstos} \lor
    \exists {\allleadstog} \notin C \suchthat \Theta \cap \powerset{\allleadstog} \wsub C^\wsub$.
  \end{definition}

  A claim $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \notin C: \exists θ \subseteq {\allleadstog} \suchthat \theta \notin C^\subseteq$,
  equivalently iff $\exists {\allleadstog} \notin C \suchthat \forall \theta \in \Theta, \theta \subseteq {\allleadstog}: \theta \in C^\subseteq$.
  (To be corrected for the trivial claims.)

  \subsection{Comparison with rejection}
  The following theorem (whose condition is that $C$ may fail to hold and still be not rejectable) is a consequence of the contrapositive of \cref{th:holdsiff}.
  \begin{corollary}
    \label{th:holdsiffCorr}
    A claim $C$ that satisfies $\exists {\allleadstog} \notin C$ such that $C$ is not rejectable given $\allleadstog$ is unfalsifiable and not trivial.
  \end{corollary}
  The following theorem (an immediate consequence of \cref{def:unfals}) is the converse of \cref{th:holdsiffCorr}.
  It shows that falsifiability or triviality is exactly the property that we need to fill the gap.
  \begin{theorem}
    \label{th:holdsiffConv}
    For any unfalsifiable and not trivial claim $C$:
    $\exists {\allleadstog} \notin C$ such that $C$ is not rejectable given $\allleadstog$.
  \end{theorem}
  \begin{remark}
    It is interesting to note that strong unfalsifiability entails necessary non rejection whereas strict unfalsifiability entails possible non rejection. Perhaps we should exploit this by showing that a given relation not in C entails non rejection using that relation, or something like that?
  \end{remark}

\subsection{Coverability}
\begin{definition}
  \label{def:covr}
  A claim $C$ is \emph{coverable} iff $\forall \theta \in \Theta \cap C^\wsub: \exists \theta' \in \Theta \cap C^\wsub \suchthat \theta \wsub \theta' \land \set{{\allleadstog} \in \allleadstos \suchthat \theta' \in \powerset{\allleadstog}} \wsub C$.
\end{definition}

\begin{theorem}
  \label{th:extStronglyU}
  For any coverable claim $C$ and strongly unfalsifiable claim $C'$:
  $\Theta \cap C^\wsub \wsub (C' \cap C)^\wsub$.
\end{theorem}
\begin{proof}
  Consider any $\theta \in \Theta \cap C^\wsub$.
  Let us show that $\theta \in \cup_{{\allleadstog} \in (C' \cap C)} \powerset{\allleadstog}$.

  Consider any $\theta' \in \Theta \cap C^\wsub \suchthat \theta \wsub \theta' \land \set{{\allleadstog} \in \allleadstos \suchthat \theta' \in \powerset{\allleadstog}} \wsub C$.
  As $\Theta \wsub {C'}^\wsub$, we have $\theta' \in \cup_{{\allleadstog} \in C'} \powerset{\allleadstog}$.
  Pick any such $\allleadstog$ (thus with $\theta' \in \powerset{\allleadstog}$ and ${\allleadstog} \in C'$). Thus, $\theta \in \powerset{\allleadstog}$.
  Also, as $\set{{\allleadstog} \in \allleadstos \suchthat \theta' \in \powerset{\allleadstog}} \wsub C$, we have ${\allleadstog} \in C$.
\end{proof}
A claim $C$ is not coverable iff $\exists \theta \in \Theta \cap C^\wsub \suchthat \forall \theta' \in \Theta \cap C^\wsub, \theta \wsub \theta': \exists {\allleadstog} \notin C \suchthat \theta' \in \powerset{\allleadstog}$.

A claim $C$ is not coverable because of the empty set iff $\forall \theta \in \Theta \cap C^\wsub: \exists {\allleadstog} \notin C \suchthat \theta \in \powerset{\allleadstog}$.

\begin{remark}
  We do not have that when $C_1 \cap C_2 \neq \emptyset$, then $\theta \in C_1^\wsub \land \theta \in C_2^\wsub \implies \theta \in (C_1 \cap C_2)^\wsub$.
  For a counter-example, consider $\hist, \hist_1, \hist_2 \in \allhists$, $i, i_1, i_2 \in I$ and $p \neq p', p_1, p_2 \in P$ and $C_1 = \set{{\allleadstog} \in \allleadstos \suchthat (\hist_1 \ileadstog[i_1] p_1 \implies \hist \ileadstog p) \land (\hist_2 \ileadstog[i_2] p_2 \implies \hist \ileadstog p')}$, $C_2 = \set{{\allleadstog} \in \allleadstos \suchthat (\hist \ileadstog p \implies \hist_2 \ileadstog[i_2] p_2)}$ and $\theta = \set{(\hist_1, i_1, p_1)}$.
\end{remark}
\begin{remark}
  The above remark extends to the case where $C_1$ is strongly unfalsifiable.
  For a counter-example, consider $C_1 = \set{{\allleadstog} \in \allleadstos \suchthat (\ar) \ileadstog \phi \lor (\ar' \ileadstog \phi)}$ and $C_2$ containing a condition requiring $(\ar) \nileadstog \phi \land (\ar' \nileadstog \phi)$.
  Add implications if desired to make $C_1$ and $C_2$ partially compatible, for example, $C_1 = \set{{\allleadstog} \in \allleadstos \suchthat (\alpha \ileadstog[j] p) \implies [(\ar) \ileadstog \phi \lor (\ar' \ileadstog \phi)]}$, so that $C_1$ and $C_2$ require their condition only when some trigger holds.
  Note that the $C_2$ condition may be arbitrarily complex so I see no hope of caracterizing these situations.
\end{remark}
\begin{remark}
  Apparently (from \cref{ex:partlyU}) we can have $C_1 \cap C_2 \neq \emptyset$ and $\overline{C_1} \cap C_2 \neq \emptyset$ but still $C_1^\wsub \cap C_2^\wsub \wsub (C_1 \cap C_2)^\wsub$?
  Though the above remark suggests that non inclusion will yield some similar incompatibility.
\end{remark}

\subsection{Extensions}
\begin{theorem}
  \label{th:struExt}
  If $C$ is strongly unfalsifiable, then any wider claim $C' \supseteq C$ is strongly unfalsifiable.
\end{theorem}
\begin{proof}
  $\Theta \wsub C^\wsub \wsub {C'}^\wsub$.
\end{proof}

\begin{theorem}
  \label{th:partlyU}
  If $C = C_1 \cap C_2$ with $C_1$ strongly unfalsifiable and $C_2$ coverable, then $C$ is unfalsifiable.
\end{theorem}
\begin{proof}
  From $C_1$ being strongly unfalsifiable and $C_2$ being coverable, we obtain
  $\Theta \cap C_2^\wsub \wsub (C_1 \cap C_2)^\wsub$ (\cref{th:extStronglyU}).

  Pick any ${\allleadstog} \in C_2$ with ${\allleadstog} \notin C_1$ (thus ${\allleadstog} \notin C$) and any $\theta \in \Theta \cap \powerset{\allleadstog}$,
  and let us show that $\theta \in C^\wsub$.

  From $\theta \sub {\allleadstog}$ we have $\theta \in C_2^\wsub$.
  Thus, $\theta \in (C_1 \cap C_2)^\wsub = C^\wsub$.
\end{proof}
\begin{theorem}
  \label{th:strongUToStrictU}
  If $C = C_1 \cap C_2$ with $C_1$ strongly unfalsifiable and $C_2$ falsifiable and coverable, then $C$ is unfalsifiable but not strongly unfalsifiable.
\end{theorem}
\begin{proof}
  The claim $C_2$ is falsifiable, thus, is not strongly unfalsifiable.
  Thus, the more precise claim $C$ is not strongly unfalsifiable (\cref{th:struExt}).
  \Cref{th:partlyU} concludes.
\end{proof}

\subsection{Draft material about falsification}
\begin{remark}
  Any relation $\allleadstog$ is equivalent to $\cup \set{\restr{\allleadstog}{O} \suchthat O \in \cup_{J \subseteq_\finite I} (\allhists)^J}$.
\end{remark}

\begin{remark}
  Any relation $\allleadstog$ is equivalent to $\cup \set{\restr{\allleadstog}{O} \suchthat O \in (\finite(\incr(\allhists))^I)}$.
  Conversely, any set $S$ of observations such that $\set{\theta^{-1}(P) \suchthat \theta \in S} = \incr(\allhists)^I$ and $\cup S$ is right-unique corresponds to a unique relation $\cup S = {\allleadstog}$.
  Thus, there is a bijection between the set $\mathcal{S}$ of such sets $S$ and the set of leads to relations.

  This is the intuitive reason that, as we will see,
  a claim $C$ is falsifiable iff $\forall S \in \mathcal{S}: S \subseteq C^\subseteq \implies \cup S \in C$.
\end{remark}

\begin{definition}
  A claim $C$ is necessarily verifiable iff $\forall {\allleadstog} \in C$: $\exists \theta \in \Theta \cap \powerset{\allleadstog} \suchthat \set{{\allleadstog'} \in \allleadstos \suchthat \theta \in \powerset{\allleadstog'}} \wsub C$.
\end{definition}

\begin{example}[Bigger disjunct, strongly unfalsifiable]
  \label{ex:biggeror}
  Pick any $i \in I$ and $\ar \in \allargs$.
  The claim $[(\ar) \ileadstog \phi] \lor [(\ar, \ar) \ileadstog \phi] \lor …$,
  thus, $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat (a)^n \ileadstog \phi}$,
  is strongly unfalsifiable.

  Let $C' = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall j: (a)^n \ileadstog[j] \phi}$ denote the claim of \cref{ex:bigor}.
  We have $C' \subseteq C$.
  It follows that ${C'}^\wsub \subseteq C^\wsub$.
  We know from \cref{ex:bigor} that $\Theta \wsub {C'}^\wsub$, thus, $\Theta \wsub C^\wsub$.
\end{example}

\begin{remark}
  I think that \cref{th:extStronglyU} is not true when replacing the assumption that $C$ is easy with assuming that $C$ is necessarily verifiable.

  Assume that $C$ is $\hist_1 \ileadstog p_1 \lor (\hist_2 \ileadstog p_2 \land \hist_3 \ileadstog p_3)$.
  Say that $\theta = {(\hist_2 \ileadstog p_2), (\hist_4 \ileadstog p_4)}$.
  Now we could have that together with $\hist_3$, this is not observable.
  And together with $\hist_3$ it could be incompatible with $C'$: we only have guaranteed compatibility with observables.
  % 
  % From $C$ being verifiable we know that for some $\theta^\mathit{cover} \in \Theta: \set{{\allleadstog} \in \allleadstos \suchthat \theta^\mathit{cover} \in \powerset{\allleadstog}} \wsub C$.
  % 
  % From $\Theta \wsub {C'}^\wsub$
  % we know that
  % $\theta \in \cup_{{\allleadstog} \in C'} \powerset{\allleadstog}$.
  % Pick any ${\allleadstog} \in C'$ such that $\theta \in \powerset{\allleadstog}$.
  % 
  % If $\theta^\mathit{cover} \wsub \theta$, then by $\theta \in \powerset{\allleadstog}$, $\theta^\mathit{cover} \in \powerset{\allleadstog}$,
  % whence ${\allleadstog} \in C$.
  % 
  % If $\theta \wsub \theta^\mathit{cover}$, using $\Theta \wsub {C'}^\wsub$ whence $\theta^\mathit{cover} \in \cup_{{\allleadstog} \in C'} \powerset{\allleadstog}$, pick any ${\allleadstog} \in C'$ such that $\theta^\mathit{cover} \in \powerset{\allleadstog}$ (hence $\theta \in \powerset{\allleadstog}$) and get ${\allleadstog} \in C$ as previously.
  % For the missing case (no inclusion relation), I probably need $C$ to be necessarily verifiable (and anyway the very notion of verifiability only makes sense for $\allleadsto$), but even that seems insufficient.
  % Consider $C = \set{{\allleadstog} \in \allleadstos \suchthat \hist_1 \ileadstog p_1 \lor (…)}$. It is verifiable thanks to $\theta = \set{(\hist_1, i, p_1)}$ but this does not help with an observation that is compatible with $C$ using the other disjunct.
\end{remark}

Can I make a strictly unfalsifiable claim more precise and obtain a strongly unfalsifiable claim?

\begin{remark}
  The claim “For each glass, it will not break on a monday or not break on a tuesday, whatever happens to it.” is not falsifiable. But one can make it stronger and falsify it, by giving a general theory of how glasses break, proving that it holds, and that it is not compatible with the claim. I should think about this.
\end{remark}

\begin{conjecture}
  $C$ is falsifiable iff exists two compatible paths in two different such forests, for a given $i$, that are not in a single forest, for that $i$, of $C$.
  This seems incorrect; need to be able to build a union relation!
  Prove that it is false with claim "i goes to phi OR i goes to not phi except for empty set going to phi"?
  Actually I need to be able to merge two relations that agree on everything but some part of the tree, in such a way that all paths crossing the boundary must be in $C$, but not the merge.
  TOo be done: caract with singleton $I$.
  % Given a claim $C$, define $S = C^\wsub \cap \Theta$.
  % We have: $\set{\cup_i(\theta_i) \in \allleadstos \suchthat (\theta_i) \in s_S} = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S}$.
\end{conjecture}
\begin{remark}
  Say that $(r1, o1)$ is determined in $C$ (to $p$) iff all relations in $C$ including $r1$ agree on ${\allleadstog}(o)$ (satisfy ${\allleadstog}(o) = p$).
  Let $C(r, o)$ equal $\set{{\allleadstog}(o) \suchthat {\allleadstog} \in C \land r \in {\allleadstog}}$.

  A claim satisfying, $\forall {\allleadstog} \in C$:
  $\forall r1 = (o1, p1) \in {\allleadstog}$, $\exists o2$ with $o1$ and $o2$ obs-compatible and $(r1, o2)$ determined in $C$ such that $\forall o3: (o2, o3)$ is obs-compatible and $(r2 = (o2, C(r1, o2)), o3)$ is determined in $C$
  %Perhaps can omit $(r1, o2)$ determined. In fact, has to satisfy that these determine only relations in $C$.
  % \[\forall (\hist, i, p) \in {\allleadstog}, \exists (\histp, j, p) \in {\allleadstog} \suchthat \forall (\histpp, k) \in \allhists × I: \set{(\histp, j), (\histpp, k)} \in O\]
  is falsifiable.

  NOPE.
  Need to say that $C$ is a union of disjoint subsets such that each subset $C'$ include all relations that are specified on a subset of observations (and left free on everything else) which satisfy something like the above condition: all points related to a point that is related to all others.
\end{remark}
\begin{theorem}
  Given $R \subseteq I × \allhists$, say that $C$ is determined on $R$ iff
  % $C = \set{\restr{\allleadstog}{R} \cup \restr{\allleadstog}{R'} \suchthat R' \in (I × \allhists) \setminus R}$.
  $\exists f \in P^R \suchthat C = \set{f \cup g \suchthat g \in P^{(I × \allhists) \setminus R}}$.

  If $\exists R \subseteq I × \allhists$ such that $C$ is determined on $R$ and satisfying $\exists o \in R \suchthat \forall o' \in R: (\set{o, o'} \in \allpriobss)$, then $C$ is falsifiable.
\end{theorem}
\begin{proof}
  Note that as $C$ is closed-under-combinations, this follows from \cref{rq:closedcomb}, but let’s design a more direct proof in hope of weakening the condition.

  Consider any ${\allleadstog}$ obs-compatible with $C^\wsub \cap \Theta$ and let us prove that ${\allleadstog} \in C$.

  % Pick any $o_1 \in R$.
  % Define $r_1 = (o_1, {\allleadstog}(o_1))$, thus with $r_1 \in {\allleadstog}$.
  Pick some $o \in R$ as in the hypothesis.
  Consider any $o' \in R$.
  As $\set{o, o'} \in \allpriobss$ and ${\allleadstog}$ being obs-compatible, we have
  % $\exists {\allleadstog'} \suchthat \restr{\allleadstog}{\set{o, o'}} = \restr{\allleadstog'}{\set{o, o'}}$.
  $\restr{\allleadstog}{\set{o, o'}} \in C^\wsub$.
  Thus, ${\allleadstog}(o') = f(o')$.

  We obtain that $\forall o' \in R: {\allleadstog}(o') = f(o')$, therefore, ${\allleadstog} \in C$.
\end{proof}
\begin{conjecture}[draft]
  Given $R \subseteq I × \allhists$, say that $C$ is determined on $R$ iff
  % $C = \set{\restr{\allleadstog}{R} \cup \restr{\allleadstog}{R'} \suchthat R' \in (I × \allhists) \setminus R}$.
  $\exists f \in P^R \suchthat C = \set{f \cup g \suchthat g \in P^{(I × \allhists) \setminus R}}$.
  Say that $R$ is coverable iff $\exists o_1, …, O_1, … \suchthat \forall o \in O_1: \set{o_1, o} \in \allpriobss \land \forall o \in O_2: \set{o_2, o} \in \allpriobss \land … \land \set{o_i, o_{i + 1}} \in \allpriobss \land \cup O_i = R \land o_i \in R$.

  If $\exists R \subseteq I × \allhists$ coverable and such that $C$ is determined on $R$, then $C$ is falsifiable.
\end{conjecture}
\begin{proof}
  Consider any ${\allleadstog}$ obs-compatible with $C^\wsub \cap \Theta$ and let us prove that ${\allleadstog} \in C$.

  Pick some $o_1 \in R$ as in the hypothesis.
  Consider any $o \in O_1 \cup \set{o_2}$.
  As $\set{o_1, o} \in \allpriobss$ and ${\allleadstog}$ being obs-compatible, we have
  $\restr{\allleadstog}{\set{o_1, o}} \in C^\wsub$.
  As also $o_1, o \in R$, ${\allleadstog}(o) = f(o)$.
  In particular, ${\allleadstog}(o_2) = f(o_2)$.
  Iterate with $o_2$ and $o \in O_2$, and so on.
  We obtain that $\forall o \in R: {\allleadstog}(o) = f(o)$, therefore, ${\allleadstog} \in C$.
\end{proof}
\begin{example}
  Pick $i \in I$ and
  define $o_1 = ((a), i)$,
  $o_2 = (\emptyset, i)$ and
  $o_3 = ((b), i)$.
  Define $R = \set{o_1, o_2, o_3}$.
  Define $f_1, f_2, f_3 \in P^R$ as follows.
  $f_1(o_1) = \phi$,
  $f_1(o_2) = ¬\phi$,
  $f_1(o_3) = \phi$.
  $f_2(o_1) = \phi$,
  $f_2(o_2) = \phi$,
  $f_2(o_3) = ¬\phi$.
  $f_3(o_1) = ¬\phi$,
  $f_3(o_2) = \phi$,
  $f_3(o_3) = \phi$.
  Define $C = \set{f \cup g \suchthat f \in \set{f_1, f_2, f_3} \land g \in P^{(I × \allhists) \setminus R}}$.
  This is the claim of \cref{ex:threeFs}.

  Note that $R$ is covered by $O_1 = \set{o_1}, O_2 = \set{o_2}, O_3 = \set{o_3}$.
  Note that $C$ is determined on $R$.
  Yet, $C$ is not falsifiable.

  Note that
  \begin{equation}
    C^\wsub \cap \Theta = \left\{\theta \in \Theta \suchthat
    \begin{aligned}
       & (\theta(o_1) \in \set{\emptyset, \phi} \land \theta(o_2) \in \set{\emptyset, ¬\phi} \land \theta(o_3) \in \set{\emptyset, \phi}) \lor \mbox{} \\
       & (\theta(o_1) \in \set{\emptyset, \phi} \land \theta(o_2) \in \set{\emptyset, \phi} \land \theta(o_3) \in \set{\emptyset, ¬\phi}) \lor \mbox{} \\
       & (\theta(o_1) \in \set{\emptyset, ¬\phi} \land \theta(o_2) \in \set{\emptyset, \phi} \land \theta(o_3) \in \set{\emptyset, \phi})
    \end{aligned}\right\}.
  \end{equation}
  Also, $\theta(o_1) \neq \emptyset \implies \theta(o_3) = \emptyset$ (and equivalently, $\theta(o_3) \neq \emptyset \implies \theta(o_1) = \emptyset$),
  whence:
  \begin{equation}
    C^\wsub \cap \Theta = \left\{\theta \in \Theta \suchthat
    \begin{aligned}
       & (\theta(o_1) = \theta(o_3) = \emptyset \land \theta(o_2) \in \set{\emptyset, ¬\phi}) \lor \mbox{} \\
       & (\theta(o_1) = \phi \land \theta(o_2) \in \set{\emptyset, ¬\phi}) \lor \mbox{}                    \\
       & (\theta(o_3) = \phi \land \theta(o_2) \in \set{\emptyset, ¬\phi}) \lor \mbox{}                    \\
       & (\theta(o_1) = \theta(o_3) = \emptyset \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}  \\
       & (\theta(o_1) = \phi \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}                     \\
       & (\theta(o_3) = ¬\phi \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}                    \\
       & (\theta(o_1) = \theta(o_3) = \emptyset \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}  \\
       & (\theta(o_1) = ¬\phi \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}                    \\
       & (\theta(o_3) = \phi \land \theta(o_2) \in \set{\emptyset, \phi})
    \end{aligned}\right\}.
  \end{equation}
  Factorizing:
  \begin{equation}
    C^\wsub \cap \Theta = \left\{\theta \in \Theta \suchthat
    \begin{aligned}
       & (\theta(o_1) = \theta(o_3) = \emptyset \land \theta(o_2) \in \set{\emptyset, \phi, ¬\phi})) \lor \mbox{} \\
       & (\theta(o_1) = \phi \land \theta(o_2) \in \set{\emptyset, \phi, ¬\phi}) \lor \mbox{}                     \\
       & (\theta(o_3) = \phi \land \theta(o_2) \in \set{\emptyset, \phi, ¬\phi}) \lor \mbox{}                     \\
       & (\theta(o_3) = ¬\phi \land \theta(o_2) \in \set{\emptyset, \phi}) \lor \mbox{}                           \\
       & (\theta(o_1) = ¬\phi \land \theta(o_2) \in \set{\emptyset, \phi})
    \end{aligned}\right\}.
  \end{equation}

  Note that the complement is as follows.
  \begin{equation}
    \overline{C^\wsub \cap \Theta} = \left\{\theta \in \Theta \suchthat
    \begin{aligned}
       & \theta(o_2) = 0 \lor \mbox{}                                                                                \\
       & \theta(o_2) = ¬\phi \land (\theta(o_1) \in \set{¬\phi, 0} \lor \theta(o_3) \in \set{¬\phi, 0}) \lor \mbox{} \\
       & \theta(o_1) \in \set{¬\phi, 0} \land \theta(o_3) \neq ¬\phi \lor \mbox{}                                    \\
       & \theta(o_1) = \emptyset \lor \mbox{}                                                                        \\
       & \theta(o_3) = \emptyset \lor \mbox{}                                                                        \\
       & \theta(o_3) \in \set{¬\phi, 0} \land \theta(o_1) \neq ¬\phi
    \end{aligned}\right\}.
  \end{equation}
  This seems incorrect: if $\theta = \emptyset$, then it belongs, even though it satisfies $\theta(o_1) = \emptyset$.

  The relations that are obs-compatible with $C^\wsub \cap \Theta$ are the ones that, for $(o_1, o_2, o_3)$, include two $\phi$ and one $¬\phi$, the one that equals all $\phi$, and the one that assigns $o_1$ and $o_3$ to $¬\phi$ and $o_2$ to $\phi$.

  It follows that the relations that are obs-compatible with $C^\wsub \cap \Theta$ and that are in $C^\wsub \cap \Theta$ are the ones that, for $(o_1, o_2, o_3)$, include two $\phi$ and one $¬\phi$; and the relations that are obs-compatible with $C^\wsub \cap \Theta$ and that are not in $C^\wsub \cap \Theta$ are the one that equals all $\phi$ and the one that assign $o_1$ and $o_3$ to $\emptyset$ and $o_2$ to $\phi$.

  Pick any $g \in P^{(I × \allhists) \setminus R}$ and define ${\allleadstog} = \set{(\emptyset, i, \phi), ((a), i, \phi), ((b), i, \phi)} \cup g$.
  This relation is obs-compatible with $C^\wsub \cap \Theta$, but ${\allleadstog} \notin C$.
\end{example}
\begin{conjecture}
  Given $R \subseteq I × \allhists$, say that $C$ is determined on $R$ iff
  $\exists f \in P^R \suchthat C = \set{f \cup g \suchthat g \in P^{(I × \allhists) \setminus R}}$.
  Say that $R$ is coverable iff $\exists k \in \N, (O_l)_{l \in \intvl{1, k}}, (o_l)_{l \in \intvl{1, k}} \suchthat
    \cup_{l \in \intvl{1, k}} O_l = R \land
    \forall l \in \intvl{1, k}: o_l \in O_l \land
    \forall o \in O_l: \set{o_l, o} \in \allpriobss \land
    \forall l \in \intvl{1, k - 1}: \set{o_l, o_{l + 1}} \in \allpriobss$.

  If $\exists R \subseteq I × \allhists$ coverable and such that $C$ is determined on $R$, then $C$ is falsifiable.
\end{conjecture}
\begin{proof}
  Consider any ${\allleadstog}$ obs-compatible with $C^\wsub \cap \Theta$ and let us prove that ${\allleadstog} \in C$.

  Pick some $k \in \N, (O_l)_{l \in \intvl{1, k}}, (o_l)_{l \in \intvl{1, k}}$ as in the hypothesis.
  Consider any $l \in \intvl{1, k - 1}$
  and any $o \in O_l \cup \set{o_{l + 1}}$.
  As $\set{o_l, o} \in \allpriobss$ and ${\allleadstog}$ being obs-compatible, we have
  $\restr{\allleadstog}{\set{o_l, o}} \in C^\wsub$.
  As also $o_l, o \in R$ (per $\cup_{l' \in \intvl{1, k}} O_{l'} = R$), ${\allleadstog}(o) = f(o)$.
  In particular, ${\allleadstog}(o_{l + 1}) = f(o_{l + 1})$.
  Iterate with $o_2$ and $o \in O_2$, and so on.
  We obtain that $\forall o \in R: {\allleadstog}(o) = f(o)$, therefore, ${\allleadstog} \in C$.
\end{proof}
\begin{conjecture}
  Given $R \subseteq I × \allhists$, define ${\sim_R} \subseteq \allleadstos × \allleadstos$ as $f \sim_R f'$ iff $\restr{f}{R} = \restr{f'}{R}$. This is an equivalence relation.
  Given $f \in \allleadstos$, define the equivalence class of $f$ for ${\sim_R}$ as ${\sim_R}(f) = \set{f' \in \allleadstos \suchthat f \sim_R f'} = \set{f' \in P^{(I × \allhists)} \suchthat \restr{f}{R} = \restr{f'}{R}}$.

  Given $R \subseteq I × \allhists$, define the a priori observable subsets of $R$ as $Z_R = \allpriobss \cap \powerset{R}$.

  Consider $R \subseteq I × \allhists$, $F \subseteq P^R$ and $C = \set{f \cup g \suchthat f \in F \land g \in P^{(I × \allhists) \setminus R}}$.
  The claim $C$ is unfalsifiable iff $\exists {\allleadstog} \in \allleadstos \suchthat \bigcap_{O \in Z_R} {\sim_O}(f) \cap C = \emptyset$.

  % Given $k \in \N$ and a family $(O_l)_{l \in \intvl{1, k}}$, consider $\exists f \in C \suchthat 
  Say that $R$ is coverable iff $\exists k \in \N, (O_l)_{l \in \intvl{1, k}}, (o_l)_{l \in \intvl{1, k}} \suchthat
    \cup_{l \in \intvl{1, k}} O_l = R \land
    \forall l \in \intvl{1, k}: o_l \in O_l \land
    \forall o \in O_l: \set{o_l, o} \in \allpriobss \land
    \forall l \in \intvl{1, k - 1}: \set{o_l, o_{l + 1}} \in \allpriobss$.

  If $\exists R \subseteq I × \allhists$ coverable and such that $C$ is determined on $R$, then $C$ is falsifiable.
\end{conjecture}
Consider: extend this so as to cover eg $C$ containing $f_1$ and $f_2$ constant functions as in example; but not covering when the domain is not suitable!
Use proof similar to \cref{ex:for}, getting started with observing that any obs-compatible $\allleadstog$ failing every element but one of $C$ at least at one point must then satisfy the last element of $C$ at some point (?).
\begin{conjecture}
  If $\forall {\allleadstog} \in C: \exists R \suchthat {\allleadstog} = f \cup g$ with $f \in P^R, g \in P^{(I × \allhists) \setminus R}$ and $R$ converable, then $C$ is falsifiable.
  Should be specified better, but also, not sure that this holds: consider C being a leads i to phi OR b leads i to phi (given i).
  Perhaps something like $\exists F \suchthat C = \cup_{f \in F} \set{f \cup g}$ where $f^{-1}(P)$ being coverable.
\end{conjecture}
\begin{proof}
  Problem is: how to get started?
\end{proof}

\begin{example}
  Consider any $i$, $\ar_1 ≠ \ar_2 \in \allargs$, $p ≠ p' \in P$ and any ${\allleadstog^1}, {\allleadstog^2}$ such that $(\ar_1) \ileadstog^1 p$ and $(\ar_2) \ileadstog^2 p'$.
  Consider the claim $C = \set{{\allleadstog^1}, {\allleadstog^2}}$. When is it falsifiable considering these two relations?
\end{example}

Consider: investigate claims that are non falsifiable but with $C^\wsub \cap \Theta \neq \Theta$.
Then exhibit a claim such that a smaller one is not the same status and a bigger one is not the same status.

Any claim containing two relations that differ only on incompatible observations is not falsifiable.
\begin{theorem}
  Any $C$ such that $\exists {\allleadstog^1}, {\allleadstog^2} \in C \suchthat$ some combination of them ${\allleadstog} \notin C$ (meaning that ${\allleadstog} \subseteq {\allleadstog^1} \cup {\allleadstog^2}$) and the set over which they differ, $\set{(i, \hist) \suchthat {\allleadstog^1}(i, \hist) ≠ {\allleadstog^2}(i, \hist)}$, does not contain two elements that are simultaneously observable (meaning that no $θ$ contains more than one of these elements), is not falsifiable.
\end{theorem}
\begin{proof}
  Consider any $θ \subset {\allleadstog}$.
  If $θ$ is fully compatible with ${\allleadstog^1}$, then ${\allleadstog^1} \in S_θ \cap C ≠ \emptyset$.
  Otherwise, $θ$ contains some element $(i, \hist, {\allleadstog^2}(i, \hist))$ with ${\allleadstog^2}(i, \hist) ≠ {\allleadstog^1}(i, \hist)$. Then, ${\allleadstog^2} \in S_θ \cap C ≠ \emptyset$, for otherwise, for at least one other element $(i', \histp)$, $θ$ contains $(i', \histp, {\allleadstog^1}(i', \histp))$ with ${\allleadstog^1}(i', \histp) ≠ {\allleadstog^2}(i', \histp)$, which violates the hypothesis.
\end{proof}
\Cref{ex:cardPart} illustrates the need for the condition about the combination in the theorem.

\begin{conjecture}
  If a claim $C$ is not falsifiable, then $\exists {\allleadstog^1}, {\allleadstog^2} \in C \suchthat$ some combination of them ${\allleadstog} \notin C$ (meaning that ${\allleadstog} \subseteq {\allleadstog^1} \cup {\allleadstog^2}$) and the set over which they differ, $\set{(i, \hist) \suchthat {\allleadstog^1}(i, \hist) ≠ {\allleadstog^2}(i, \hist)}$, does not contain two elements that are simultaneously observable (meaning that no $θ$ contains more than one of these elements).
\end{conjecture}
\begin{conjecture}
  If $\forall {\allleadstog^1}, {\allleadstog^2} \in C \suchthat$ some combination of them ${\allleadstog} \notin C$ (meaning that ${\allleadstog} \subseteq {\allleadstog^1} \cup {\allleadstog^2}$), the set over which they differ, $\set{(i, \hist) \suchthat {\allleadstog^1}(i, \hist) ≠ {\allleadstog^2}(i, \hist)}$, contains at least two elements that are simultaneously observable (meaning that some $θ$ contains more than one of these elements), then $C$ is falsifiable.
\end{conjecture}
\begin{proof}
  We need a stronger hypothesis.
  Consider $one$ and $two$ differ on $(i, (a))$, $(i, (b))$ and $(i, (a, a))$. And some $three$ not in $C$ has $(i, (a)), (i, (a, a))$ as in $one$ and $(i, (b))$ as in $two$. Then $C$ is not falsifiable.
\end{proof}
\begin{conjecture}
  If $\forall {\allleadstog^1}, {\allleadstog^2} \in C \suchthat$ some combination of them ${\allleadstog} \notin C$ (meaning that ${\allleadstog} \subseteq {\allleadstog^1} \cup {\allleadstog^2}$), the set over which they differ, $\set{(i, \hist) \suchthat {\allleadstog^1}(i, \hist) ≠ {\allleadstog^2}(i, \hist)}$, we can find some $(i, \hist) \in {\allleadstog^1}, (j, \hist') \in {\allleadstog^2}$ obs-compatible, then $C$ is falsifiable.
\end{conjecture}
\begin{proof}
  This fails. Could have $one$, $two$, $three$, each sending everything to $p$, $¬p$ and $0$, respectively, and any combination of two of these in $C$, but no comb of three in $C$. Then $four$ a combination of the three ($i, a, p, i, b, ¬p, i, c, 0$), but $C$ not falsifiable. Actually, even if we can observe two among the three (using $i, a, p, i, aa, ¬p, i, b, 0$), we can’t falsifiy $four$.
\end{proof}

\begin{remark}
  Odd length to $p$ and even length to $¬p$. Then we can know that $P_i = \emptyset$.
Or, if $i$ finds $a_1 > a_2 > a_3 > a_1$ (more convincing relation), then $P_i = \emptyset$ and we can know it.
\end{remark}

\subsection{Probably redundant}
NB the following results do not consider trivial claims.

\begin{theorem}
  A claim $C$ is falsifiable iff its complement $\Cbar = \allleadstos \setminus C$ satisfies:
  $\forall {\allleadstog} \in \Cbar, \exists \theta \subseteq {\allleadstog} \suchthat S_θ \subseteq \Cbar$, or equivalently:
  $\forall {\allleadstog} \in \Cbar, \exists O \in (\allhists)^I \suchthat S_{\restr{\allleadstog}{O_O}} \subseteq \Cbar$.
\end{theorem}
\begin{proof}
  The first phrasing follows from the definition of falsifiability.
  Also, the second phrasing implies the first one: pick $\theta = \restr{\allleadstog}{O_O}$.

  Let us prove that the first phrasing implies the second one.
  Given $\theta$, define $O = \set{(i, \argmax_{\hist \in (\theta(i) \cup \emptyset)} \card{\hist})}$ as associating to each $i$ the longest corresponding $\hist$ from $\theta$, or the empty sequence if $\theta(i)$ is empty.
  It follows that $\theta \subseteq \restr{\allleadstog}{O_O}$.
  Thus, $S_{\restr{\allleadstog}{O_O}} \subseteq S_θ$.
  As $S_θ \subseteq \Cbar$, we obtain $S_{\restr{\allleadstog}{O_O}} \subseteq \Cbar$.
\end{proof}

\begin{corollary}
  A claim $C$ is falsifiable iff $\forall f \in C^{\allpriobss} \suchthat \forall O, O' \in \allpriobss: \restr{f(O)}{O \cap O'} = \restr{f(O')}{O \cap O'}$:
  \[\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C.\]
\end{corollary}
\begin{proof}
  Consider a falsifiable claim $C$, consider such an $f$, and let us prove that $\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C$.
  Define ${\allleadstog} \subseteq (I × \allhists) × P$ as $\cup_{O \in \allpriobss} \restr{f(O)}{O}$. Its domain equals $I × \allhists$, and it is right-unique thanks to the condition on $f$.
  Thus, \cref{thm:caractFals} applies and ${\allleadstog} \in C$.

  Conversely, assume that the statement about all $f$ holds and let us prove that $C$ is falsifiable.
  Consider any ${\allleadstog} \in \allleadstos$ such that $\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq$.
  Given any $O$, we have $\restr{\allleadstog}{O} \in C^\subseteq$ and therefore there is some $e_O \in C$ such that $\restr{\allleadstog}{O} = \restr{e_O}{O}$.
  Define $f$ as $f(O) = e_O$.
  As this $f$ satisfies the condition in the hypothesis, we obtain that $\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C$, hence, $\cup_{O \in \allpriobss} \restr{\allleadstog}{O} \in C$.
  Therefore, ${\allleadstog} \in C$.
  Apply \cref{thm:caractFals} to conclude that $C$ is falsifiable.
\end{proof}

\begin{corollary}
  \label{th:corffals}
  A claim $C$ is falsifiable iff $\forall f \in \left(C^\subseteq\right)^{\allpriobss} \suchthat \forall O, O' \in \allpriobss: \restr{f(O)}{O \cap O'} = \restr{f(O')}{O \cap O'}$:
  \[\cup_{O \in \allpriobss} \restr{f(O)}{O} \in C.\]
\end{corollary}
\begin{corollary}
  \label{th:corffalsdom}
  A claim $C$ is falsifiable iff $\forall f \in \cup_{O \in \allpriobss} \set{\set{O} × (P^O \cap C^\subseteq)} \suchthat \forall O, O' \in \allpriobss: \restr{f(O)}{O \cap O'} = \restr{f(O')}{O \cap O'}$:
  \[\cup_{O \in \allpriobss} f(O) \in C.\]
\end{corollary}
\begin{corollary}
  \label{th:veryunfals}
  Any non-trivial claim $C$ satisfying $\Theta \subseteq C^\subseteq$ is not falsifiable.
\end{corollary}
\begin{proof}
  Pick any ${\allleadstog} \notin C$.
  Per hypothesis, $\forall O \in \allpriobss$: $\restr{\allleadstog}{O} \in C^\subseteq$.
  Define $f$ as $\forall O \in \allpriobss: f(O) = \restr{\allleadstog}{O}$.
  Observe that $\cup_{O \in \allpriobss} f(O) = {\allleadstog}$ and hence $\cup_{O \in \allpriobss} f(O) \notin C$.
  Apply \cref{th:corffalsdom}.
\end{proof}

\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)} = C$.
\end{corollary}

A relation $\allleadstog$ is obs-compatible with $C$ iff $\forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq$, equivalently, iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)$.

\begin{remark}
  The question of falsifiability is whether the set of relations $\allleadstog$ “compatible” with $C^\subseteq \cap \Theta$ equals $C$.
  Compatibility needs to consider the observables, not a subset of it.
  Define $\Theta_1 \subset \Theta$ as the observations that observe only one pair $(i, \hist)$.
  Define $R_1 = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta_1: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)}$.
  Note that $R \subseteq R_1$.
  We know that $C$ is falsifiable iff $R \subseteq C$. Thus, $C$ is not falsifiable implies $C \subsetneq R \subseteq R_1$ thus $C \subsetneq R_1$. But we do not have that $C \subsetneq R_1$ implies $C$ not falsifiable: could have $C \subsetneq R_1$ but still $R = C$.
\end{remark}

\hbadness=10000
\bibliography{simple}
\end{document}